### 2021-06-03 README.Rmd
Created a README.md from Rmd file

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "75%"
)
```

# onsetsync - Analysis and Visualisation of Synchronisation of Music Onset Data

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`onsetsync` is a R package for musical dynamics involving synchrony. There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.

## Installation

You can install the current version of `onsetsync` from Github by
entering the following commands into R:

```{r,eval=FALSE}
if (!require(devtools)) install.packages("devtools")
devtools::install_github("tuomaseerola/onsetsync")
```

## Usage

Note that `onsetsync` is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g. Librosa and Python, or Matlab, or Sonic Visualiser) using well-known algorithms. Here we take it as granted that we have extracted the onsets elsewhere and we have the onset times recorded into the csv files.

```{r, results='hide', message=FALSE, warning=FALSE,echo=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(httr)
library(onsetsync)
```

```{r message=FALSE,warning=FALSE,eval=TRUE}
# Onsets
invisible(GET("https://osf.io/8a347/?action=download", 
              write_disk("CSS_Song2_Onsets_Selected.csv", overwrite = T)))
CSS_Song2_Onset<-read.csv("CSS_Song2_Onsets_Selected.csv",header=T,sep=",")
knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple",
             caption = 'Onset data structure (see text for details).')

# Annotations
invisible(GET("https://osf.io/4cdpr/?action=download", 
              write_disk("CSS_Song2_Metre.csv", overwrite = TRUE)))
CSS_Song2_Metre <- read.csv("CSS_Song2_Metre.csv", header=T, sep=",")
knitr::kable(head(CSS_Song2_Metre[1:4,]),format = ,"simple")
CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
                                 Label.SD,SD,Clave,Bass,Guitar,Tres) 
```

As the onsets and annotations are in different files, let's first combine the raw onset and annotation with `onsetsync`. Here we first add annotations about the cycles into the onset data and then we also add isochronous beat times to the data frame, since are useful reference points for future calculations.

```{r message=FALSE,warning=FALSE,eval=TRUE}
# Add annotations about the cycle to the data frame
CSS_Song2 <- add_annotation(df = CSS_Song2_Onset,
                            annotation = CSS_Song2_Metre$Cycle,
                            time = CSS_Song2_Metre$Time,
                            reference = 'Label.SD')
# Add isochronous beats to the data frame
CSS_Song2 <- add_isobeats(df = CSS_Song2, 
                          instr = 'CycleTime', 
                          beat = 'SD')

print(knitr::kable(head(CSS_Song2),format = "simple",digits = 2))
```

Before moving onto the analysis, let's summarise the onset structures in this piece.

```{r message=FALSE,warning=FALSE,eval=TRUE}
tab1 <- summarise_onsets(df = CSS_Song2, 
                         instr = c('Clave','Bass','Guitar','Tres'))
print(knitr::kable(tab1,digits = 1,
     caption = 'Descriptives for the onset time differences (ms)'))
```

s a broad overview, we can visualise the relative synchrony to equal division subdivision of the beat for each instrument across the time. This gives a summary about when the instrument are playing relative to the beat division and the sections of the song.

```{r synch2isochron,fig.width=9, fig.height=7.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
fig1 <- plot_by_beat(df = CSS_Song2, 
                     instr = c('Bass','Clave','Guitar','Tres'), 
                     beat = 'SD', 
                     virtual='Isochronous.SD.Time',
                     pcols=2)
print(fig1)
```

### 2021-06-04 README.Rmd
New function (get_OSF_csv) to simplify getting data from OSF

@@
 
-```{r, results='hide', message=FALSE, warning=FALSE,echo=FALSE}
+```{r, message=FALSE, warning=FALSE,echo=TRUE}
+library(onsetsync)
+library(httr)
 library(dplyr)
 library(ggplot2)
-library(reshape2)
-library(httr)
-library(onsetsync)
 ```
 
-```{r message=FALSE,warning=FALSE,eval=TRUE}
-# Onsets
-invisible(GET("https://osf.io/8a347/?action=download", 
-              write_disk("CSS_Song2_Onsets_Selected.csv", overwrite = T)))
-CSS_Song2_Onset<-read.csv("CSS_Song2_Onsets_Selected.csv",header=T,sep=",")
+```{r message=FALSE,warning=FALSE,eval=TRUE,results='asis'}
+CSS_Song2_Onset <- get_OSF_csv('8a347') # Onsets
 knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple",
@@
 
-# Annotations
-invisible(GET("https://osf.io/4cdpr/?action=download", 
-              write_disk("CSS_Song2_Metre.csv", overwrite = TRUE)))
-CSS_Song2_Metre <- read.csv("CSS_Song2_Metre.csv", header=T, sep=",")
+CSS_Song2_Metre <- get_OSF_csv('4cdpr') # Annotations
 knitr::kable(head(CSS_Song2_Metre[1:4,]),format = ,"simple")

### 2021-06-04 R/get_OSF_csv.R

@@ -1,0 +1,32 @@
+#' Get csv file from OSF
+#'
+#' This function gets CSV file from OSF using httr GET method.
+#'
+#' @param address OSF address code () to be retrieved. Works for CSV files.
+#' @param filename Filename (optional)
+#' @return data
+#' @import httr
+
+get_OSF_csv <- function(address = NULL, filename = NULL) {
+  
+  GET <- write_disk <- NULL
+  # patch together the address
+  combined_address <-
+    paste0('https://osf.io/', address, '/?action=download')
+  
+  # create a temporary filename if needed
+  if (!is.null(filename)) {
+    fn <- filename
+  } else {
+    fn <- tempfile()
+  }
+  
+  # get the file using httr GET method
+  invisible(GET(combined_address, write_disk(fn, overwrite = TRUE)))
+  
+  # read data
+  data <- read.csv(fn, header = T, sep = ",")
+  
+  return <- data
+  
+}

### 2021-06-04 README.Rmd
Made the README more elaborate

@@
 Note that `onsetsync` is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g.
-Librosa and Python, or Matlab, or Sonic Visualiser) using well-known algorithms.
-Here we take it as granted that we have extracted the onsets elsewhere and we have the onset times recorded into the csv files.
+[Librosa](https://librosa.org), or [Mir Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms.
+Here we take it as granted that we have extracted the onsets in some of these programs, probably checked them by hand, and we have the onset times recorded into the csv files.
 
@@
 CSS_Song2_Onset <- get_OSF_csv('8a347') # Onsets
-knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple",
-             caption = 'Onset data structure (see text for details).')
+knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple")
 
 CSS_Song2_Metre <- get_OSF_csv('4cdpr') # Annotations
-knitr::kable(head(CSS_Song2_Metre[1:4,]),format = ,"simple")
 CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
@@
 As the onsets and annotations are in different files, let's first combine the raw onset and annotation with `onsetsync`.
-Here we first add annotations about the cycles into the onset data and then we also add isochronous beat times to the data frame, since are useful reference points for future calculations.
+Here we first add annotations (using `add_annotation` function) about the cycles into the onset data.
+We then add isochronous beat times to the data frame using `add_isobeats`, since these are useful reference points for synchrony calculations.
 
-```{r message=FALSE,warning=FALSE,eval=TRUE}
+```{r message=FALSE,warning=FALSE,eval=TRUE,results='asis'}
 # Add annotations about the cycle to the data frame
@@
 
-```{r message=FALSE,warning=FALSE,eval=TRUE}
+```{r message=FALSE,warning=FALSE,eval=TRUE,results='asis'}
 tab1 <- summarise_onsets(df = CSS_Song2, 
@@
 
-s a broad overview, we can visualise the relative synchrony to equal division subdivision of the beat for each instrument across the time.
-This gives a summary about when the instrument are playing relative to the beat division and the sections of the song.
+As a broad overview, we can visualise the relative synchrony to equal division subdivision of the beat for each instrument across the time.
 
-```{r synch2isochron,fig.width=9, fig.height=7.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
+```{r synch2isochron,fig.width=9, fig.height=7.0}
 fig1 <- plot_by_beat(df = CSS_Song2, 
@@
 ```
+
+There are several variants of this summary, but let's move on.
+
+To what degree are the pairs of instruments synchronised in this example?
+Since the instruments usually play widely different amounts of onsets in a piece, and these are bound to be at different beats sub-divisions, the mutual amount of comparable onsets for each pair varies often dramatically.
+In order to keep the mean and standard deviations comparable, we will randomly sample joint onsets for both instruments.
+
+```{r paired1}
+set.seed(1201) # set random seed
+N <- 200 # Let's select 200 onsets
+d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass',N,1,'SD',TRUE)
+print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
+    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
+```
+Let's visualise the synchrony of all pairings of the instruments in this example.
+
+```{r fig4,fig.width=4.5, fig.height=4.5, warning=FALSE}
+inst<-c('Clave','Bass','Guitar','Tres') # Define instruments 
+dn <- sync_execute_pairs(CSS_Song2,inst,N,10,'SD')
+fig2 <- plot_by_pair(dn)  # plot
+print(fig2)  
+```
+
+For more examples, see [vignette IN PROGRESS](http://) and associated [paper IN PROGRESS](http://).

### 2021-06-04 README.Rmd
Update to dependancies (hms for plots).

@@
 [Librosa](https://librosa.org), or [Mir Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms.
-Here we take it as granted that we have extracted the onsets in some of these programs, probably checked them by hand, and we have the onset times recorded into the csv files.
+Here we take it as granted that we have extracted the onsets in some of these programs, probably checked them by hand, and we have the onset times recorded into csv files.
 

### 2021-06-04 R/plot_by_beat.R

@@ -13,6 +13,7 @@
 #' @return Graphic output
 #' @import ggplot2
 #' @import dplyr
+#' @import hms
 
 plot_by_beat <-
   function(df = NULL,

### 2021-06-04 R/plot_timeline.R

@@ -10,6 +10,7 @@
 #' @return Graphic output
 #' @export
 #' @import ggplot2
+#' @import hms
 
 # T. Eerola, Durham University, IEMP project
 # 22/4/2021

### 2021-06-04 doc/onsetsync.R

@@ -5,5 +5,14 @@ knitr::opts_chunk$set(
 )
 
 ## ----setup--------------------------------------------------------------------
+if (!require(devtools)) install.packages("devtools")
+devtools::install_github("tuomaseerola/onsetsync")
+
 library(onsetsync)
+library(httr)
+library(dplyr)
+library(ggplot2)
+
+CSS_Song2_Onset <- get_OSF_csv('8a347') # Onsets
+knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple")
 

### 2021-06-04 README.Rmd
added docs folder

@@
 print(knitr::kable(tab1,digits = 1,
-     caption = 'Descriptives for the onset time differences (ms)'))
+     caption = 'Descriptives for the onset time differences (ms).'))
 ```
@@
 
-There are several variants of this summary, but let's move on.
+There are several variants of this summary that create different summaries of the onsets across the beats, but let's move on.
 
@@
 
-```{r fig4,fig.width=4.5, fig.height=4.5, warning=FALSE}
+```{r fig4,warning=FALSE}
 inst<-c('Clave','Bass','Guitar','Tres') # Define instruments 
@@
 
-For more examples, see [vignette IN PROGRESS](http://) and associated [paper IN PROGRESS](http://).
+For more examples, see [docs IN PROGRESS](https://tuomaseerola.github.io/onsetsync/) and associated [paper IN PROGRESS](http://).

### 2021-06-04 docs/index.Rmd
Rename docs files to index

--- 2021-06-04 README.Rmd
@@
 ---
-output: github_document
+title: "onsetsync"
+output: rmarkdown::html_vignette
+vignette: >
+  %\VignetteIndexEntry{onsetsync}
+  %\VignetteEngine{knitr::rmarkdown}
+  %\VignetteEncoding{UTF-8}
 ---
-
-<!-- README.md is generated from README.Rmd. Please edit that file -->
 
@@
   collapse = TRUE,
-  comment = "#>",
-  fig.path = "man/figures/README-",
-  out.width = "75%"
+  comment = "#>"
 )
 ```
-
-# onsetsync - Analysis and Visualisation of Synchronisation of Music Onset Data
-
-<!-- badges: start -->
-[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
-<!-- badges: end -->
 

### 2021-06-04 joss-paper/paper.Rmd
Uploaded the first paper draft

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'onsetsync: An R Package for Onset Synchrony and Periodicity Analysis'
tags:
  - R
  - music
  - entrainment
  - periodicity
  - synchrony
authors:
  - name: Tuomas Eerola
    orcid: 0000-0002-2896-929X
    affiliation: 1 # (Multiple affiliations must be quoted)
  - name: Martin Clayton
    orcid: 0000-0000-0000-0000
    affiliation: 1
affiliations:
 - name: Department of Music, Durham University
   index: 1
citation_author: Eerola et. al.
date: 1 July 2021
year: 2021
bibliography: paper.bib
output: rticles::joss_article
#csl: apa.cls
citation_package: "default"
keep_tex: FALSE
journal: JOSS
---

# 1 Introduction and statement of need

Music performance relies on tight yet flexible timing between the performers. This entrainment between two or more performers can be analysed from the note onsets obtained from recorded performances. The onset synchrony characterises the timing accuracy and dynamics from the onsets of the performance. The synchrony between performers is influenced by various factors such as the genre of music, performer skill level, intention, and phrase and beat structures of the music. The analysis of synchrony in music benefits from shared tools as there are a number of common operations that need to be carried in every dataset (e.g., comparing individual onsets to a virtual beat, or assessing the synchrony across other variables such as tempo, metrical hierarchy, or phrasing) and the amount of datasets containing raw onsets has recently increased.

## 1.1 Statement of need

`onsetsync` is a R package for musical dynamics involving synchrony. There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property. These functions will make the analyses of the onset corpora more transparent and comparable and will allow more scholars to carry out exploratory investigation of entraiment in music with minimal skills in computing.

`onsetsync` was designed to be used by both empirical music researchers and by students in courses on music and science and empirical musicology.

> NOTE 1: Here more about synchronization and entrainment. For instance, sensorimotor synchronisation [@repp2008sensorimotor], synchrony indices from Wing, or from @polak2016both. And we need to mention a real world example such as [@lucas2011inter] and cue in Martin's influential papers included as well [@clayton2020;@clayton_et_al_2018b].

And the EMR article [@Clayton2020emr].

> NOTE 2: Different indices for asynchrony, absolute asynchrony, relative asynchrony? We talk about synchrony in all examples, though.

## 1.2 Availability

```{r secret_load,eval=TRUE,echo=FALSE}
# LOADS THIS SECRECTLY FROM THE LOCAL FOLDER
library(onsetsync)
ver <- packageVersion('onsetsync')
```

`onsetsync` is available at Github and can be loaded and installed using the code below. The current version is `r ver`.

```{r availability,eval=FALSE}
library(devtools)
devtools::install_github("tuomaseerola/onsetsync")
library(onsetsync)
packageVersion('onsetsync')
```

# 2 Core functionality

`onsetsync` has five types of functionalities; (1) adding information, (2) summarising onset data, (3) calculating synchrony, (4) estimating periodicity, and (5) visualising all of these. Here we will take example onset data from an open access repository at Open Science Framework (OSF) involving Cuban Son and Salsa performances, annotated and processed as part of [Interpersonal Entrainment in Music Performance](https://musicscience.net/projects/timing/iemp/) project available at [https://osf.io/sfxa2/](https://osf.io/sfxa2/). We will carry out operations that demonstrate all the functionalities of the package.

Note that `onsetsync` is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g. [Librosa](https://librosa.org), or [Mir Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms. Here we take it as granted that we have extracted the onsets in some of these programs, probably checked them by hand, and we have the onset times recorded into csv files.

> NOTE 3: It might be beneficial to be able to replicate some analyses reported in the IEMP papers (Clayton et al 2018 in Musicae Scientiae is the best one, although the Music Perception one could also work) but coordinating this might be a pain (and the results might not be identical).

## 2.1 Read data from an online repository

```{r, results='hide', message=FALSE, warning=FALSE,echo=FALSE}
library(dplyr)
library(ggplot2)
library(reshape2)
library(httr)
```

Retrieve onset data and annotation of metre from an online repository of performances (IEMP at OSF, insert link and refs):

```{r getdata1,message=FALSE,warning=FALSE}
CSS_Song2_Onset <- get_OSF_csv('8a347') # Onsets
knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple")

CSS_Song2_Metre <- get_OSF_csv('4cdpr') # Annotations
CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
                                 Label.SD,SD,Clave,Bass,Guitar,Tres) 
```

The Table 1 shows a portion of the onset data structure where the last three columns refer to the onset times of the specific instruments (`Clave`, `Bass`, and `Guitar`). Note that we only show the first eight columns and 6 rows. The first five columns are meta-data, referring to the `Piece`, beat sub-divisions (`SD` and `Label.SD`, which is a unique label combining cycle and beat information), `Section` indicating the specific part of the piece. Finally, in this example we have `Clave_` column, which is a reference to a clave pattern. The onset times for the instruments are recorded in seconds, although we prefer to use milliseconds in the analyses of synchrony to avoid reporting many digits.

In the second data frame (`CSS_Song2_Metre`) we have annotations, namely the cycle times, which are the exact time when the cycles start (containing all sub-divisions of the beat). This is important information and want to combine the actual onsets and the cycles many analyses.

HERE IT COULD BE EXPLAINED IN MORE DETAILS WHERE THE CYCLE INFORMATION COMES FROM.

> Note 4: We need to have some datasets that with the package to avoid always loading the stuff from OSF.

## 2.2 Combine onsets and annotations

As the onsets and annotations are in different files, let's first combine the raw onset and annotation with `onsetsync`. Here we first add annotations about the cycles into the onset data and then we also add isochronous beat times to the data frame, since are useful reference points for future calculations.

```{r pre,echo=TRUE,results='asis'}
library(onsetsync)
packageVersion('onsetsync')

# Add annotations about the cycle to the data frame
CSS_Song2 <- add_annotation(df = CSS_Song2_Onset,
                            annotation = CSS_Song2_Metre$Cycle,
                            time = CSS_Song2_Metre$Time,
                            reference = 'Label.SD')
# Add isochronous beats to the data frame
CSS_Song2 <- add_isobeats(df = CSS_Song2, 
                          instr = 'CycleTime', 
                          beat = 'SD')

print(knitr::kable(head(CSS_Song2),format = "simple",digits = 2))
```

At this point we have the basic elements for most analyses shown in Table 3; We have the onset data (coded under instrument names, here `Clave`, `Bass`, and `Tres`) and information about the beat cycles  (`CycleTime` and `Cycle`), and we also have timing information for isochronous beat divisions (`Isochronous.SD.Time`).

These virtual beat structures that are now based on annotations can be systematically off by a certain amount, so perhaps some it is advantageous for some analyses to infer the timing information related to the beat structure from the existing onsets themselves and not rely on external reference. `onsetsynch_add_isobeats` can also accomplish this if given the instruments from which it will calculate the mean onset time for the first beat of the cycle, which will be used to set the isochronous beat timings for the rest of the beats within the cycle.

Before moving onto the analysis, let's summarise the onset structures in this piece.

```{r summary_period,results='asis'}
tab1 <- summarise_onsets(df = CSS_Song2, 
                         instr = c('Clave','Bass','Guitar','Tres'))
print(knitr::kable(tab1,digits = 1,
     caption = 'Descriptives for the onset time differences (ms)'))
```

Table 4 summarise these and we can see that we have a varying number of onsets in different instruments and the typical time difference between the onsets range from X to Y. We will look at the periodicity analysis later, but these basics are good to keep in mind when analysing synchrony.

Let's move on and explore the synchrony between the performers.

## 2.3 Analysis of synchrony

### 2.3.1 Synchrony with the beat sub-divisions

As a broad overview, we can visualise the relative synchrony to equal division subdivision of the beat for each instrument across the time. This gives a summary about when the instrument are playing relative to the beat division and the sections of the song.

```{r synch2isochron,fig.width=9, fig.height=7.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
fig1 <- plot_by_beat(df = CSS_Song2, 
                     instr = c('Bass','Clave','Guitar','Tres'), 
                     beat = 'SD', 
                     virtual='Isochronous.SD.Time',
                     pcols=2)
print(fig1)
```

This six-minute piece does not contain much variation in the patterns of onsets across the beats. We can see that the bass is elaborating the beats during the last minute or so while the clave, guitar and tres keep playing their patterns throughout the piece.

Alternatively, we can show how much the onsets deviate from the isochronous beat or from the mean onset times. Here's an example of the former for two instruments, bass and tres.

```{r synch2isochron2,fig.width=9, fig.height=5.5,fig.cap='Deviations from beat sub-divisions.'}
fig2 <- plot_by_beat(df = CSS_Song2,
                     instr = c('Bass','Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     pcols=1,
                     griddeviations = TRUE)
print(fig2)
```

Perhaps we do not want to follow the trajectories across time but are more interested in finding out the overall synchrony of the instruments across the beat sub-divisions. A boxplot of the synchronies can be obtained using the same function by adding box is TRUE parameter:

```{r synch2isochron3,fig.width=7, fig.height=3.5,fig.cap='Synchrony acrocc sub-division of beat'}
fig3 <- plot_by_beat(df = CSS_Song2,
                     instr = 'Tres',
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     pcols=1,
                     griddeviations = TRUE,
                     box = TRUE)
print(fig3)
```

### 2.3.2 Synchrony between the instruments

To what degree are the pairs of instruments synchronised in this example? Since the instruments usually play widely different amounts of onsets in a piece, and these are bound to be at different beats sub-divisions, the mutual amount of comparable onsets for each pair varies often dramatically. In order to keep the mean and standard deviations comparable, we will randomly sample joint onsets for both instruments.

```{r paired1}
set.seed(1201) # set random seed
N <- 100 # Let's select 100 onsets
d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass',N,1,'SD',TRUE)
print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
```
In this example there are at least 241 shared onset times between the clave and the bass. Let's redo the random sampling 10 times so we get more observations whilst still always main the sample of 100 joint onsets.
```{r paired2}
d10 <- sync_sample_paired(CSS_Song2,'Clave','Bass',N,10,'SD',TRUE)
print(paste('Asynchrony M =',round(mean(d10$asynch*1000),1),
      'ms & SD =',round(sd(d10$asynch*1000),2)))
```

You can carry out this operation easily for all possible pairings of the instruments with `sync_execute_pairs` and visualise the results with a related function (`plot_by_pair`).

```{r fig4,fig.width=4.5, fig.height=4.5, warning=FALSE,fig.cap='Asynchronies across the instrument pairs.'}
inst<-c('Clave','Bass','Guitar','Tres') # Define instruments 
dn <- sync_execute_pairs(CSS_Song2,inst,N,10,'SD')
fig4 <- plot_by_pair(dn)  # plot
print(fig4)  
```

In Figure 4 we see how different pairs of instruments have different mutual synchronies; Bass is consistently ahead of guitar and clave is behind the tres to pick some of the extreme examples from the visualisation.

It is also possible to carry out the comparison with across the beat sub-divisions. Here we choose only three instruments to keep the number of comparisons sensible for a plot (instruments times sub-divisions) and keep the rest of the analysis settings the same.

```{r fig5, fig.width=7.5, fig.height=5.0, warning=FALSE,fig.cap='Asynchronies across the beat sub-divisions and instrument pairs.'}
inst<-c('Bass','Guitar','Tres')
dn <- sync_execute_pairs(CSS_Song2,inst,N,10,'SD')
fig5 <- plot_by_pair(dn,bybeat=TRUE)
print(fig5)  
```

From the breakdown of the sub-divisions across the instruments shown in Figure 5, it is clear that the first sub-division usually has the least amount of asynchrony and less variation in asynchrony, but the actual patterns fluctuate between instrument pairs.

Let's calculate statistics (t-test) about the onset synchronies for these instrument pairs.

```{r stats,results='asis'}
table3 <- data.frame(stats_by_pair(dn))
print(knitr::kable(table3,digits = 1))
```

Table 5 suggests all deviations between these instrument combinations are statistically significant (and this includes a correction for multiple comparisons).

\pagebreak

### 2.3.3 Synchrony with other variables

It is relatively straightforward to explore whether synchrony is linked with another variable. Some meaningful comparisons between synchrony could be carried out with duration of the cycles (as a proxy of musical tempo), dynamics, or note density. Next we try this out with the tempo, here defined as the duration of each cycle in seconds.

```{r sync_vs_tempo,fig.width=3, fig.height=3, fig.cap='Synchrony vs tempo.'}

CSS_Song2 <- CSS_Song2 %>%
  group_by(Cycle) %>%
  mutate(Duration = max(Isochronous.SD.Time) - min(Isochronous.SD.Time))
CSS_Song2 <- ungroup(CSS_Song2) # drop grouping structure

d1 <- sync_sample_paired(df = CSS_Song2,
                         INSTR1 = 'Clave',
                         INSTR2 = 'Bass',
                         N = 200,
                         BNum = 1, 
                         beat = 'Duration')
fig6 <- plot_by_variable(d1,
                       meta = 'Clave-Bass Synchrony',
                       xlab='Cycle duration (in s)')
print(fig6)
```

As figure 6 shows, there does not seem to be any clear associations between the synchrony between bass and clave and the tempo as measured by the cycle durations.

### 2.3.4 Synchrony across separate performances

So far our analysis has consider one examples and synchrony within the example. We can take two different performances, choose the instrument pairings, and carry out the desired comparison, provided that this is conceptually meaningful and technically feasible (similar types of onset and annotation data available). In this example, the column names are slightly different to the first example and the extract is taken from the middle of a performance, so the timing information could be adjusted to remove the long lead time until the first onset.

```{r moredat, eval=TRUE}

invisible(GET("https://osf.io/a4yw7/?action=download", 
    write_disk("CSS_Song3_Onsets_Selected.csv", overwrite = TRUE)))
CSS_Song3_Onset<-read.csv("CSS_Song3_Onsets_Selected.csv",
                          header=T, sep=",")
CSS_Song3_Onset <- dplyr::select(CSS_Song3_Onset,
                                 Label.SD,SD,Clave,Bass,Guitar,Tres)

invisible(GET("https://osf.io/czu6m/?action=download", 
              write_disk("CSS_Song3_Metre.csv", overwrite = TRUE)))
CSS_Song3_Metre <- read.csv("CSS_Song3_Metre.csv", header=T, sep=",")

# Add annotations
CSS_Song3 <- add_annotation(df=CSS_Song3_Onset,
                            annotation = CSS_Song3_Metre$Cycle,
                            time = CSS_Song3_Metre$Time,
                            reference = 'Label.SD')
# Add isochronous beats
CSS_Song3 <- add_isobeats(df=CSS_Song3,instr = 'CycleTime', beat = 'SD')
```

Now that we have another Cuban Salsa performance from the same band in a data frame called `CSS_Song3`, let's calculate the synchrony across the instruments in both performances.

```{r compare,echo=TRUE,eval=TRUE}
d1 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='SD') 
d1 <- data.frame(d1); d1$dataset<-'CSS_Song2'

d2 <- sync_sample_paired(CSS_Song3,'Tres','Bass',N=0,beat='SD')
d2 <- data.frame(d2);d2$dataset<-'CSS_Song3'

D <- suppressMessages(full_join(d1,d2)); # merge the outputs
D$beatL<-factor(D$beatL); D$dataset<-factor(D$dataset) # as factor
D$abs_asynch_ms <- abs(D$asynch*1000) # convert into absolute errors and to ms
```

After these operations, the data frame `D` contains onset synchronies from both performances. We can create a simple boxplot of the combined data ignoring the beat sub-divisions, but then also calculate a simple summary statistics about the differences and show the means across the sub-divisions.

Let's first look at the means.

```{r showmeans, eval=TRUE,results='asis'}
grandmeans <- summarise(group_by(D,dataset),
                        M = mean(abs(asynch*1000)),
                        SD = sd(abs(asynch*1000)))
knitr::kable(grandmeans,
             caption="Grand mean asynchronies.",
             format="simple", digits=2)
# Summary statistics
output <- aov(abs(asynch) ~ beatL * dataset, data=D)
print(knitr::kable(xtable::xtable(output), caption = "Anova table."))
```

There seems to be a statistically significant difference between the pieces in synchrony but no main effect of beat sub-division. However, the interaction between the piece and sub-division seems to be significant, so some of the beats differ between the two pieces.

```{r showplot,eval=TRUE,fig.width=3.0, fig.height=3.0,fig.cap='Absolute asynchronies across the piece.'}
fig10 <- plot_by_dataset(D,'abs_asynch_ms','dataset', box = TRUE)
print(fig10)
```

### 2.3.5 Synchrony across time

Appy the analyses of synchrony across specific segments (cycle, section, etc.). Let's look at synchrony between tres and bass across time (across cycles).

> NOTE 6: REDO THE PLOTS

```{r sync_time1, fig.width=6, fig.height=3.5, fig.cap='Synchrony across cycles.'}
d2 <- sync_sample_paired(df = CSS_Song2,
                         INSTR1 = 'Tres',
                         INSTR2 = 'Bass',
                         N = 0,
                         beat = 'Cycle')
tmp <- data.frame(asynch=d2$asynch*1000,Cycle=d2$beatL)

#fig12 <- plot_by_var_time(df=tmp,var1 = 'Cycle',var2 = 'asynch')

fig12 <- ggplot(tmp,aes(x=Cycle,y=asynch))+
  geom_point(colour='orange') +
  stat_smooth(aes(x=Cycle,y=asynch), method = lm, 
              formula = y ~ splines::bs(x, 12), se = TRUE)+
  xlab('Cycle')+
  ylab('Asynchrony (ms)')+
  theme_linedraw()
print(fig12)
```

We can also explore whether the synchrony (here _absolute asynchrony_) might be related to _tempo_ changes across time. Here define the tempo by the length of the cycles.

```{r sync_time2, fig.width=6, fig.height=3.0,fig.cap='Absolute asynchrony across cycle length.'}

CSS_Song2 <- CSS_Song2 %>%
  group_by(Cycle) %>%
  mutate(Duration = max(Isochronous.SD.Time) - min(Isochronous.SD.Time))

d2 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='Duration')
d3 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='Isochronous.SD.Time')
tmp <- data.frame(asynch=d2$asynch*1000,Duration=d2$beatL,Time=d3$beatL)

fig13 <- ggplot(tmp,aes(x=Time,y=abs(asynch),colour=Duration))+
  geom_line() +
  stat_smooth(aes(x=Time,y=abs(asynch)), method = lm, 
              formula = abs(y) ~ splines::bs(x, 12), se = FALSE)+
  xlab('Time (min:sec)')+
  ylab('Absolute asynchrony (ms)')+
  scale_x_time(breaks = seq(0,350,by=60))+
  scale_color_continuous(name='Cycle Duration',low='red',high = 'blue')+
  theme_linedraw()+
  theme(legend.position="top")

print(fig13)  
```

It seems that there could be some association between increase in tempo and increased asynchrony (at least there is an increase in both around 05:10-5:30).

\pagebreak

## 2.4 Analysis of periodicity

Let's take the bass onsets in the first example (CSS_Song2) and look at the first 20 seconds. Let's add the cycles in the plot to make the pattern more visible.

```{r discrete_rep,echo=TRUE,fig.width=5.5,fig.height=2,fig.cap='Discrete onset times for bass.'}
extract <- dplyr::filter(CSS_Song2,Guitar >= 0 & Guitar < 10)
fig11 <- plot_timeline(data=extract, instr = 'Guitar')
print(fig11)
```

We can see that bass is keeping steady beat, playing a sequence of two notes separated by about 0.40 seconds with 1.3 intervals between the pairs of beats.

For the periodicity analyses, we want to transform the the discrete onsets to continuous time representation with an uniform sampling rate. This makes the comparisons of the models easier and allows to handle time more explicitly. Here we transform the onsets from one performance into onsets represented by time and gaussian distributions centred on the onsets.

```{r continuous_rep,echo=TRUE,fig.width=5.5,fig.height=1.8,fig.cap='Onset curve (continuous representation) for bass.'}
fig12 <- gaussify_onsets(extract$Guitar,sr=250, time=TRUE, plot = TRUE)
```

In this section we use multiple techniques such as auto-correlation, fast fourier transform (FFT), wavelet analysis and onset differences to estimate the periodicity of the onset structures.

### 2.4.1 Estimate periodicity using autocorrelation

```{r acf,echo=TRUE,fig.width=4.0,fig.height=3.0,fig.cap='Autocorrelation of the onset signal between 0 and 0.8 seconds.'}
CSS_Song2 <- ungroup(CSS_Song2) # drop grouping structure
P1 <- periodicity(CSS_Song2,
                  instr = 'Guitar', method='acf',
                  freq_range = c(0,1))
print(P1$Figure)
PM <- periodicity_moments(P1$Curve)
print(PM$Per)
BPM <- period_to_BPM(PM$Per)
print(BPM)
```

Figure 6 displays the auto-correlation function for the onsets and the next calculation (`periodicity_moments`) takes different descriptors of the function (such as the _period_, which is the highest peak in the plot). In this example it is `r PM$Per * 1000` in milliseconds. This is slightly different from the early calculation of the typical onset time differences reported in Table 4, which suggested that the mean onset time difference for guitar was 224 ms and the median was 245 ms. In terms of beats per minute (BPM), this is `r round(BPM)`, which was obtained via `period_to_BPM` function. This tempo in BPM is very likely misleading as the beat division it used is based on 16 sub-division in a cycle, which suggests that guitar plays eight notes (quavers) so the tempo in BPM in is more probably half of the suggested, `r round(BPM/2)` BPM.

### 2.4.2 Create signals varying in periodicity

```{r periodic_examples,fig.width=8,fig.height=5,fig.cap='Four example signals.'}
## Example 1: Random
ex1 <- runif(120,min = 0,max=60)

## Example 2: Accelerating sequence
ex2<-NULL
i<-0.1
for (k in 1:6) {
  tmp <- seq(0,10,by=0.92-i)
  ex2 <-c(ex2,tmp+(k-1)*10)
  i<-i+0.1
}

## Example 3: Periodic
ex3 <- seq(0.5,60,by=0.5)

## Example 4: Period with random noise
ex4<- ex3 + runif(120,min=0.01,max = 0.2)-0.1

## Put them into one data.frame
df <- data.frame(Random = ex1, Accelerating = ex2, 
                 Periodic2Hz = ex3, Periodic2HzNoise = ex4)
fig13 <- plot_timeline(data = df, instr = as.character(names(df)))
print(fig13)
```

### 2.4.4 Period estimation with different techniques

Let's apply four different techniques to the periodic signal that contains a bit of noise (`ex4`).

```{r,fig.width=8,fig.height=5,fig.cap='Periodicity estimation with ACF, FFT, Wavelet and onset time difference.',warning=FALSE,message=FALSE}
P1 <- periodicity(df = data.frame(onset = ex4), 
                  instr = 'onset', method='acf',
                  freq_range = c(0,1),
                  title = '2 Hz signal with ACF')
P2 <- periodicity(df = data.frame(onset = ex4), 
                  instr = 'onset', method='fft',
                  freq_range = c(0,1),
                  title = '2 Hz signal with FFT')
P3 <- periodicity(df = data.frame(onset = ex4), 
                  instr = 'onset', method='wavelet',
                  freq_range = c(0,1),
                  title = '2 Hz signal with Wavelet')
P4 <- periodicity(df = data.frame(onset = ex4), 
                  instr = 'onset', method='diff',
                  freq_range = c(0,1),
                  title = '2 Hz signal with onset time difference')

library(cowplot)
plot_grid(P1$Figure, P2$Figure, P3$Figure, P4$Figure, ncol = 2)
```

Despite the different shape of the function, all methods are able to find the asssumed 2 Hz (0.5 s) peak with `periodicity_moments`, provided that we choose the appropriate method of picking the peak of the curve (for auto-correlation, the default option of `max` would yield half the periodicity (1 s) in this case).

```{r report_periodicity}
P1M <- periodicity_moments(P1$Curve,method='first')
cat(paste('Dominant periodicity according to ACF:',round(P1M$Per*1000),'ms'))
P2M <- periodicity_moments(P2$Curve,method='max')
cat(paste('Dominant periodicity according to FFT:',round(P2M$Per*1000),'ms'))
P3M <- periodicity_moments(P3$Curve,method='max')
cat(paste('Dominant periodicity according to Wavelet:',round(P3M$Per*1000),'ms'))
P4M <- periodicity_moments(P4$Curve,method='max')
cat(paste('Dominant periodicity according to Difference:',round(P4M$Per[1]*1000),'ms'))
```
### 2.4.3 Degree of periodicity with different techniques

We can also ask how different are these example signals in terms of their periodicity.

```{r report_periodicity_max,echo=TRUE,results='asis'}
P_ALL <- matrix(0,4,4)
P_METHOD <-c('acf','fft','wavelet','diff')
for (k in 1:4) {
  for (l in 1:length(P_METHOD)) {
    P <- periodicity(df = df, instr = as.character(names(df)[k]),
                     method = P_METHOD[l],freq_range = c(0,1))
    PM <- periodicity_moments(P$Curve,method='max')
    P_ALL[k,l]<-as.numeric(PM$Max[1])
  }
}

rownames(P_ALL) <- names(df)
colnames(P_ALL) <- P_METHOD
print(knitr::kable(P_ALL,digits = 3,
                   caption = 'Maximum amplitude of the peak 
                   period across four signals and methods.'))
```

Table 8 suggests similar ranking where the periodic signal has the clearest periodicity, periodic with noise is the second (except in difference-based method), and the random signal is the worst. These examples were of course artificial, but they serve to illustrate two aspects of the periodicity estimation, the actual periodicity and its amplitude. Other statistical moments such as skewness and kurtosis can also be explored from this representation. Periodicity in the actual music performances will be much more nuanced and usually contains multiple hierarchies, so picking the right measures to describe these aspects of the beat structures requires further work.

## Conclusions

Short summary to be written.

# Acknowledgements

We acknowledge contributions from IEMP Fellows (RP, etc) and Simone Tarsitani here. And Kelly Jakubowski.

# References

### 2022-06-20 README.Rmd
New version 0.4.0

@@
 `onsetsync` is a R package for musical dynamics involving synchrony.
-There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
+<img src="man/figures/logo.png" align="right" height="139"/> There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
 
@@
 
-Note that `onsetsync` is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g.
-[Librosa](https://librosa.org), or [Mir Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms.
-Here we take it as granted that we have extracted the onsets in some of these programs, probably checked them by hand, and we have the onset times recorded into csv files.
+Note that `onsetsync` is not dedicated to extraction of onsets from audio as that should be done using other tools (e.g.
+[Librosa](https://librosa.org), or [MIR Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using established onset detection algorithms).
+Here we take it as granted that we have extracted the onsets of the music from a recording in some of these tools, probably checked them by hand, and we have the onset times recorded into csv files.
 
@@
 library(ggplot2)
+packageVersion("onsetsync")
 ```
+Read onsets of one Cuban Son performance titled _Palo Santo_ from _IEMP_ dataset about _Cuban Son and Salsa_ at [https://osf.io/sfxa2/](https://osf.io/sfxa2/):
+
+Add annotations of the metric cycles to the data, also obtained from OSF dataset:
 
@@
 
-There are several variants of this summary that create different summaries of the onsets across the beats, but let's move on.
+There are several variants of this visualisation that create alternative summaries of the onsets across the beats, but let's move on.
 
-To what degree are the pairs of instruments synchronised in this example?
+To what degree are the pairs of instruments synchronised to each other in this example?
 Since the instruments usually play widely different amounts of onsets in a piece, and these are bound to be at different beats sub-divisions, the mutual amount of comparable onsets for each pair varies often dramatically.

### 2022-06-20 R/CSS_IEMP.R
0.4.0 Update

@@ -1,0 +1,12 @@
+#' CSS_IEMP
+#'
+#' Cuban Salsa and Son mini corpus, refer to IEMP collection for full details.
+#'
+#' @format A list of 5 containing individual pieces:
+#' \describe{
+#'   \item{names}{Name of the piece}
+#'   \item{id}{ID}
+#'   \item{genre}{Genre of the piece}
+#'   \item{class}{corpus}
+#' }
+"CSS_IEMP"

### 2022-06-20 R/add_isobeats.R

@@ -24,7 +24,7 @@ add_isobeats <- function(df = NULL,
   mean_onset <- all_of <- NULL
   
   #_ beat<-'SD'
-  beat_N <- max(as.matrix(df[,which(colnames(df)==beat)]))  
+  beat_N <- max(as.matrix(df[,which(colnames(df)==beat)]),na.rm = T)  # fix to avoid NAs
   
   # Optional, recode new number for each cycle
   L <- round(nrow(df)/beat_N)

### 2022-06-20 R/periodicity.R

@@ -9,6 +9,7 @@
 #' @param freq_range Frequency range to be included in the periodicity analysis (in seconds)
 #' @param resolution Resolution for some of the analyses (in seconds)
 #' @param ignore_period_under Ignore period lower than this (in seconds)
+#' @param ignore_period_above Ignore period above than this (in seconds)
 #' @param colour Line colour for plotting
 #' @param title Title for plotting
 #' @seealso \code{\link{periodicity_moments}}
@@ -28,6 +29,7 @@ periodicity <-
            freq_range = c(0, 2),
            resolution = 0.01,
            ignore_period_under = 0.2,
+           ignore_period_above = 5,
            colour = 'navyblue',
            title = NULL) {
     # T. Eerola, Durham University, IEMP project
@@ -53,7 +55,7 @@ periodicity <-
     
     onset <- Diff <- Ampl <- Time <- NULL
         
-    tmp <- dplyr::select(df, instr)
+    tmp <- dplyr::select(df, tidyr::all_of(instr)) # added all_of
     colnames(tmp) <- 'onset'
     tmp <- tidyr::drop_na(tmp)    # drop NAs
 
@@ -76,11 +78,11 @@ periodicity <-
        #  forced_max_ampl <- NA
        #  }
        # 
-      fig <- ggplot2::ggplot(Per, ggplot2::aes(x = Time, y = Ampl)) +
+      fig <- ggplot2::ggplot(Per, ggplot2::aes(x = Time, y = Ampl/max(Ampl))) +
         ggplot2::geom_line(colour=colour) +
         ggplot2::xlab('Time (s)') +
+        ggplot2::ylab('Normalised Ampl.') +
         ggplot2::scale_x_continuous(limits = c(freq_range[1], freq_range[2]),breaks=seq(freq_range[1],freq_range[2],by=0.2)) +
-#        ggplot2::scale_y_continuous(limits = c(NA,forced_max_ampl))+
         ggplot2::ggtitle(title)+
         ggplot2::theme_linedraw()+
         ggplot2::theme(plot.title = ggplot2::element_text(hjust = 0.5))
@@ -99,11 +101,11 @@ periodicity <-
                    Time = AC$lag / sampling_rate)
       # filter zero lag or close
[…]

### 2022-06-20 R/plot_by_beat.R

@@ -10,6 +10,7 @@
 #' @param griddeviations Add deviation from the virtual beats in 
 #' @param boxplot Do the graphics by boxplot
 #' @param colour colour for the boxplot
+#' @param colourpalette colors for dots in the timeline, use 'grey' for bw
 #' @return Graphic output
 #' @import ggplot2
 #' @import dplyr
@@ -23,7 +24,8 @@ plot_by_beat <-
            pcols = 1,
            griddeviations = FALSE,
            boxplot = FALSE,
-           colour = 'lightblue') {
+           colour = 'lightblue',
+           colourpalette = 'Set1') {
     # T. Eerola, Durham University, IEMP project
     # 23/1/2018
     # needs work
@@ -85,10 +87,10 @@ plot_by_beat <-
         y = instr,
         colour = name
       )) +
-        ggplot2::geom_point(size = 1, na.rm = TRUE,show.legend = FALSE,alpha=0.85) +
+        ggplot2::geom_point(size = 1, na.rm = TRUE,show.legend = FALSE,alpha=0.85) + # colour='black'
         ggplot2::scale_x_continuous(breaks = seq(1, max(DF$beat))) +
         ggplot2::scale_color_brewer(name = "Instrument",
-                           palette = 'Set1',
+                           palette = colourpalette,
                            type = "div") +
         ggplot2::facet_wrap( ~ name, ncol = pcols) +
         ggplot2::xlab(paste('Beat (', beat, ')', sep = '')) +
@@ -96,6 +98,9 @@ plot_by_beat <-
         ggplot2::ylab('Time (s)') +
         ggplot2::theme_linedraw()
       g1
+      if(colourpalette=='Greys'){
+        g1$layers[[1]]$aes_params$colour <-  "gray20"
+      }
     }
     
     ## PLOT with metrical grid deviations
@@ -112,23 +117,26 @@ plot_by_beat <-
         ggplot2::geom_point(size = 1, na.rm = TRUE,show.legend = FALSE,alpha=0.85) +
         ggplot2::scale_x_continuous(breaks = seq(1, max(DF$beat))) +
         ggplot2::scale_color_brewer(name = "Instr.",
-                           palette = 'Set1',
-                           type = "div") + # was Spectral
+                           palette = colourpalette,
+                           type = "div") + 
         ggplot2::facet_wrap( ~ name, ncol = pcols) +
         ggplot2::xlab(paste('Beat (', beat, ')', sep = '')) +
[…]

### 2022-06-20 R/plot_by_dataset.R

@@ -65,7 +65,7 @@ plot_by_dataset <-
                                     palette = 'Set1',
                                     type = "div") +
         ggplot2::scale_fill_brewer(palette = "Accent") +
-        ggplot2::xlab('Dataset') +
+        ggplot2::xlab('Performance') +
         ggplot2::ylab('Asynchrony (ms)') +
         ggplot2::annotate(
           "text",

### 2022-06-20 R/plot_by_pair.R

@@ -5,6 +5,7 @@
 #' @param df data frame to be processed
 #' @param bybeat Optional (FALSE by default) which adds the beats to the plot
 #' @param reference Optional (0 by default) which adds the reference line to the plot
+#' @param colourpalette Optional colourpalette, Default
 #' @seealso \code{\link{plot_by_variable}}
 #' @return Graphic output
 #' @import ggplot2
@@ -13,12 +14,16 @@
 
 plot_by_pair <-function(df=NULL,
                         bybeat=FALSE,
-                        reference=0){
+                        reference=0,
+                        colourpalette='Pastel1'){
 
   # T. Eerola, Durham University, IEMP project
   # 14/1/2018  
   Instrument <- ms <- beatL <- NULL
   
+  palettetype <- 'qual'
+  if(colourpalette=='Greys' | colourpalette=='Blues'){palettetype <- 'seq'}
+
   if(bybeat==FALSE){ # asynchronies
   
   m <- suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
@@ -26,9 +31,14 @@ plot_by_pair <-function(df=NULL,
 #  colnames(m)<-c('Instrument','ms')
   
   g1<-ggplot2::ggplot(m,ggplot2::aes(x=reorder(x=Instrument,-ms,mean),y=ms,fill=Instrument))+
-    ggplot2::geom_violin(scale = "width",show.legend = FALSE,na.rm=TRUE)+
-    ggplot2::geom_jitter(height = 0, width = 0.15,size=0.15,show.legend = FALSE,na.rm=TRUE)+
-    ggplot2::geom_hline(yintercept = reference,color='red')+
+    ggplot2::geom_violin(scale = "width",show.legend = FALSE,na.rm=TRUE,alpha=0.7)+
+    ggplot2::geom_jitter(height = 0, width = 0.15,size=0.15,show.legend = FALSE,na.rm=TRUE,colour='grey50')+
+    ggplot2::stat_summary(fun = "mean",
+                 geom = "crossbar",
+                 width = 0.4,
+                 color = "black",show.legend = FALSE)+
+    ggplot2::geom_hline(yintercept = reference,color='black',linetype='dashed')+
+    ggplot2::scale_fill_brewer(palette = colourpalette,type = palettetype)+
     ggplot2::coord_flip()+
[…]

### 2022-06-20 R/sync_sample_paired.R

@@ -23,6 +23,9 @@ sync_sample_paired <- function(df = NULL,
   
   # T. Eerola, Durham University, IEMP project
   # 14/1/2018  
+  # Updated 18/6/2022 to incorporate corpus structure
+
+  if(is.data.frame(df)==TRUE){
   
   instr1 <- as.matrix(df[,which(colnames(df)==INSTR1)])  
   instr2 <- as.matrix(df[,which(colnames(df)==INSTR2)])  
@@ -84,5 +87,28 @@ sync_sample_paired <- function(df = NULL,
     }    
   }
   
-  return<-list(asynch=D,beatL=beat_L)
+  } # ends normal single df processing
+  
+  if(is.data.frame(df)==FALSE){
+    print('Calculating across the corpora') # This is not very effective, but  
+    NAMES<-names(df)
+    D<-list()
+    L<-NULL
+    for (i in 1:length(df)) {
+      tmp_df <- df[[i]]
+      D[[i]] <- sync_sample_paired(df = tmp_df,INSTR1,INSTR2,N,BNum,beat)
+      beat_L<-NULL
+      L[i] <- length(D[[i]]$asynch)
+    }
+    R3<-dplyr::bind_rows(D)
+    NAME_INDEX<-NULL
+    for (x in 1:length(L)) {
+      V<-rep(NAMES[x],L[x])
+      NAME_INDEX<-c(NAME_INDEX,V)
+    }
+    R3$name<-NAME_INDEX
+    D<-R3
+  }
+  
+  return<-list(asynch=D,beatL=beat_L)  
 }

### 2022-06-20 R/plot_by_dataset.R
0.4.1 minor fixes

@@ -43,7 +43,7 @@ plot_by_dataset <-
     
     g1 <-
       ggplot2::ggplot(DF, ggplot2::aes(asynch, fill = data), colour = 'black') +
-      ggplot2::geom_density(alpha = 1) +
+      ggplot2::geom_density(alpha = 1, show.legend = FALSE) +
       ggplot2::scale_color_brewer(name = "Dataset",
                                   palette = 'Set1',
                                   type = "div") +
@@ -54,26 +54,30 @@ plot_by_dataset <-
     #g1
     
     if (box == TRUE) {
-      m <-
-        dplyr::summarise(dplyr::group_by(DF, data),
-                         m = median(asynch),
-                         count = n())
+
+      # for N
+        meds <- c(by(DF$asynch, DF$data, median))
+        nlabels <- table(DF$data)
       
-      g1 <- ggplot2::ggplot(DF, ggplot2::aes(data, asynch, fill = data)) +
-        ggplot2::geom_boxplot(outlier.colour = "gray80", outlier.size = 0.25, show.legend = FALSE) +
-        ggplot2::scale_color_brewer(name = "Dataset",
-                                    palette = 'Set1',
-                                    type = "div") +
-        ggplot2::scale_fill_brewer(palette = "Accent") +
+        g1 <- ggplot2::ggplot(DF, ggplot2::aes(data,asynch,fill=data)) +
+        ggplot2::geom_violin(show.legend = FALSE) +
+        ggplot2::stat_summary(fun=median, geom="crossbar", size=.3, color="black",show.legend = FALSE)+
+        ggplot2::geom_text(data = data.frame(), ggplot2::aes(x = names(meds) , y = meds + 2, 
+                                             label = paste0("n=", nlabels)), hjust = 0, size = 3)+
+                  # ggplot2::scale_color_brewer(name = "Dataset",
+        #                             palette = 'Set1',
+        #                             type = "div") +
+        ggplot2::scale_fill_brewer(palette = "Set1") +
         ggplot2::xlab('Performance') +
         ggplot2::ylab('Asynchrony (ms)') +
-        ggplot2::annotate(
-          "text",
-          x = as.numeric(m$data),
-          y = m$m + m$m * 0.1,
-          label = paste("N=", m$count, sep = ""),
-          size = 5
-        ) +
+        ggplot2::coord_flip()+
+        # ggplot2::annotate(
+        #   "text",
[…]

### 2022-06-20 R/plot_by_var_time.R

@@ -9,6 +9,7 @@
 #' @param var2 Variable 2
 #' @param var3 Variable 3 (optional)
 #' @param colour Colour option
+#' @param smoothline_colour Colour option
 #' @param xlabel label for X axis
 #' @param ylabel label for Y axis
 #' @param zlabel label for Z axis
@@ -22,25 +23,26 @@ plot_by_var_time <- function(df,
                              var2 = 'Synchrony',
                              var3 = NULL,
                              colour = 'orange',
+                             smoothline_colour = 'navyblue',
                              xlabel = NULL,
                              ylabel = NULL,
                              zlabel = NULL) {
   # T. Eerola, Durham University, IEMP project
-  # 1/6/2022
+  # 20/6/2022
 
-    var1 <- var2 <- var3 <- Time <- VAR2 <- VAR3 <- NULL
-#  lm <- NULL
+ .data <- NULL # <- var2 <- var3 <- Time <- VAR2 <- VAR3 <- NULL
+
   if(is.null(xlabel)){xlabel <- var1}
   if(is.null(ylabel)){ylabel <- var2}
   if(is.null(zlabel)){zlabel <- var3}
 
   if(is.null(var3)){
-    DF<-df %>% dplyr::select(tidyr::all_of(c(var1, var2)))
-    colnames(DF) <- c('Time', 'VAR2')
+#    DF <- df %>% dplyr::select(tidyr::all_of(c(var1, var2)))
+#    colnames(DF) <- c('Time', 'VAR2')
     g1 <-
-      ggplot2::ggplot(DF, ggplot2::aes(x = Time, y = VAR2)) +
+      ggplot2::ggplot(df, ggplot2::aes(x = .data[[var1]], y = .data[[var2]])) +
       ggplot2::geom_point(colour=colour) +
-      ggplot2::geom_smooth(ggplot2::aes(x = Time, y = VAR2),method = lm,formula = y ~ splines::bs(x, 12), se = FALSE, colour=colour) +
+      ggplot2::geom_smooth(ggplot2::aes(x = .data[[var1]], y = .data[[var2]]),method = lm,formula = y ~ splines::bs(x, 12), se = FALSE, colour=smoothline_colour) +
       ggplot2::xlab(xlabel)+
       ggplot2::ylab(ylabel)+
       #   ggplot2::scale_x_time()+
[…]

### 2022-06-20 README.Rmd
revised README

@@
 
-`onsetsync` is a R package for musical dynamics involving synchrony.
-<img src="man/figures/logo.png" align="right" height="139"/> There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
+`onsetsync` is a R package for musical assessing synchrony between onsets in music.
+<img src="man/figures/logo.png" align="right" height="139"/> There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
 
@@
 
-Note that `onsetsync` is not dedicated to extraction of onsets from audio as that should be done using other tools (e.g.
-[Librosa](https://librosa.org), or [MIR Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using established onset detection algorithms).
-Here we take it as granted that we have extracted the onsets of the music from a recording in some of these tools, probably checked them by hand, and we have the onset times recorded into csv files.
-
 ```{r, message=FALSE, warning=FALSE,echo=TRUE}
 library(onsetsync)
-library(httr)
 library(dplyr)
@@
 ```
-Read onsets of one Cuban Son performance titled _Palo Santo_ from _IEMP_ dataset about _Cuban Son and Salsa_ at [https://osf.io/sfxa2/](https://osf.io/sfxa2/):
+### Reading in data
 
-Add annotations of the metric cycles to the data, also obtained from OSF dataset:
+Read onsets of one Cuban Son performance titled _Palo Santo_ from _IEMP_ dataset at [https://osf.io/sfxa2/](https://osf.io/sfxa2/).
+This song has the onsets and the annotations about the metric cycles already extracted and defined and **comes with the package**.
 
 ```{r message=FALSE,warning=FALSE,eval=TRUE,results='asis'}
-CSS_Song2_Onset <- get_OSF_csv('8a347') # Onsets
-knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple")
-
-CSS_Song2_Metre <- get_OSF_csv('4cdpr') # Annotations
-CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
-                                 Label.SD,SD,Clave,Bass,Guitar,Tres) 
-```
-
-As the onsets and annotations are in different files, let's first combine the raw onset and annotation with `onsetsync`.
-Here we first add annotations (using `add_annotation` function) about the cycles into the onset data.
-We then add isochronous beat times to the data frame using `add_isobeats`, since these are useful reference points for synchrony calculations.
-
-```{r message=FALSE,warning=FALSE,eval=TRUE,results='asis'}
-# Add annotations about the cycle to the data frame
-CSS_Song2 <- add_annotation(df = CSS_Song2_Onset,
-                            annotation = CSS_Song2_Metre$Cycle,
-                            time = CSS_Song2_Metre$Time,
-                            reference = 'Label.SD')
-# Add isochronous beats to the data frame
-CSS_Song2 <- add_isobeats(df = CSS_Song2, 
-                          instr = 'CycleTime', 
-                          beat = 'SD')
-
+CSS_Song2 <- onsetsync::CSS_IEMP[[2]]   # Read one song from internal data
+CSS_Song2 <- dplyr::select(CSS_Song2,Label.SD,SD,Clave,Bass,Guitar,Tres,
+                           CycleTime,Cycle,Isochronous.SD.Time) # Select some columns
 print(knitr::kable(head(CSS_Song2),format = "simple",digits = 2))
@@
 
-Before moving onto the analysis, let's summarise the onset structures in this piece.
+Reading data from is easy either from CSV files in your computer or directly from OSF using `get_OSF_csv` function.
 
-```{r message=FALSE,warning=FALSE,eval=TRUE,results='asis'}
-tab1 <- summarise_onsets(df = CSS_Song2, 
-                         instr = c('Clave','Bass','Guitar','Tres'))
-print(knitr::kable(tab1,digits = 1,
-     caption = 'Descriptives for the onset time differences (ms).'))
-```
+### Visualise onsets structures
 
-As a broad overview, we can visualise the relative synchrony to equal division subdivision of the beat for each instrument across the time.
+As an overview, we can visualise the onsets across the beat sub-divisions for each instrument and do this across the time. Note that time run vertically (from bottom to up) here.
 
-```{r synch2isochron,fig.width=9, fig.height=7.0}
+```{r synch2isochron,fig.width=7, fig.asp=.75}
 fig1 <- plot_by_beat(df = CSS_Song2, 
@@
 
-There are several variants of this visualisation that create alternative summaries of the onsets across the beats, but let's move on.
+### Calculate asynchronies
 
-To what degree are the pairs of instruments synchronised to each other in this example?
-Since the instruments usually play widely different amounts of onsets in a piece, and these are bound to be at different beats sub-divisions, the mutual amount of comparable onsets for each pair varies often dramatically.
-In order to keep the mean and standard deviations comparable, we will randomly sample joint onsets for both instruments.
+To what degree are the pairs of instruments synchronised to each other? Let's visualise the synchrony of all pairings of the instruments in this example.
 
-```{r paired1}
-set.seed(1201) # set random seed
-N <- 200 # Let's select 200 onsets
-d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass',N,1,'SD',TRUE)
-print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
-    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
-```
-Let's visualise the synchrony of all pairings of the instruments in this example.
-
-```{r fig4,warning=FALSE}
-inst<-c('Clave','Bass','Guitar','Tres') # Define instruments 
-dn <- sync_execute_pairs(CSS_Song2,inst,N,10,'SD')
+```{r fig2,warning=FALSE,fig.width=7, fig.asp=.7}
+inst <- c('Clave','Bass','Guitar','Tres') # Define instruments 
+dn <- sync_execute_pairs(CSS_Song2,inst,0,1,'SD')
 fig2 <- plot_by_pair(dn)  # plot
@@
 
-For more examples, see [docs IN PROGRESS](https://tuomaseerola.github.io/onsetsync/) and associated [paper IN PROGRESS](http://).
+As we saw in the first figure, the instruments usually play widely different amounts of onsets in a piece, and these are bound to be at different beats sub-divisions, the mutual amount of comparable onsets for each pair often varies dramatically.
+Comparison of mean asynchronies across sub-divisions can be facilitated by taking random samples of the joint onsets.
+Here we choose a random 200 matching onsets and re-calculate the comparison of asynchrony with this subset 1000 times.
+
+```{r paired1}
+set.seed(1234) # set random seed
+N <- 200 # Let's select 200 onsets
+Bootstrap <- 1000
+d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass',N,Bootstrap,'SD',TRUE)
+print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
+    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
+```
+There are other measures to summarise the asynchronies and visualise them.
+
+### Calculate synchrony across performances
+
+We can apply the measures to a corpus of performances. Here we load five Cuban Son and Salsa performances and run the same analysis as above across the performances.
+
+```{r corpus, warning=FALSE, message=FALSE, eval=TRUE,fig.width=6.0}
+corpus <- onsetsync::CSS_IEMP
+D <- sync_sample_paired(corpus,'Tres','Bass',N=0,beat='SD')
+D <- D$asynch
+D$asynch_abs <- abs(D$asynch)*1000
+fig3 <- plot_by_dataset(D,'asynch_abs','name', box = TRUE)
+print(fig3)
+```
+
+For more examples, see [documentation](https://tuomaseerola.github.io/onsetsync/) and associated [paper IN PROGRESS](http://).
+
+#### Note: How do I get onsets from my music?
+
+Note that `onsetsync` is not dedicated to extraction of onsets from audio as that should be done using other tools (e.g.
+[Librosa](https://librosa.org), or [MIR Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using established onset detection algorithms).
+Here we assume that you have extracted the onsets of the music from a recording already in one of these tools, and preferably checked them by hand.
+It is already more meaningful to carry out analyses of synchrony when you have the metrical information (cycles and beats) identified.
+If your starting point is MIDI, getting the onsets is just a conversion operation away (I would recommend [music21](https://web.mit.edu/music21/doc/moduleReference/moduleConverter.html) for this purpose), although annotation might still be required.

### 2022-06-20 README.Rmd
added link to audio

@@
 This song has the onsets and the annotations about the metric cycles already extracted and defined and **comes with the package**.
+
+<iframe src="https://mfr.osf.io/render?url=https://osf.io/z9uxs/?direct%26mode=render%26action=download%26public_file=True&amp;initialWidth=700&amp;childId=mfrIframe&amp;parentTitle=OSF%20%7C%20CSS_Song2_StereoMix.wav&amp;parentUrl=https%3A%2F%2Fosf.io%2Fz9uxs%2F" width="100%" scrolling="yes" marginheight="0" frameborder="0" allowfullscreen="" webkitallowfullscreen="" sandbox="allow-scripts allow-popups allow-same-origin" height="150px"></iframe>
 

### 2022-06-20 README.Rmd
audio

--- 2022-06-20 README.Rmd
@@
 This song has the onsets and the annotations about the metric cycles already extracted and defined and **comes with the package**.
+
+<div>
+<iframe src="https://mfr.osf.io/render?url=https://osf.io/z9uxs/?direct%26mode=render%26action=download%26public_file=True&amp;initialWidth=700&amp;childId=mfrIframe&amp;parentTitle=OSF%20%7C%20CSS_Song2_StereoMix.wav&amp;parentUrl=https%3A%2F%2Fosf.io%2Fz9uxs%2F" width="100%" scrolling="yes" marginheight="0" frameborder="0" allowfullscreen="" webkitallowfullscreen="" sandbox="allow-scripts allow-popups allow-same-origin" height="150px"></iframe>
+</div>
 

### 2022-07-04 R/gaussify_onsets.R
Removed wavelet and fixed gaussify issue

@@ -2,7 +2,7 @@
 #'
 #' This is a conversion function that turns onsets into sampled time-series.
 #'
-#' @param data Data frame to be processed
+#' @param data Data frame to be processed with discrete time points (onsets)
 #' @param sr Sampling rate (default 50 Hz)
 #' @param wlen Window length of percentage of the sr
 #' @param plot If a plot is needed
@@ -13,11 +13,10 @@
 #' @import tidyr
 
 gaussify_onsets <- function(data = NULL,
-                            sr = 250,
-                            wlen = 0.2,
-                            plot = FALSE,
-                            time = TRUE) {
-  
+                                 sr = 250,
+                                 wlen = 0.3,
+                                 plot = FALSE,
+                                 time = TRUE) {
   onsetcurve <- NULL
   X <- Y <- NULL
   # remove NAs
@@ -26,14 +25,17 @@ gaussify_onsets <- function(data = NULL,
   # define time
   mintime <- min(data)
   maxtime <- max(data)
-  t <- seq(mintime, (maxtime + 1 / sr), by = 1 / sr) # Time
+  t <- seq(mintime, (maxtime + 0 / sr), by = 1 / sr) # Time
   signal <- rep(0, 1, length(t))          # Empty vector
-  #  length(t)
-  #  length(signal)
   
   # add 1 to locations of onsets
   signal[round((data - mintime) * sr) + 1] <- 1
-  #  length(signal)
+  #  sometimes odd and even lengths of the signal result in different length t and signal
+  # fix it by adding one frame to the end of t
+  N <- length(signal) - length(t)
+  if (N == 1) {
+    t <- c(t, t[length(t)] + 1 / sr)
+  }
   
   # define the window length
   WLEN <- round(sr * wlen)
@@ -42,22 +44,18 @@ gaussify_onsets <- function(data = NULL,
   signal <- c(array(0, WLEN), signal, array(0, WLEN))
   #minbuffer <- seq(mintime - WLEN / sr, mintime - 1 / sr, by = 1 / sr)
   #maxbuffer <- seq(maxtime + 2 / sr, maxtime + WLEN+1 / sr, by = 1 / sr)
-
+  
   minbuffer <- array(NA, WLEN)
   maxbuffer <- array(NA, WLEN)
   
   t <- c(minbuffer, t, maxbuffer)
- #   length(t)
[…]

### 2022-07-04 R/periodicity.R

@@ -4,7 +4,7 @@
 #'
 #' @param df data frame to be processed
 #' @param instr Instrument name to be processed
-#' @param method Periodicity analysis method (diff, acf, fft, wavelet)
+#' @param method Periodicity analysis method (diff, acf, fft, per)
 #' @param sampling_rate Sampling rate (Hz)
 #' @param freq_range Frequency range to be included in the periodicity analysis (in seconds)
 #' @param resolution Resolution for some of the analyses (in seconds)
@@ -18,7 +18,6 @@
 #' @import ggplot2
 #' @import tidyr
 #' @import seewave
-#' @import WaveletComp
 #' @seealso \code{\link{periodicity_moments}} for summary function.
 
 periodicity <-
@@ -36,12 +35,12 @@ periodicity <-
     # 22/4/2021
     
     ## FOR EACH INSTRUMENT, CALCULATE DEVIATION AND CREATE MATRIX
-
+    
     # If no title is given, use the data frame name...
-        
-    if(is.null(title)){
-      title=deparse(substitute(df))
-      }
+    
+    if (is.null(title)) {
+      title = deparse(substitute(df))
+    }
     
     ## DEBUG
     # df <- CSS_Song2
@@ -54,16 +53,17 @@ periodicity <-
     # colour <- 'navyblue'
     
     onset <- Diff <- Ampl <- Time <- NULL
-        
+    
     tmp <- dplyr::select(df, tidyr::all_of(instr)) # added all_of
     colnames(tmp) <- 'onset'
     tmp <- tidyr::drop_na(tmp)    # drop NAs
-
+    
     #### ONSET DIFF HISTOGRAM ------------------------------------
     if (method == 'diff') {
       tmp <- dplyr::mutate(tmp, Diff = c(NA, diff(onset)))
       tmp <-
-        dplyr::filter(tmp, Diff >= freq_range[1] & Diff <= freq_range[2])
+        dplyr::filter(tmp, Diff >= freq_range[1] &
+                        Diff <= freq_range[2])
       # estimate density via histogram
       h <-
         hist(
@@ -73,111 +73,148 @@ periodicity <-
         )
       Per <- data.frame(Time = h$mids, Ampl = h$density)
       #H$density<-H$density/max(H$density)  # normalise
-     
-       # if(is.null(forced_max_ampl)){
[…]

### 2022-07-04 R/plot_by_beat.R

@@ -11,6 +11,7 @@
 #' @param boxplot Do the graphics by boxplot
 #' @param colour colour for the boxplot
 #' @param colourpalette colors for dots in the timeline, use 'grey' for bw
+#' @param pointsize size of the dots (defaults 1)
 #' @return Graphic output
 #' @import ggplot2
 #' @import dplyr
@@ -25,7 +26,8 @@ plot_by_beat <-
            griddeviations = FALSE,
            boxplot = FALSE,
            colour = 'lightblue',
-           colourpalette = 'Set1') {
+           colourpalette = 'Set1',
+           pointsize = 1) {
     # T. Eerola, Durham University, IEMP project
     # 23/1/2018
     # needs work
@@ -87,7 +89,7 @@ plot_by_beat <-
         y = instr,
         colour = name
       )) +
-        ggplot2::geom_point(size = 1, na.rm = TRUE,show.legend = FALSE,alpha=0.85) + # colour='black'
+        ggplot2::geom_point(size = pointsize, na.rm = TRUE,show.legend = FALSE,alpha=0.85) + # colour='black'
         ggplot2::scale_x_continuous(breaks = seq(1, max(DF$beat))) +
         ggplot2::scale_color_brewer(name = "Instrument",
                            palette = colourpalette,
@@ -114,7 +116,7 @@ plot_by_beat <-
         y = instr,
         colour = name
       )) +
-        ggplot2::geom_point(size = 1, na.rm = TRUE,show.legend = FALSE,alpha=0.85) +
+        ggplot2::geom_point(size = pointsize, na.rm = TRUE,show.legend = FALSE,alpha=0.85) +
         ggplot2::scale_x_continuous(breaks = seq(1, max(DF$beat))) +
         ggplot2::scale_color_brewer(name = "Instr.",
                            palette = colourpalette,

### 2022-07-04 R/summary_measures.R

@@ -8,7 +8,7 @@
 #' @import tidyr
 #' @import magrittr
 
-summary_measures <-function(df=NULL,relative=FALSE){
+summary_measures <- function(df=NULL,relative=FALSE){
 
   # T. Eerola, Durham University, IEMP project
   # 20/6/2022

### 2022-11-07 README.Rmd
div
Link to OSF audio
Updated README
Big rename of 5 functions

--- 2022-06-20 README.Rmd
@@
 `onsetsync` is a R package for musical assessing synchrony between onsets in music.
-<img src="man/figures/logo.png" align="right" height="139"/> There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
+There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
 
@@
 This song has the onsets and the annotations about the metric cycles already extracted and defined and **comes with the package**.
+
+Go and <A HREF="https://osf.io/z9uxs/" target="_blank">listen to the song at OSF</A>.
 
@@
 ```
+
 There are other measures to summarise the asynchronies and visualise them.
 
-### Calculate synchrony across performances
-
-We can apply the measures to a corpus of performances. Here we load five Cuban Son and Salsa performances and run the same analysis as above across the performances.
-
-```{r corpus, warning=FALSE, message=FALSE, eval=TRUE,fig.width=6.0}
-corpus <- onsetsync::CSS_IEMP
-D <- sync_sample_paired(corpus,'Tres','Bass',N=0,beat='SD')
-D <- D$asynch
-D$asynch_abs <- abs(D$asynch)*1000
-fig3 <- plot_by_dataset(D,'asynch_abs','name', box = TRUE)
-print(fig3)
-```
-
-For more examples, see [documentation](https://tuomaseerola.github.io/onsetsync/) and associated [paper IN PROGRESS](http://).
-
-#### Note: How do I get onsets from my music?
-
-Note that `onsetsync` is not dedicated to extraction of onsets from audio as that should be done using other tools (e.g.
-[Librosa](https://librosa.org), or [MIR Toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using established onset detection algorithms).
-Here we assume that you have extracted the onsets of the music from a recording already in one of these tools, and preferably checked them by hand.
-It is already more meaningful to carry out analyses of synchrony when you have the metrical information (cycles and beats) identified.
-If your starting point is MIDI, getting the onsets is just a conversion operation away (I would recommend [music21](https://web.mit.edu/music21/doc/moduleReference/moduleConverter.html) for this purpose), although annotation might still be required.
+For more examples, see [Get started](docs/articles/onsetsync.html).

### 2022-11-07 R/gaussify_onsets.R

@@ -4,7 +4,7 @@
 #'
 #' @param data Data frame to be processed with discrete time points (onsets)
 #' @param sr Sampling rate (default 500 Hz)
-#' @param wlen Window length of percentage of the sr (default 4%)
+#' @param wlen Window length of percentage of the sr (default 4)
 #' @param plot If a plot is needed
 #' @param time Time samples to be included in the output
 #' @return Output contain vector of onset times

### 2022-11-07 R/nPVI.R

@@ -1,47 +1,0 @@
-#' Measure of durational variability of events
-#'
-#' This function calculates of durational variability of events (nPVI).
-#'
-#' This measure is borne out of language research. It has been noted that 
-#' the variability of vowel duration is greater in stress- vs. syllable-timed 
-#' languages (Grabe & Low, 2002). This measure accounts for the 
-#' variability of durations and is also called "normalized Pairwise Variability 
-#' Index" (nPVI). Patel & Daniele have applied it to music (2003) by comparing
-#' whether the prosody of different languages is also reflected in music. There is 
-#' a clear difference between a sample of works by French and English composers.
-#' 
-#' @param df data frame to be processed
-#' @param instr Instrument to be processed
-#' @return integer value
-#' @export
-#' @import tidyr
-
-nPVI <-
-  function(df = NULL,
-           instr = NULL) {
-    # T. Eerola, Durham University, IEMP project
-    # 25/6/2021
-    
-    
-    tmp <- dplyr::select(df, instr)
-    colnames(tmp) <- 'onset'
-#    tmp<-data.frame(onset=c(0,0.2,0.4,0.6,0.8,1.0)) ## EVEN
-#    tmp<-data.frame(onset=c(0,0.4,0.6,1.0,1.2,1.6)) ## variable
-    tmp <- tidyr::drop_na(tmp)    # drop NAs
-    dur <- diff(tmp$onset)
-    
-# from Miditoolbox ###################
-#    a=dur(nmat,'sec');
-#for i=2:length(a)
-#  s(i)=(a(i-1)-a(i))/((a(i-1)+a(i))/2);
-#end
-#n = 100/(length(a)-1) * sum(abs(s));
-########################################
-  s <- NULL
-  for (i in 2:length(dur)) {
-    s[i-1] <- (dur[i-1]-dur[i])/((dur[i-1]+dur[i])/2);
-  }
-n = (100/(length(dur)-1)) * sum(abs(s))
-    
-    return <- n
-  }

### 2022-11-07 R/period_to_BPM.R

@@ -4,7 +4,7 @@
 #'
 #' @param period data frame to be processed
 #' @param beats_per_minute baseline assumption of how many beats there are in min.
-#' @seealso \code{\link{periodicity_moments}},\code{\link{periodicity}}
+#' @seealso \code{\link{summarise_periodicity}},\code{\link{periodicity}}
 #' @return Numeric output
 #' @export
 #' @examples

### 2022-11-07 R/periodicity.R

@@ -10,13 +10,13 @@
 #' @param resolution Resolution for some of the analyses (in seconds)
 #' @param colour Line colour for plotting 
 #' @param title Title for plotting
-#' @seealso \code{\link{periodicity_moments}}
+#' @seealso \code{\link{summarise_periodicity}}
 #' @return Graphic output
 #' @export
 #' @import ggplot2
 #' @import tidyr
 #' @import seewave
-#' @seealso \code{\link{periodicity_moments}} for summary function.
+#' @seealso \code{\link{summarise_periodicity}} for summary function.
 
 periodicity <-
   function(df = NULL,

### 2022-11-07 R/periodicity_moments.R

@@ -1,63 +1,0 @@
-#' Summarise onset periodicities by calculating the moments (mean, etc.)
-#'
-#' This function plots the calculated asynchronies of instruments and labels these by instruments.
-#'
-#' @param data data frame to be processed
-#' @param method method of choosing the peaks (max default, or first)
-#' @return Graphic output
-#' @seealso \code{\link{periodicity}}
-#' @import moments
-#' @export
-
-periodicity_moments <- function(data = NULL,
-                                method = 'max') {
-  # T. Eerola, Durham University, IEMP project
-  # 22/4/2021
-  P <- Ampl <- Time <- NULL
-  # Extract moments
-
-  find_peaks <- function (x, m = 3) {
-    shape <- diff(sign(diff(x, na.pad = FALSE)))
-    pks <- sapply(
-      which(shape < 0),
-      FUN = function(i) {
-        z <- i - m + 1
-        z <- ifelse(z > 0, z, 1)
-        w <- i + m + 1
-        w <- ifelse(w < length(x), w, length(x))
-        if (all(x[c(z:i, (i + 2):w)] <= x[i + 1]))
-          return(i + 1)
-        else
-          return(numeric(0))
-      }
-    )
-    pks <- unlist(pks)
-    pks
-  }
-  
-  if (method == 'first') {
-    Per_peaks <- data$Time[find_peaks(data$Ampl)]
-    PerAmpl_peaks <- data$Ampl[find_peaks(data$Ampl)]
-    Summary <-
-      dplyr::summarise(
-        data,
-        Max = PerAmpl_peaks[1],
-        Sum = sum(Ampl),
-        Kurt = moments::kurtosis(Ampl),
-        Skew = moments::skewness(Ampl),
-        Per = Per_peaks[1]
-      )
-  }
-  if (method == 'max') {
-    Summary <-
-      dplyr::summarise(
-        data,
-        Max = max(Ampl),
-        Sum = sum(Ampl),
-        Kurt = moments::kurtosis(Ampl),
-        Skew = moments::skewness(Ampl),
-        Per = Time[Ampl == max(Ampl)]
-      )
-  }
-  return <- Summary
-}

### 2022-11-07 R/periodicity_nPVI.R

@@ -1,0 +1,47 @@
+#' Measure of durational variability of events
+#'
+#' This function calculates of durational variability of events (nPVI).
+#'
+#' This measure is borne out of language research. It has been noted that 
+#' the variability of vowel duration is greater in stress- vs. syllable-timed 
+#' languages (Grabe & Low, 2002). This measure accounts for the 
+#' variability of durations and is also called "normalized Pairwise Variability 
+#' Index" (nPVI). Patel & Daniele have applied it to music (2003) by comparing
+#' whether the prosody of different languages is also reflected in music. There is 
+#' a clear difference between a sample of works by French and English composers.
+#' 
+#' @param df data frame to be processed
+#' @param instr Instrument to be processed
+#' @return integer value
+#' @export
+#' @import tidyr
+
+periodicity_nPVI <-
+  function(df = NULL,
+           instr = NULL) {
+    # T. Eerola, Durham University, IEMP project
+    # 25/6/2021
+    
+    
+    tmp <- dplyr::select(df, instr)
+    colnames(tmp) <- 'onset'
+#    tmp<-data.frame(onset=c(0,0.2,0.4,0.6,0.8,1.0)) ## EVEN
+#    tmp<-data.frame(onset=c(0,0.4,0.6,1.0,1.2,1.6)) ## variable
+    tmp <- tidyr::drop_na(tmp)    # drop NAs
+    dur <- diff(tmp$onset)
+    
+# from Miditoolbox ###################
+#    a=dur(nmat,'sec');
+#for i=2:length(a)
+#  s(i)=(a(i-1)-a(i))/((a(i-1)+a(i))/2);
+#end
+#n = 100/(length(a)-1) * sum(abs(s));
+########################################
+  s <- NULL
+  for (i in 2:length(dur)) {
+    s[i-1] <- (dur[i-1]-dur[i])/((dur[i-1]+dur[i])/2);
+  }
+n = (100/(length(dur)-1)) * sum(abs(s))
+    
+    return <- n
+  }

### 2022-11-07 R/stats_by_pair.R

@@ -1,104 +1,0 @@
-#' Calculate summary statistics for asynchronies by instrument pairs
-#'
-#' This function calculates simple t-test of the asynchronies of instruments (whether it differs from zero)
-#'
-#' @param df data frame to be processed
-#' @param bybeat logical value whether beat information is used
-#' @param adjust Adjust p values with Bonferroni correction (default)
-#' @seealso \code{\link{sync_sample_paired}}, \code{\link{sync_execute_pairs}}
-#' @return Statistics output (Mean, Standard deviation, T-stat, P value)
-#' @import ggplot2
-#' @import tidyr
-#' @import magrittr
-#' @import scales
-
-stats_by_pair <-function(df,
-                         bybeat=FALSE,
-                         adjust='bonferroni'){
-
-# T. Eerola, Durham University, IEMP project
-# 14/1/2018  
-  Instrument <- summarise <- across <- NULL
-
-  if(bybeat==FALSE){ # asynchronies
-    m<-suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
-    m$ms<-m$ms*1000
-    #colnames(m)<-c('Instrument','ms')
-   # head(m)
-    
-    # check whether instrument is different from 0
-    L<-levels(m$Instrument)
-#    L
-    T<-NULL
-    for(k in 1:length(L)){
-      tmp<-m$ms[m$Instrument==L[k]]
-      if(sum(is.na(tmp))<length(tmp)){
-        t<-t.test(tmp,mu=0)
-        T$tval[k]<-as.numeric(t$statistic)
-        T$pval[k]<-as.numeric(t$p.value)
-      }
-      if(sum(is.na(tmp))==length(tmp)){
-        T$tval[k]<-NA
-        T$pval[k]<-NA
-      }
-    }
-  }
-  
-  if(bybeat==TRUE){
- #   print('by beat')
-    m <- suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
-    m$ms<-m$ms*1000
-#    colnames(m)<-c('Instrument','ms')
-#    head(m)
-    b<-suppressMessages(reshape2::melt(df$beatL))
-    m$beatL<-b$value
- #   str(m)
-    m$beatL<-factor(m$beatL)
-    
-    
[…]

### 2022-11-07 R/summarise_periodicity.R

@@ -1,0 +1,63 @@
+#' Summarise onset periodicities by calculating the moments (mean, etc.)
+#'
+#' This function plots the calculated asynchronies of instruments and labels these by instruments.
+#'
+#' @param data data frame to be processed
+#' @param method method of choosing the peaks (max default, or first)
+#' @return Graphic output
+#' @seealso \code{\link{periodicity}}
+#' @import moments
+#' @export
+
+summarise_periodicity <- function(data = NULL,
+                                method = 'max') {
+  # T. Eerola, Durham University, IEMP project
+  # 22/4/2021
+  P <- Ampl <- Time <- NULL
+  # Extract moments
+
+  find_peaks <- function (x, m = 3) {
+    shape <- diff(sign(diff(x, na.pad = FALSE)))
+    pks <- sapply(
+      which(shape < 0),
+      FUN = function(i) {
+        z <- i - m + 1
+        z <- ifelse(z > 0, z, 1)
+        w <- i + m + 1
+        w <- ifelse(w < length(x), w, length(x))
+        if (all(x[c(z:i, (i + 2):w)] <= x[i + 1]))
+          return(i + 1)
+        else
+          return(numeric(0))
+      }
+    )
+    pks <- unlist(pks)
+    pks
+  }
+  
+  if (method == 'first') {
+    Per_peaks <- data$Time[find_peaks(data$Ampl)]
+    PerAmpl_peaks <- data$Ampl[find_peaks(data$Ampl)]
+    Summary <-
+      dplyr::summarise(
+        data,
+        Max = PerAmpl_peaks[1],
+        Sum = sum(Ampl),
+        Kurt = moments::kurtosis(Ampl),
+        Skew = moments::skewness(Ampl),
+        Per = Per_peaks[1]
+      )
+  }
+  if (method == 'max') {
+    Summary <-
+      dplyr::summarise(
+        data,
+        Max = max(Ampl),
+        Sum = sum(Ampl),
+        Kurt = moments::kurtosis(Ampl),
+        Skew = moments::skewness(Ampl),
+        Per = Time[Ampl == max(Ampl)]
+      )
+  }
+  return <- Summary
+}

### 2022-11-07 R/summarise_sync.R

@@ -1,0 +1,43 @@
+#' Summarise synchronies
+#'
+#' This function summarises the calculated asynchronies.
+#'
+#' @param df data frame to be processed
+#' @param relative Whether calculation is relative measure
+#' @return Integer output
+#' @import tidyr
+#' @import magrittr
+
+summarise_sync <- function(df=NULL,relative=FALSE){
+
+  # T. Eerola, Durham University, IEMP project
+  # 20/6/2022
+  #Instrument <- ms <- beatL <- NULL
+  
+  d_in_ms <- df$asynch * 1000 # convert to ms
+  asynch <- NULL  
+  if(relative==FALSE){
+    asynch$pairwise_asynchronization <- sd(d_in_ms)
+    asynch$mean_absolute_asynchrony <- mean(abs(d_in_ms))
+    asynch$groupwise_asynchronization <- sqrt(mean(d_in_ms))
+    asynch$mean_pairwise_asynchrony <- mean(d_in_ms)
+  }
+  if(relative==TRUE){
+    asynch$mean_pairwise_asynchrony <- mean(d_in_ms)
+  }
+  # # Mean relative asynchrony requires computing the mean
+  # CSS_Song2 <- dplyr::select(CSS_Song2,-Isochronous.SD.Time)
+  # CSS_Song2 <- add_isobeats(df = CSS_Song2, 
+  #                           instr = c('Guitar','Bass','Tres','Clave'), 
+  #                           beat = 'SD')
+  # # then comparing clave and bass to this
+  # d_clave <- sync_sample_paired(CSS_Song2,'Clave','Isochronous.SD.Time',
+  #                               N=0,beat = 'SD')
+  # d_bass <- sync_sample_paired(CSS_Song2,'Bass','Isochronous.SD.Time',
+  #                              N=0,beat = 'SD')
+  # asynch$MS[5] <- mean(c(d_clave$asynch*1000,d_bass$asynch*1000))
+  
+  
+return <- asynch
+  
+}

### 2022-11-07 R/summarise_sync_by_pair.R

@@ -1,0 +1,104 @@
+#' Calculate summary statistics for asynchronies by instrument pairs
+#'
+#' This function calculates simple t-test of the asynchronies of instruments (whether it differs from zero)
+#'
+#' @param df data frame to be processed
+#' @param bybeat logical value whether beat information is used
+#' @param adjust Adjust p values with Bonferroni correction (default)
+#' @seealso \code{\link{sync_sample_paired}}, \code{\link{sync_execute_pairs}}
+#' @return Statistics output (Mean, Standard deviation, T-stat, P value)
+#' @import ggplot2
+#' @import tidyr
+#' @import magrittr
+#' @import scales
+
+summarise_sync_by_pair <-function(df,
+                         bybeat=FALSE,
+                         adjust='bonferroni'){
+
+# T. Eerola, Durham University, IEMP project
+# 14/1/2018  
+  Instrument <- summarise <- across <- NULL
+
+  if(bybeat==FALSE){ # asynchronies
+    m<-suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
+    m$ms<-m$ms*1000
+    #colnames(m)<-c('Instrument','ms')
+   # head(m)
+    
+    # check whether instrument is different from 0
+    L<-levels(m$Instrument)
+#    L
+    T<-NULL
+    for(k in 1:length(L)){
+      tmp<-m$ms[m$Instrument==L[k]]
+      if(sum(is.na(tmp))<length(tmp)){
+        t<-t.test(tmp,mu=0)
+        T$tval[k]<-as.numeric(t$statistic)
+        T$pval[k]<-as.numeric(t$p.value)
+      }
+      if(sum(is.na(tmp))==length(tmp)){
+        T$tval[k]<-NA
+        T$pval[k]<-NA
+      }
+    }
+  }
+  
+  if(bybeat==TRUE){
+ #   print('by beat')
+    m <- suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
+    m$ms<-m$ms*1000
+#    colnames(m)<-c('Instrument','ms')
+#    head(m)
+    b<-suppressMessages(reshape2::melt(df$beatL))
+    m$beatL<-b$value
+ #   str(m)
+    m$beatL<-factor(m$beatL)
+    
+    
[…]

### 2022-11-07 R/summary_measures.R

@@ -1,43 +1,0 @@
-#' Plot synchronies by instrument pairs
-#'
-#' This function plots the calculated asynchronies of instrument pairs.
-#'
-#' @param df data frame to be processed
-#' @param relative Whether calculation is relative measure
-#' @return Integer output
-#' @import tidyr
-#' @import magrittr
-
-summary_measures <- function(df=NULL,relative=FALSE){
-
-  # T. Eerola, Durham University, IEMP project
-  # 20/6/2022
-  #Instrument <- ms <- beatL <- NULL
-  
-  d_in_ms <- df$asynch * 1000 # convert to ms
-  asynch <- NULL  
-  if(relative==FALSE){
-    asynch$pairwise_asynchronization <- sd(d_in_ms)
-    asynch$mean_absolute_asynchrony <- mean(abs(d_in_ms))
-    asynch$groupwise_asynchronization <- sqrt(mean(d_in_ms))
-    asynch$mean_pairwise_asynchrony <- mean(d_in_ms)
-  }
-  if(relative==TRUE){
-    asynch$mean_pairwise_asynchrony <- mean(d_in_ms)
-  }
-  # # Mean relative asynchrony requires computing the mean
-  # CSS_Song2 <- dplyr::select(CSS_Song2,-Isochronous.SD.Time)
-  # CSS_Song2 <- add_isobeats(df = CSS_Song2, 
-  #                           instr = c('Guitar','Bass','Tres','Clave'), 
-  #                           beat = 'SD')
-  # # then comparing clave and bass to this
-  # d_clave <- sync_sample_paired(CSS_Song2,'Clave','Isochronous.SD.Time',
-  #                               N=0,beat = 'SD')
-  # d_bass <- sync_sample_paired(CSS_Song2,'Bass','Isochronous.SD.Time',
-  #                              N=0,beat = 'SD')
-  # asynch$MS[5] <- mean(c(d_clave$asynch*1000,d_bass$asynch*1000))
-  
-  
-return <- asynch
-  
-}

### 2022-11-07 docs/index.R

@@ -1,18 +1,0 @@
-## ---- include = FALSE---------------------------------------------------------
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
-
-## ----setup--------------------------------------------------------------------
-if (!require(devtools)) install.packages("devtools")
-devtools::install_github("tuomaseerola/onsetsync")
-
-library(onsetsync)
-library(httr)
-library(dplyr)
-library(ggplot2)
-
-CSS_Song2_Onset <- get_OSF_csv('8a347') # Onsets
-knitr::kable(head(CSS_Song2_Onset[,1:8,]),format = "simple")
-

### 2022-11-10 onsetsync.Rmd
Added onsetsync.Rmd

---
title: "onsetsync demo"
author: "Tuomas Eerola"
date: "2022-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Onsetsync demo as Rmd notebook**

A Rmarkdown notebook with code examples in R for [onsetsync](https://github.com/tuomaseerola/onsetsync) by [Tuomas Eerola](https://www.durham.ac.uk/staff/tuomas-eerola/), [Music and Science Lab]() at [Durham University](https://www.durham.ac.uk) and [Music and Science Lab](https://musicscience.net).

To run the code in your RStudio, download the Rmd file and open it in your Studio. Here's the link: [https://github.com/tuomaseerola/onsetsync/onsetsync.Rmd](https://github.com/tuomaseerola/onsetsync/onsetsync.Rmd). It assumes that you have `devtools` package installed to be able to install packages from github.

File `onsetsync.Rmd` | Version `9/11/2022` | [Back to Index](https://github.com/tuomaseerola/onsetsync/)

# Load libraries

```{r}
library(ggplot2,quietly = TRUE)
library(tidyverse,quietly = TRUE)
require(devtools)
devtools::install_github("tuomaseerola/onsetsync")
library(onsetsync,quietly = TRUE)
ver <- packageVersion('onsetsync')
print(ver)
```

## Load data (internal)

Here we take a Cuban Son performance titled *Palo Santo* from *IEMP* dataset [@Clayton2020emr] at [https://osf.io/sfxa2/](https://osf.io/sfxa2/):

```{r}
CSS_Song2 <- onsetsync::CSS_IEMP[[2]]   # Read one song from internal data
CSS_Song2 <- dplyr::select(CSS_Song2,Label.SD,SD,Clave,Bass,Guitar,Tres,
                           Cycle,Isochronous.SD.Time) # Select some columns
print(head(CSS_Song2))
```

## Load Data -- From OSF (optional)

```{r}
require(httr)                           # Additional library needed
CSS_Song2_Onset <- get_OSF_csv('8a347') # Get onsets
print(head(CSS_Song2_Onset[,1:8,]))
```

## Summarise Raw Data

```{r}
tab1 <- summarise_onsets(df = CSS_Song2, 
                         instr = c('Clave','Bass','Guitar','Tres'))
print(knitr::kable(tab1,format = "markdown", digits = 1,
     caption = 'Onsets counts (N) and mean IOIs in ms.'))
```

## Visualisation

## Visualise Onset Structures

```{r}
fig1 <- plot_by_beat(df = CSS_Song2, 
                     instr = c('Bass','Clave','Guitar','Tres'), 
                     beat = 'SD', 
                     virtual = 'Isochronous.SD.Time',
                     pcols = 2)
print(fig1)
```

## Visualise Deviations from Isochrony

```{r}
fig2 <- plot_by_beat(df = CSS_Song2,
                     instr = c('Bass','Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     pcols=1,
                     griddeviations = TRUE)
print(fig2)
```

## Visualise Across Sub-beats

```{r}
fig2 <- plot_by_beat(df = CSS_Song2,
                     instr = c('Tres'),
                     beat = 'SD',
                     virtual = 'Isochronous.SD.Time', 
                     pcols=1,
                     griddeviations = TRUE, box = TRUE)
print(fig2)
```

# Synchrony

## Classic Measures of Synchrony

See definitions in classic literature (Rasch, Wing, Clayton).

```{r}
d <- sync_sample_paired(CSS_Song2,'Clave','Bass',N=0,beat = 'SD')
s <- summarise_sync(d)
S <- data.frame(s); St<-t(S); colnames(St) <- 'ms'
knitr::kable(St,digits = 1,caption = 'Classic asynchrony measures.')
```

## Synchrony Between Instruments

Number of _shared onsets_ may vary across instruments/beats. Comparisons with `sampling` and `bootstrapping`.

```{r}
set.seed(1201) # set random seed
N <- 100       # Let's select random 100 onsets from both
Boot <- 1      # No multiple bootstraps
d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass', N, Boot,'SD',TRUE)
print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
```

## Synchrony Between Instruments (plot)

```{r}
inst <- c('Clave','Bass','Guitar','Tres') # Define instruments 
dn <- sync_execute_pairs(CSS_Song2,inst,N,10,'SD')
print(plot_by_pair(dn))  # plot
```

## Descriptive Statistics

```{r}
table3 <- data.frame(summarise_sync_by_pair(dn))
print(knitr::kable(table3,
                   digits = 1,
                   caption = 'Descriptives, t-test and p-value.'))
```

## Synchrony with other Variables

Is there a link between beat durations and synchrony?

```{r}
CSS_Song2 <- CSS_Song2 %>% 
  group_by(Cycle) %>%
  mutate(Duration = max(Isochronous.SD.Time) - min(Isochronous.SD.Time))
CSS_Song2 <- ungroup(CSS_Song2) # drop grouping structure
d1 <- sync_sample_paired(df = CSS_Song2,
                         INSTR1 = 'Clave', INSTR2 = 'Bass',
                         N = 200, BNum = 1, beat = 'Duration')
fig6 <- plot_by_variable(d1,meta = 'Clave-Bass Synchrony',xlab = 'Cycle duration (in s)')
print(fig6)
```

## Synchrony with other Variables

Are there specific trends in synchrony across the performance?

```{r}
d2 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='Duration')
d3 <- sync_sample_paired(CSS_Song2,'Tres','Bass',N=0,beat='Isochronous.SD.Time')
tmp <- data.frame(asynch=d2$asynch*1000,Duration=d2$beatL,Time=d3$beatL)
fig9 <- plot_by_var_time(df = tmp,
                          var1 = 'Time',
                          var2 = 'asynch',
                          var3 = 'Duration',
                          ylabel = 'Asynchrony (ms)')
fig9
```

## Synchrony across Several Performances (corpus)

```{r}
corpus <- onsetsync::CSS_IEMP                              # This is a small corpus
D <- sync_sample_paired(corpus,'Tres','Bass',N=0,beat='SD')
D <- D$asynch
D$asynch_abs <- abs(D$asynch)*1000                         # Convert to ms
fig7 <- plot_by_dataset(D,'asynch_abs','name', box = TRUE)
print(fig7)
```

# Periodicity

## Periodicity of Guitar Onsets

```{r}
extract <- dplyr::filter(CSS_Song2, Guitar >= 60 & Guitar < 65)
fig10 <- gaussify_onsets(extract$Guitar, wlen = 0.2, plot = TRUE)
print(fig10$fig)
```

## Estimate Periodicity

Four methods to estimate periodicity (autocorrelation, FFT, smoothed FFT, or difference)

```{r, warning=FALSE}
per <- periodicity(extract,instr='Guitar',freq_range=c(0.1,0.6))
print(per$Figure)
```

## Summarise Periodicity

Convenience functions to extract peak of periodicity curve.

```{r}
psumm <- summarise_periodicity(per$Curve)
print(paste('Period is:', round(psumm$Per,3), 'seconds'))
```

Or convert to BPM

```{r}
print(paste('In BPM:',period_to_BPM(psumm$Per*2))) # knowing these are sub-beats
```

## Estimate Periodicity

Use **autocorrelation** this time...

```{r}
per2 <- periodicity(extract,instr='Guitar',method='acf',freq_range=c(0.1,0.6))
print(per2$Figure)
```

## Other Sources of Onset Data {.smaller}

### Real performances

  - [CompMusic datasets at UPF: Hindustani (North India), Carnatic (South India), Turkish-makam (Turkey), Arab-Andalusian (Maghreb), and Beijing Opera (China)](https://compmusic.upf.edu/datasets)
  - [Tap & Fiddle: Scandinavian Fiddle Tunes with Accompanying Foot-Tapping](https://zenodo.org/record/4308731)
  - [The MAESTRO (MIDI and Audio Edited for Synchronous Tracks and Organisation)](https://magenta.tensorflow.org/datasets/maestro)

    - 200 hours of real recordings with MIDI and audio (Piano competition)

  - [University of Rochester Multi-Modal Musical Performance](http://www2.ece.rochester.edu/projects/air/projects/URMP.html) 
  - [Carnatic Music Rhythm Dataset](http://compmusic.upf.edu/carnatic-varnam-dataset)

## Credits

::: callout-note
Thanks for **Adrian Poole** and **Simone Tarsitani** for preparing [IEMP Cuban](10.17605/osf.io/s437t) data included here.
:::

# References

Clayton, M., K. Jakubowski, T. Eerola, P. Keller, A. Camurri, G. Volpe, and P. Alborno. (2020). Interpersonal Entrainment in Music Performance: Theory, Method and Model. _Music Perception 38 (2)_, 136–94.

Clayton, M., S. Tarsitani, R. Jankowsky, L. Jure, L. Leante, R. Polak, A. Poole, et al. (2021). The Interpersonal Entrainment in Music Performance Data Collection. _Empirical Musicology Review 16 (1),_ 65–84. [http://dx.doi.org/10.18061/emr.v16i1.7555](http://dx.doi.org/10.18061/emr.v16i1.7555)

Rasch, Rudolf A. (1979). Synchronization in Performed Ensemble Music. _Acta Acustica United with Acustica 43 (2)_, 121–31.

Repp, Bruno H, and Yi-Huang Su. (2013). Sensorimotor Synchronization: A Review of Recent Research (2006–2012). _Psychonomic Bulletin & Review 20 (3)_, 403–52.

Wilkinson, Mark D, Michel Dumontier, IJsbrand Jan Aalbersberg, Gabrielle Appleton, Myles Axton, Arie Baak, Niklas Blomberg, et al. (2016). The FAIR Guiding Principles for Scientific Data Management and Stewardship. _Scientific Data 3 (1)_, 1–9.

Wing, Alan M, Satoshi Endo, Adrian Bradbury, & Dirk Vorberg. (2014). Optimal Feedback Correction in String Quartet Synchronization. _Journal of The Royal Society Interface 11 (93)_, 20131125.

### 2022-11-10 README.Rmd
Added links to Rmd and ipynb

@@
 
-For more examples, see [Get started](docs/articles/onsetsync.html).
+## More information
+
+For more examples, see [Get started](https://tuomaseerola.github.io/onsetsync/articles/onsetsync.html).
+
+- [Tutorial in Rmarkdown](https://github.com/tuomaseerola/onsetsync/blob/master/onsetsync.Rmd) which requires you to run the code in RStudio.
+
+- [Tutorial as Jupyter notebook](https://github.com/tuomaseerola/onsetsync/blob/master/onsetsync.ipynb) which allows you to run the code in Colab (in your browser).

### 2022-11-10 README.Rmd
Updated readme to include links to notebooks

--- 2022-11-07 README.Rmd
@@
 There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
+
+For trying the package with notebooks, jump to <A HREF="#more">more information</A>.
 
@@
 
-For more examples, see [Get started](docs/articles/onsetsync.html).
+## More information
+<a id="more">
+
+For more examples, see [Get started](https://tuomaseerola.github.io/onsetsync/articles/onsetsync.html) or see the [function reference](https://tuomaseerola.github.io/onsetsync/reference/index.html).
+
+- [Tutorial as a Rmarkdown notebook](https://github.com/tuomaseerola/onsetsync/blob/master/onsetsync.Rmd) which requires you to run the code in RStudio.
+
+- [Tutorial as a Jupyter notebook](https://github.com/tuomaseerola/onsetsync/blob/master/onsetsync.ipynb) which allows you to run the code in Colab (in your browser).

### 2022-12-23 R/add_annotation.R
0.4.7 update

@@ -18,15 +18,7 @@ add_annotation <-
     # T. Eerola, Durham University, IEMP project
     # 18/5/2021
     # New function to integrate annotations of cycles with onsets
-    
-    #### DEBUG ----------------------------------------
-    # print('debugging')
-    # df <- asereOnset
-    # annotation <- 'Cycle'
-    # reference <- 'Label.SD'
 
-    # match first Subdivision
-    
     if(reference=='Label.SD'){
       tmp <- df$Label.SD
       ind <- grep(pattern = ":1$|\\|1$", x=tmp) # gets the cycle beginnings (fixed to include | in June 2022)

### 2022-12-23 R/add_isobeats.R

@@ -42,15 +42,7 @@ add_isobeats <- function(df = NULL,
   df$mean_onset<-df2$mean_onset
   df$mean_onset[df$mean_onset=='NaN']<-NA
   
-  # df3<-df2 %>%
-  #   mutate(
-  #     iso = case_when(
-  #       iso>1 ~ seq(iso,lag(iso,16)),
-  #       TRUE                      ~  0
-  #     )
-  #   )
-  
-  DF<-NULL
+  DF <- NULL
   for(k in 1:(max(df$cycle,na.rm = TRUE)-1)){
   #    print(k)
     tmp<-dplyr::filter(df,cycle==k)
@@ -66,6 +58,6 @@ add_isobeats <- function(df = NULL,
   }
   # relabel output iso
   names(DF)[which(names(DF)=='iso')]<-beatlabel
-  DF<-dplyr::select(DF,-cycle,-mean_onset)    # drop unless needed explicitly
-  return<-DF
+  DF <- dplyr::select(DF,-cycle,-mean_onset)    # drop unless needed explicitly
+  return <- DF
 }

### 2022-12-23 R/periodicity.R

@@ -158,7 +158,7 @@ periodicity <-
       
       fig <-
         ggplot2::ggplot(Per, ggplot2::aes(x = Time, y = Ampl / max(Ampl))) +
-        ggplot2::geom_line(colour = colour) +
+        ggplot2::geom_line(colour = colour, na.rm = TRUE) +
         ggplot2::xlab('Time (s)') +
         ggplot2::ylab('Normalised Ampl.') +
         ggplot2::scale_x_continuous(
@@ -205,7 +205,7 @@ periodicity <-
       
       fig <-
         ggplot2::ggplot(Per, ggplot2::aes(x = Time, y = Ampl / max(Ampl))) +
-        ggplot2::geom_line(colour = colour) +
+        ggplot2::geom_line(colour = colour,na.rm = TRUE) +
         ggplot2::xlab('Time (s)') +
         ggplot2::ylab('Normalised Ampl.') +
         ggplot2::scale_x_continuous(

### 2022-12-23 R/plot_by_pair.R

@@ -5,17 +5,17 @@
 #' @param df data frame to be processed
 #' @param bybeat Optional (FALSE by default) which adds the beats to the plot
 #' @param reference Optional (0 by default) which adds the reference line to the plot
-#' @param colourpalette Optional colourpalette, Default 'PuOr' (Purple to Orange)
+#' @param colourpalette Optional colour palette, Default 'PuOr' (Purple to Orange)
 #' @seealso \code{\link{plot_by_variable}}
 #' @return Graphic output
 #' @import ggplot2
 #' @import tidyr
 #' @import magrittr
 
-plot_by_pair <-function(df=NULL,
-                        bybeat=FALSE,
-                        reference=0,
-                        colourpalette='PuOr'){
+plot_by_pair <-function(df = NULL,
+                        bybeat = FALSE,
+                        reference = 0,
+                        colourpalette = 'PuOr'){
 
   # T. Eerola, Durham University, IEMP project
   # 14/1/2018  
@@ -26,12 +26,10 @@ plot_by_pair <-function(df=NULL,
 
   if(bybeat==FALSE){ # asynchronies
   
-#  m <- suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
   m <- tidyr::pivot_longer(data=df$asynch,cols=c(1:dim(df$asynch)[2]),names_to = 'Instrument', values_to = 'ms')
   m$ms <- m$ms*1000
-#  colnames(m)<-c('Instrument','ms')
-  
-  g1<-ggplot2::ggplot(m,ggplot2::aes(x=reorder(x=Instrument,-ms,mean),y=ms,fill=Instrument))+
+
+  g1 <- ggplot2::ggplot(m,ggplot2::aes(x=reorder(x=Instrument,-ms,mean),y=ms,fill=Instrument))+
     ggplot2::geom_violin(scale = "width",show.legend = FALSE,na.rm=TRUE,alpha=0.7)+
     ggplot2::geom_jitter(height = 0, width = 0.15,size=0.15,show.legend = FALSE,na.rm=TRUE,colour='grey50')+
     ggplot2::stat_summary(fun = "mean",
@@ -53,30 +51,24 @@ plot_by_pair <-function(df=NULL,
     
     m$ms<-m$ms*1000
     
-    #colnames(m)<-c('Instrument','ms')
-    #head(m)
[…]

### 2022-12-23 R/plot_by_variable.R

@@ -14,7 +14,7 @@
 #' @import tidyr
 #' @import magrittr
 
-plot_by_variable <-function(df,
+plot_by_variable <- function(df,
                             meta='empty',
                             xlab='Tempo'){
 

### 2022-12-23 R/plot_timeline.R

@@ -4,8 +4,8 @@
 #'
 #' @param data data frame to be processed
 #' @param instr Instrument name to be processed
-#' @param Ampl Amplitude of the onsets (optional)
-#' @param Grouping Grouping factor (optional)
+#' @param ampl Amplitude of the onsets (optional)
+#' @param grouping Grouping factor (optional)
 #' @seealso \code{\link{plot_by_variable}}
 #' @return Graphic output
 #' @export
@@ -18,62 +18,54 @@
 plot_timeline <-
   function(data = NULL,
            instr = NULL,
-           Ampl = NULL,
-           Grouping = NULL) {
+           ampl = NULL,
+           grouping = NULL) {
     
     
-    ## debug start
-    #data <- CSS_Song2
-    #instr <- 'Tres'
-    #Ampl <- NULL
-    #Grouping <- 'Cycle'
-    ## debug end  
-    
-    
-    if (is.null(Ampl)) {
-      data$Ampl <- 1
+    if (is.null(ampl)) {
+      data$ampl <- 1
     }
-    if (is.null(Grouping)) {
-      data$Grouping <- 1
+    if (is.null(grouping)) {
+      data$grouping <- 1
     }
     
       
-    Onset <- NULL
+    onset <- NULL
     showlegend <- TRUE
     DF <- NULL
   #str(data)
-    if(!is.null(Grouping)){
+    if(!is.null(grouping)){
       for (k in 1:length(instr)) {
-        tmp <- dplyr::select(dplyr::ungroup(data), c(instr[k], Ampl, dplyr::all_of(Grouping)))
-        colnames(tmp)[1] <- 'Onset'
-        colnames(tmp)[2] <- 'Ampl'
-        colnames(tmp)[3] <- 'Grouping'
+        tmp <- dplyr::select(dplyr::ungroup(data), c(instr[k], ampl, dplyr::all_of(grouping)))
+        colnames(tmp)[1] <- 'onset'
+        colnames(tmp)[2] <- 'ampl'
+        colnames(tmp)[3] <- 'grouping'
         tmp$instr <- instr[k]
         DF <- rbind(DF, tmp)
       }
-      DF$Grouping <- factor(DF$Grouping)
+      DF$grouping <- factor(DF$grouping)
 
       # if more than 16 levels, remove factors
-      Grouping_Length <- length(levels(DF$Grouping))
-      Grouping_Length
-      if(Grouping_Length > 16){
-        DF$Grouping <- as.integer(DF$Grouping)
[…]

### 2022-12-23 R/summarise_sync_by_pair.R

@@ -1,6 +1,6 @@
 #' Calculate summary statistics for asynchronies by instrument pairs
 #'
-#' This function calculates simple t-test of the asynchronies of instruments (whether it differs from zero)
+#' This function calculates simple t-test of the asynchronies of instruments (whether they differ from zero)
 #'
 #' @param df data frame to be processed
 #' @param bybeat logical value whether beat information is used
@@ -20,18 +20,15 @@ summarise_sync_by_pair <- function(df,
   Instrument <- summarise <- across <- beatL <- Subdivision <- Asynchrony <- NULL
   
   if (bybeat == FALSE) {
-    # asynchronies
-    # m <-
-    #   suppressMessages(reshape2::melt(
-    #     df$asynch,
-    #     variable.name = 'Instrument',
-    #     value.name = 'ms'
-    #   ))
     m <- tidyr::pivot_longer(data=df$asynch,cols=c(1:dim(df$asynch)[2]),names_to = 'Instrument', values_to = 'ms')
     m$ms <- m$ms * 1000
+ #   print(head(m))
     # check whether instrument is different from 0
-    L <- levels(m$Instrument)
-    #    L
+#    L <- levels(m$Instrument)
+#    print(paste('Levels:',L))
+    L <- unique(m$Instrument)
+#    print(paste('Levels 2:',L))
+    
     T <- NULL
     for (k in 1:length(L)) {
       tmp <- m$ms[m$Instrument == L[k]]
@@ -50,7 +47,7 @@ summarise_sync_by_pair <- function(df,
       summarise(across(n, mean))
     SD <- df$asynch %>%
       summarise(across(n, sd))
-    
+
     T2 <- data.frame(M = t(M), SD = t(SD))
     T2$T <- T$tval
     T2$pval <- T$pval
@@ -60,43 +57,41 @@ summarise_sync_by_pair <- function(df,
     T2$M <- T2$M * 1000                            # milliseconds
     T2$SD <- T2$SD * 1000                          # milliseconds
     T2$pval <- scales::pvalue(T2$pval)            # prettify
+  #  print(T2)
   }
   
   
   if (bybeat == TRUE) {
-    #  print('by beat')
-    # m <-
-    #   suppressMessages(reshape2::melt(
-    #     df$asynch,
[…]

### 2022-12-23 R/sync_execute_pairs.R

@@ -7,27 +7,33 @@
 #'
 #' @param df data frame to be processed
 #' @param instruments Instrument names to be processed
-#' @param n_samp Number of samples to be drawn from the pool of joint onsets
-#' @param bnum How many bootstraps are drawn
-#' @param beats Beat structure to be included
+#' @param n Number of samples to be drawn from the pool of joint onsets
+#' @param bootn How many bootstraps are drawn (NULL default)
+#' @param beat Beat structure (subdivisions) to be included
 #' @return Output contain asynchronies and beat levels
 #' @seealso \code{\link{sync_sample_paired}}
 #' @import ggplot2
 #' @import tidyr
 #' @import magrittr
 
-sync_execute_pairs <- function(df=NULL,
-                               instruments=NULL,
-                               n_samp=100,
-                               bnum=1,
-                               beats=NULL){
+sync_execute_pairs <- function(df = NULL,
+                               instruments = NULL,
+                               n = 0,
+                               bootn = NULL,
+                               beat = NULL){
 
 # T. Eerola, Durham University, IEMP project
 # 14/1/2018  
-
+  
+  if(is.null(bootn)==TRUE){
+    bootn <- 1
+  }
+  
   # Get all combinations
-  c<-combn(instruments,2)
-  N<-dim(c)[2]
+  c <- combn(instruments,2)
+#  print(c)
+  N <- dim(c)[2]
+#  print(N)
   col_labels <- t(c)
   COL_LABELS<-NULL
   for(k in 1:N){
@@ -37,14 +43,15 @@ sync_execute_pairs <- function(df=NULL,
   DF<-NULL
   BE<-NULL
   for(k in 1:N){
-    DF <- cbind(DF,sync_sample_paired(df,c[,k][1],c[,k][2],n_samp,bnum,beats,FALSE)$asynch)
-    BE <- cbind(BE,sync_sample_paired(df,c[,k][1],c[,k][2],n_samp,bnum,beats,FALSE)$beatL)
+#    print(paste('k:',k,'1:',c[,k][1],'2:',c[,k][2],'3:',n,'4:',bootn,'5:',beat))
+    DF <- cbind(DF,sync_sample_paired(df,c[,k][1],c[,k][2],n,bootn,beat,FALSE)$asynch)
[…]

### 2022-12-23 R/sync_sample_paired.R

@@ -3,21 +3,21 @@
 #' taken N samples of two instruments (where they both have onsets)
 #'
 #' @param df data frame to be processed
-#' @param INSTR1 Instrument 1 name to be processed
-#' @param INSTR2 Instrument 2 name to be processed
-#' @param N Number of samples to be drawn from the pool of joint onsets. If 0, do not sample!
-#' @param BNum How many bootstraps are drawn
-#' @param beat Beat structure to be included
+#' @param instr1 Instrument 1 name to be processed
+#' @param instr2 Instrument 2 name to be processed
+#' @param n Number of samples to be drawn from the pool of joint onsets. If 0, do not sample (default)
+#' @param bootn How many bootstraps are drawn (default none, which is 1)
+#' @param beat Beat structure (subdivisions) to be included
 #' @param verbose Display no. of shared onsets (default FALSE)
 #' @seealso \code{\link{sync_execute_pairs}}
 #' @return List containing asynchronies and beat structures
 #' @export
 
 sync_sample_paired <- function(df = NULL,
-  INSTR1 = NULL,
-  INSTR2 = NULL, 
-  N = 100,
-  BNum = 1,
+  instr1 = NULL,
+  instr2 = NULL, 
+  n = 0,
+  bootn = NULL,
   beat = NULL,
   verbose = FALSE){
 
@@ -27,13 +27,17 @@ sync_sample_paired <- function(df = NULL,
 
   if(is.data.frame(df)==TRUE){
   
-  instr1 <- as.matrix(df[,which(colnames(df)==INSTR1)])  
-  instr2 <- as.matrix(df[,which(colnames(df)==INSTR2)])  
-  beat <- as.matrix(df[,which(colnames(df)==beat)])  
+  inst1 <- as.matrix(df[,which(colnames(df)==instr1)])  
+  inst2 <- as.matrix(df[,which(colnames(df)==instr2)])  
+  beat <- as.matrix(df[,which(colnames(df)==beat)])
+
+    if(is.null(bootn)==TRUE){
+      bootn <- 1
+    }
   
   D <- NULL
-  if(BNum==1){
-    ind<-!is.na(instr1) & !is.na(instr2)
+  if(bootn==1){
+    ind<-!is.na(inst1) & !is.na(inst2)
     len_joint<-length(which(ind))
     if(verbose==TRUE){
[…]

### 2022-12-23 README.Rmd
updated link
updated readme

--- 2022-11-07 README.Rmd
@@
 There are functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between performers, and assessing periodicity of the onsets, and visualising synchrony across cycles, time, or another property.
+
+For documentation, see [ Quick Guide](https://tuomaseerola.github.io/onsetsync/articles/onsetsync.html).
 
@@
 inst <- c('Clave','Bass','Guitar','Tres')              # Define instruments 
-dn <- sync_execute_pairs(CSS_Song2,inst,0,1,'SD')
-fig2 <- plot_by_pair(dn)  # plot
+dn <- sync_execute_pairs(CSS_Song2, inst, beat = 'SD') # Calculate ansychr.
+fig2 <- plot_by_pair(dn)                               # Plot
 print(fig2)  
@@
 N <- 200 # Let's select 200 onsets
-Bootstrap <- 1000
-d1 <- sync_sample_paired(CSS_Song2,'Clave','Bass',N,Bootstrap,'SD',TRUE)
-print(paste('Mean asynchrony of',round(mean(d1$asynch*1000),1),
-    'ms & standard deviation of',round(sd(d1$asynch*1000),1),'ms'))
+Bootstrap <- 100 # Let's bootstrap this 100 times
+d1 <- sync_sample_paired(df=CSS_Song2,
+                         instr1 = 'Clave',
+                         instr2 = 'Bass',
+                         n = N,
+                         bootn = Bootstrap,
+                         beat = 'SD')
+dplyr::summarise(data.frame(d1), N=n(), M = mean(asynch*1000), SD = sd(asynch*1000))
 ```
@@
 
-For more examples, see [Get started](docs/articles/onsetsync.html).
+For more examples, see [Get started](https://tuomaseerola.github.io/onsetsync/articles/onsetsync.html).

### 2022-12-27 doc/adding_annotations.Rmd
Updates

---
title: "Adding Annotations"
output: rmarkdown::html_vignette
description: >
  Learn how to add annotations to onsets.
vignette: >
  %\VignetteIndexEntry{Adding Annotations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Typically you have raw onset data when loading files with `onsetsync` which might not any other information within them (beat subdivisions, etc.). Here is an example of how you can add reference beat subdivisions and calculate reference beats such as mean of all onsets in a sub-division.

## Load raw onsets from OSF

```{r load, warning=FALSE,message=FALSE}
library(onsetsync)
require(httr)
CSS_Song2_Onset <- get_OSF_csv('8a347') # Get onsets
print(head(CSS_Song2_Onset[,1:8,]))
```

The data frame shows a portion of the onset data structure where the last three columns refer to the onset times of the specific instruments (`Clave`, `Bass`, and `Guitar`). Note that we only show the first eight columns and 6 rows. The first five columns are **meta-data**, referring to the piece, beats and cycles and sections. `Label.SD` combines cycle and beat information (first integer refers to cycle and the second integer refers to beat sub-division). `SD` represents the _sub-division_ explicitly. In this music there are 16 subdivisions per cycle (which can be felt as 4 beats $\times$ 4 subdivisions). `Section` indicates the specific part of the piece. Finally, in this example we have `Clave_` column, which is a reference to a clave pattern: strokes of the clave (woodblocks), whether sounded or not, are noted here; there are normally 5 strokes per cycle. The onset times for the instruments are recorded in seconds, although we prefer to use milliseconds (ms) in the analyses of synchrony to avoid reporting unnecessarily many digits. The onset times of the instruments are the result of an automatic onset detection (that has been subjected manual check) and the exact technique may depend on instrument used and source of data.

## Add annotations for cycles

```{r showannotations, echo=TRUE}
CSS_Song2_Metre <- get_OSF_csv('4cdpr') # Annotations for cycles
print(head(CSS_Song2_Metre))
```

Let's simplify this example (select only few columns).

```{r simplify_example, echo=TRUE}
CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
                  Label.SD,SD,Clave,Bass,Guitar,Tres) 
```

In the second data frame (`CSS_Song2_Metre`), shown in Table 2, we have annotations, namely the cycle times, which are the manually-estimated times when the cycles start (containing all sub-divisions of the beat). For instance, here the first cycle starts at 5.037333, the second cycle at 8.601000) and so on. This is important information for some type of analyses and often we want to combine the automatically detected onset times and the annotated cycles. It is important to realise that the information about the cycles comes from an annotation of the music by an expert (or in some cases by several experts). To get an annotation such as this, software such _Sonic Visualiser_ can be used to tap with the perceived beats of the audio track to obtain the beat times and then the first beat of the cycle is manually labelled to the data. The sub-divisions are then inferred from the cycle beginnings. This annotation process requires expertise with the musical style in question as the same patterns of onsets could potentially give rise to multiple interpretations of the metre.

## Combine onsets and annotations

As the onsets and annotations are in different files, we first combine the onsets and annotations with `onsetsync`. Here we first add annotations about the cycles into the onset data and then we also add isochronous beat times to the data frame, since are useful reference points for future calculations. The metre annotations have already informed the allocation of onsets to beat positions; here we add more information from the metre annotations to the dataframe to allow further analysis options.

```{r enrich1, echo=TRUE}
# Add annotations about the cycle to the data frame
CSS_Song2 <- add_annotation(df = CSS_Song2_Onset,
                            annotation = CSS_Song2_Metre$Cycle,
                            time = CSS_Song2_Metre$Time,
                            reference = 'Label.SD')
```

## Add isochronous beat subdivisions based on annotation

We add an isochronous reference subdivision of beat subdivisions based on annotated cycle start times.

```{r enrich2, echo=TRUE}
# Add isochronous beats to the data frame
CSS_Song2 <- add_isobeats(df = CSS_Song2, 
                          instr = 'CycleTime', 
                          beat = 'SD')
print(head(CSS_Song2[,2:9]))
```

At this point we have the basic elements for most analyses. In the data frame `CSS_Song2` we have the onset data (coded under instrument names, here `Clave`, `Bass`, and `Tres`) and information about the beat cycles (`CycleTime` and `Cycle`), and we also have timing information for isochronous beat divisions (`Isochronous.SD.Time`).

These annotation-based virtual beat structures can be systematically off by a certain amount, so perhaps some it may be advantageous for some analyses to infer the timing information related to the beat structure from the existing onsets themselves and not rely on external reference. `add_isobeats` function can accomplish this if given the instruments from which it will calculate the mean onset time for the first beat of the cycle, which will then be used to set the isochronous beat timings for the rest of the beats within the cycle.

## Add isochronous beat subdivisions based on mean onset times

```{r enrich3, echo=TRUE}
colnames(CSS_Song2)[9] <- 'Ann.Iso' # Rename
# Add isochronous beats based on mean timing of guitar, tres, and clave

CSS_Song2 <- add_isobeats(df = CSS_Song2, 
                          instr = c('Guitar','Tres','Clave'), 
                          beat = 'SD')
# Show the newly calculated isochronous beat times 
# from the second cycle onwards.
print(head(CSS_Song2[17:22,c(2,6:10)]))
```

These operations allow many to create a number of variant reference points to be used in the calculation of synchrony.

### 2022-12-27 doc/synthesise_onsets.Rmd
Updates

---
title: "Synthesise onsets"
output: rmarkdown::html_vignette
description: >
  Create audio files from the onset data.
vignette: >
  %\VignetteIndexEntry{Synthesise onsets}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

To create an audio file of your onset data, you can call `synthesise_onsets` function.

## Example data

Load a Cuban salsa example. Let's take a 10 second example. To eliminate the first 5, where the instruments are not playing, let's redefine the time to start from the first annotated cycle.

```{r load, warning=FALSE,message=FALSE}
library(onsetsync)
data <- CSS_IEMP[[2]]                                                          # example
data <- dplyr::filter(data,Isochronous.SD.Time > 5 & Isochronous.SD.Time < 15) # 10 s
data$Guitar <- data$Guitar - min(data$Isochronous.SD.Time,na.rm = TRUE)        # trim time     
data$Bass <- data$Bass - min(data$Isochronous.SD.Time,na.rm = TRUE)            # trim time
```

## Synthesise Guitar and

Let's take two instruments and create their onsets as high-pitched sound (Guitar) and short noise burst (Bass).

```{r createwave, echo=TRUE}
inst <- c('Guitar','Bass')
fs <- 22050
w <- synthesise_onsets(data=data,instruments = inst, sr = fs, type =c('synth','noise'))
```

Plot the signal. This is downsampled for plotting in `ggplot2`.

```{r, echo=TRUE,fig.width=6.5,fig.height=2.5}
scaling <- 20
ws <- signal::resample(w,fs,fs*scaling)
tmp <- data.frame(Amplitude=ws,Time=seq(0,length(ws)-1))
tmp$Time <- tmp$Time/fs*scaling
ggplot2::ggplot(tmp,ggplot2::aes(x=Time,y=Amplitude))+
  ggplot2::geom_line(color='blue')+
  ggplot2::theme_linedraw()
```

Save as wave file using `seewave` library.

```{r savefile, echo=TRUE, eval=FALSE}
seewave::savewav(wave = w, f = fs, channel = 1, filename = 'example.wav')
```

Further options relate to sound generators and frequency.

### 2022-12-27 R/plot_by_beat.R

@@ -26,7 +26,7 @@ plot_by_beat <-
            griddeviations = FALSE,
            boxplot = FALSE,
            colour = 'lightblue',
-           colourpalette = 'Set1',
+           colourpalette = 'Set2',
            pointsize = 1) {
     # T. Eerola, Durham University, IEMP project
     # 23/1/2018

### 2022-12-27 R/plot_by_dataset.R

@@ -8,6 +8,7 @@
 #' @param pcols Number of columns for multiple plots (default 1)
 #' @param box Do the graphics by boxplot
 #' @param colour colour for the boxplot
+#' @param colorpalette colourpalette for the boxplot
 #' @seealso \code{\link{plot_by_variable}}, \code{\link{plot_by_pair}}, \code{\link{plot_by_beat}}
 #' @return Graphic output
 #' @import ggplot2
@@ -20,7 +21,8 @@ plot_by_dataset <-
            data = NULL,
            pcols = 1,
            box = FALSE,
-           colour = 'lightblue') {
+           colour = 'lightblue',
+           colorpalette = 'Set2') {
     # T. Eerola, Durham University, IEMP project
     # 23/1/2018
     # needs work
@@ -45,7 +47,7 @@ plot_by_dataset <-
       ggplot2::ggplot(DF, ggplot2::aes(asynch, fill = data), colour = 'black') +
       ggplot2::geom_density(alpha = 1, show.legend = FALSE) +
       ggplot2::scale_color_brewer(name = "Dataset",
-                                  palette = 'Set1',
+                                  palette = colorpalette,
                                   type = "div") +
       ggplot2::facet_wrap( ~ data, pcols) +
       ggplot2::xlab(paste('Asynchrony (ms)')) +
@@ -67,7 +69,7 @@ plot_by_dataset <-
                   # ggplot2::scale_color_brewer(name = "Dataset",
         #                             palette = 'Set1',
         #                             type = "div") +
-        ggplot2::scale_fill_brewer(palette = "Set1") +
+        ggplot2::scale_fill_brewer(palette = colorpalette) +
         ggplot2::xlab('Performance') +
         ggplot2::ylab('Asynchrony (ms)') +
         ggplot2::coord_flip()+

### 2022-12-27 R/plot_by_pair.R

@@ -15,7 +15,7 @@
 plot_by_pair <-function(df = NULL,
                         bybeat = FALSE,
                         reference = 0,
-                        colourpalette = 'PuOr'){
+                        colourpalette = 'Set2'){
 
   # T. Eerola, Durham University, IEMP project
   # 14/1/2018  
@@ -29,7 +29,9 @@ plot_by_pair <-function(df = NULL,
   m <- tidyr::pivot_longer(data=df$asynch,cols=c(1:dim(df$asynch)[2]),names_to = 'Instrument', values_to = 'ms')
   m$ms <- m$ms*1000
 
-  g1 <- ggplot2::ggplot(m,ggplot2::aes(x=reorder(x=Instrument,-ms,mean),y=ms,fill=Instrument))+
+  m$Instrument <- factor(m$Instrument)
+
+  g1 <- ggplot2::ggplot(m,ggplot2::aes(x=Instrument,y=ms,fill=Instrument))+
     ggplot2::geom_violin(scale = "width",show.legend = FALSE,na.rm=TRUE,alpha=0.7)+
     ggplot2::geom_jitter(height = 0, width = 0.15,size=0.15,show.legend = FALSE,na.rm=TRUE,colour='grey50')+
     ggplot2::stat_summary(fun = "mean",
@@ -48,21 +50,19 @@ plot_by_pair <-function(df = NULL,
   if(bybeat==TRUE){
     #m <- suppressMessages(reshape2::melt(df$asynch,variable.name='Instrument',value.name='ms'))
     m <- tidyr::pivot_longer(data=df$asynch,cols=c(1:dim(df$asynch)[2]),names_to = 'Instrument', values_to = 'ms')
-    
     m$ms<-m$ms*1000
-    
     b <- tidyr::pivot_longer(data=df$beatL,cols=c(1:dim(df$beatL)[2]),names_to = 'variable', values_to = 'value')
-    
     
     m$beatL <- b$value
     m$beatL <- factor(m$beatL)
-
+    m$Instrument <- factor(m$Instrument)
+    
     # Make sure there are sufficient number of colour palette categories for subdivs
     colpal <- RColorBrewer::brewer.pal(4,colourpalette)
     number_of_beats <- length(unique(m$beatL))
     colourpalette <- colorRampPalette(colpal)(number_of_beats)
 
[…]

### 2022-12-27 R/sync_execute_pairs.R

@@ -25,8 +25,13 @@ sync_execute_pairs <- function(df = NULL,
 # T. Eerola, Durham University, IEMP project
 # 14/1/2018  
 
+
   if(is.null(bootn)==TRUE){
     bootn <- 1
+  }
+
+  if(n < bootn & bootn>1){
+    stop("More bootstraps (bootn) are specified than samples (n)!", call. = FALSE)
   }
   
   # Get all combinations

### 2022-12-27 R/sync_sample_paired_relative.R

@@ -1,0 +1,28 @@
+#' Compare instrument onset times to mean onset times from several instruments
+#'
+#' @param df data frame to be processed
+#' @param instr Instrument name to be processed
+#' @param instr_ref Instrument reference names to be processed
+#' @param beat Beat structure (subdivisions) to be included
+#' @seealso \code{\link{sync_execute_pairs}}
+#' @return difference in ms 
+#' @export
+
+sync_sample_paired_relative <- function(df = NULL,
+  instr = NULL,
+  instr_ref = NULL, 
+  beat = NULL){
+
+  Isochronous.SD.Time <- NULL
+  
+  # T. Eerola, Durham University, IEMP project
+  # 21/12/2022  
+
+  df <- dplyr::select(df,-Isochronous.SD.Time)
+  df <- add_isobeats(df = df, instr = instr_ref, beat = beat)
+
+  # then compare onset times to this new reference
+  df2 <- sync_sample_paired(df,instr1 = instr, instr2 = 'Isochronous.SD.Time', beat = beat)
+  output <- summarise_sync(df2)
+  return <- output 
+}

### 2022-12-27 R/synthesise_onsets.R

@@ -1,0 +1,136 @@
+#' Synthesise onsets with noises and clicks
+#'
+#' Take the specified instruments and create wave file with sounds where the onsets are
+#'
+#' @param data data frame to be processed
+#' @param instruments Instruments names to be included
+#' @param sr sampling rate (default 22050)
+#' @param dur duration of sounds (default 30 ms)
+#' @param type type of sounds (default noise, try click or synth from seewave)
+#' @param freq freq of sounds (if synth is given as input)
+#' @import seewave
+#' @return wave file
+#' @export
+
+synthesise_onsets <-
+  function(data = NULL,
+           instruments  = NULL,
+           sr = 22050,
+           dur = 0.03,
+           type = 'noise',
+           freq = 2500) {
+    WAVE <- NULL
+    # T.E. 25/12/2022
+    
+    # # # # DEBUG
+    #  data<-data.frame(Guitar=c(0.1,0.3,0.5,0.7),Bass=c(0.0,NA,0.5,1.0))
+    #  instruments<-c('Guitar','Bass')
+    #  sr <- 22050
+    #  dur <- 0.1
+    #  type<-'click'
+    # # # # DEBUG
+
+noisew <- NULL
+
+    # define a function to be used when merging different instrument onsets        
+    absmax <- function(x) {
+      x[which.max(abs(x))]
+    }
+    
+    # if less type parameters, create as many as instruments
+    input_nro <- length(instruments) - length(type)
+    if (input_nro != 0) {
+      type <- rep(type[1], length(instruments))
+    }
+    
+    ## arbitrary pitches
+    FR <- seq(200, 2000, by = 40)
+    if (is.null(freq) == TRUE) {
+      freq <- FR[1:length(instruments)]
+    }
+    
+    # trim data only to instruments specified
+    data <- dplyr::select(data,dplyr::all_of(instruments))
+
+    # create a vector for the whole track
+    maxdur <- max(data, na.rm = T) + dur
+#    print(paste("maxdur:",maxdur))
+#    print(paste("fs:",sr))
+#    print(paste("tracks:",length(instruments)))
+    #  WAVE <- rep(0, maxdur * fs) # WAVE for each instrument
+    WAVE <-
+      matrix(0, (maxdur * sr), length(instruments)) # WAVE for each instrument
+    # Loop across instruments
+    for (k in 1:length(instruments)) {
+   #    print(paste("instrument:",instruments[k]))
+  #     print(paste("type:",type[k]))
[…]

### 2022-12-27 doc/adding_annotations.R

@@ -1,0 +1,39 @@
+## ----load, warning=FALSE,message=FALSE----------------------------------------
+library(onsetsync)
+require(httr)
+CSS_Song2_Onset <- get_OSF_csv('8a347') # Get onsets
+print(head(CSS_Song2_Onset[,1:8,]))
+
+## ----showannotations, echo=TRUE-----------------------------------------------
+CSS_Song2_Metre <- get_OSF_csv('4cdpr') # Annotations for cycles
+print(head(CSS_Song2_Metre))
+
+## ----simplify_example, echo=TRUE----------------------------------------------
+CSS_Song2_Onset <- dplyr::select(CSS_Song2_Onset,
+                  Label.SD,SD,Clave,Bass,Guitar,Tres) 
+
+## ----enrich1, echo=TRUE-------------------------------------------------------
+# Add annotations about the cycle to the data frame
+CSS_Song2 <- add_annotation(df = CSS_Song2_Onset,
+                            annotation = CSS_Song2_Metre$Cycle,
+                            time = CSS_Song2_Metre$Time,
+                            reference = 'Label.SD')
+
+## ----enrich2, echo=TRUE-------------------------------------------------------
+# Add isochronous beats to the data frame
+CSS_Song2 <- add_isobeats(df = CSS_Song2, 
+                          instr = 'CycleTime', 
+                          beat = 'SD')
+print(head(CSS_Song2[,2:9]))
+
+## ----enrich3, echo=TRUE-------------------------------------------------------
+colnames(CSS_Song2)[9] <- 'Ann.Iso' # Rename
+# Add isochronous beats based on mean timing of guitar, tres, and clave
+
+CSS_Song2 <- add_isobeats(df = CSS_Song2, 
+                          instr = c('Guitar','Tres','Clave'), 
+                          beat = 'SD')
+# Show the newly calculated isochronous beat times 
+# from the second cycle onwards.
+print(head(CSS_Song2[17:22,c(2,6:10)]))
+

### 2022-12-27 doc/onsetsync.R

@@ -1,9 +1,50 @@
-## ---- include = FALSE---------------------------------------------------------
-knitr::opts_chunk$set(
-  collapse = TRUE,
-  comment = "#>"
-)
+## ----eval=FALSE---------------------------------------------------------------
+#  if (!require(devtools)) install.packages("devtools")
+#  devtools::install_github("tuomaseerola/onsetsync")
 
-## ----setup--------------------------------------------------------------------
+## ---- message=FALSE, warning=FALSE,echo=TRUE----------------------------------
 library(onsetsync)
+library(dplyr)
+library(ggplot2)
+packageVersion("onsetsync")
+
+## ----message=FALSE,warning=FALSE,eval=TRUE,results='asis'---------------------
+CSS_Song2 <- onsetsync::CSS_IEMP[[2]]   # Read one song from internal data
+CSS_Song2 <- dplyr::select(CSS_Song2,Label.SD,SD,Clave,Bass,Guitar,Tres,
+                           CycleTime,Cycle,Isochronous.SD.Time) # Select some columns
+print(knitr::kable(head(CSS_Song2),format = "simple",digits = 2))
+
+## ----synch2isochron,fig.width=7, fig.asp=.75----------------------------------
+fig1 <- plot_by_beat(df = CSS_Song2, 
+                     instr = c('Bass','Clave','Guitar','Tres'), 
+                     beat = 'SD', 
+                     virtual='Isochronous.SD.Time',
+                     pcols=2)
+print(fig1)
+
+## ----fig2,warning=FALSE,fig.width=7, fig.asp=.7-------------------------------
+inst <- c('Clave','Bass','Guitar','Tres') # Define instruments 
+dn <- sync_execute_pairs(CSS_Song2,inst,beat = 'SD')
+fig2 <- plot_by_pair(dn)  # plot
+print(fig2)  
+
+## ----paired1------------------------------------------------------------------
+set.seed(1234) # set random seed
+N <- 200 # Let's select 200 onsets
+Bootstrap <- 100
+d1 <- sync_sample_paired(df = CSS_Song2, 
+                         instr1 = 'Clave',
+                         instr2 = 'Bass',
+                         n = N, 
+                         bootn = Bootstrap,
+                         beat = 'SD')
[…]

### 2022-12-27 doc/synthesise_onsets.R

@@ -1,0 +1,24 @@
+## ----load, warning=FALSE,message=FALSE----------------------------------------
+library(onsetsync)
+data <- CSS_IEMP[[2]]                                                          # example
+data <- dplyr::filter(data,Isochronous.SD.Time > 5 & Isochronous.SD.Time < 15) # 10 s
+data$Guitar <- data$Guitar - min(data$Isochronous.SD.Time,na.rm = TRUE)        # trim time     
+data$Bass <- data$Bass - min(data$Isochronous.SD.Time,na.rm = TRUE)            # trim time
+
+## ----createwave, echo=TRUE----------------------------------------------------
+inst <- c('Guitar','Bass')
+fs <- 22050
+w <- synthesise_onsets(data=data,instruments = inst, sr = fs, type =c('synth','noise'))
+
+## ---- echo=TRUE,fig.width=6.5,fig.height=2.5----------------------------------
+scaling <- 20
+ws <- signal::resample(w,fs,fs*scaling)
+tmp <- data.frame(Amplitude=ws,Time=seq(0,length(ws)-1))
+tmp$Time <- tmp$Time/fs*scaling
+ggplot2::ggplot(tmp,ggplot2::aes(x=Time,y=Amplitude))+
+  ggplot2::geom_line(color='blue')+
+  ggplot2::theme_linedraw()
+
+## ----savefile, echo=TRUE, eval=FALSE------------------------------------------
+#  seewave::savewav(wave = w, f = fs, channel = 1, filename = 'example.wav')
+

### 2023-01-01 README.Rmd
minor

@@
 
-```{r synch2isochron,fig.width=7, fig.asp=.75}
+```{r synch2isochron,fig.width=7.4, fig.asp=.75}
 fig1 <- plot_by_beat(df = CSS_Song2, 
@@
 
-```{r fig2,warning=FALSE,fig.width=7, fig.asp=.7}
+```{r fig2,warning=FALSE,fig.width=7.4, fig.asp=.7}
 inst <- c('Clave','Bass','Guitar','Tres')              # Define instruments 

### 2023-01-17 paper/paper.Rmd
Added paper directory with contents.

---
# Example from https://joss.readthedocs.io/en/latest/submitting.html
title: 'onsetsync: An R Package for Onset Synchrony Analysis'
tags:
  - R
  - music
  - entrainment
  - periodicity
  - synchrony
authors:
  - name: Tuomas Eerola
    orcid: 0000-0002-2896-929X
    affiliation: 1 # (Multiple affiliations must be quoted)
  - name: Martin Clayton
    orcid: 0000-0002-9670-5077
    affiliation: 1
affiliations:
 - name: Department of Music, Durham University
   index: 1
citation_author: Eerola et. al.
date: 17 January 2023
year: 2023
bibliography: paper.bib
output: rticles::joss_article
csl: /Users/tuomaseerola/Documents/computational/R/under_dev/onsetsync_DEV/JOSS/apa.csl
citation_package: "default"
keep_tex: TRUE
journal: JOSS
---

# 1 Introduction

Music performance relies on tight yet flexible timing coordination
between performers. An important aspect of this interpersonal
entrainment is the synchronisation of musical events between performers,
and this can be analysed from the note onsets obtained from recorded
performances. Analysis of the asynchronies between onsets gives an
insight into synchronisation between performers, and the extent to which
it is accurate (how close together the events are in time) and precise
(how stable the relationship is). The synchrony between performers is
influenced by various factors such as the genre of music, performer
skill level, intention, and phrase and beat structures of the music. The
analysis of synchrony in music benefits from shared tools as there are a
number of common operations that need to be carried out in every dataset
(e.g., calculating pairwise synchrony, or assessing the
synchrony across other variables such as tempo, metrical hierarchy, or
phrasing) and the number of datasets containing onset timing information has recently increased.

## 1.1 Statement of need

`onsetsync` is a R package for assessment of musical synchrony
through note onsets. Data is ingested in the form of CSV files with individual instrument onset times are aligned with labelled beat positions; separate metre annotations comprising lists of time estimates for the downbeats are also used for some functions. The package includes functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, assessing the periodicity of the onsets, and visualising synchrony across metrical cycles, time, or another property. These functions make the analysis of onset corpora transparent and will allow more scholars to carry out investigations of entrainment in music.

## 1.2 Onset datasets

`onsetsync` package comes with example performances from two traditions, Cuban Son and Salsa performances [@Poole_Tarsitani_Clayton_2022] and North Hindustani performances [@Clayton_Leante_Tarsitani_2021]. The examples are taken from the Interpersonal Entrainment in Music Performance (IEMP) project collection that has compiled small ensemble performance data from six traditions (North Indian ragas, Malian jembe, Tunisian stambeli, Uruguayan candombe, European string quartet, and Cuban son and salsa). These collection contain tens of hours annotated and verified onsets and are available from Open Science Framework (OSF). There are also noteworthy datasets containing audio and annotated contents of music relevant to assessing synchrony such as _Carnatic Music Rhythm Dataset_ [@srinivasamurthy2015particle], _Hindustani Music Rhythm Dataset_ [@srinivasamurthy2016generalized], _Arab-Andalusian (Maghreb) dataset_ [@sordo2014creating], _Scandinavian Fiddle Tunes with Accompanying Foot-Tapping_ [@lordelo2020adversarial] and _MAESTRO (MIDI and Audio Edited for Synchronous Tracks and Organisation)_ dataset which is an extensive collection (200+ hours) of professional piano performances [@hawthorne2018enabling]. Some datasets such as Chopin performances compiled by [@goebl2010investigations] facilitate study of the synchrony between the two hands of the pianist.

## 1.3 Terminology and measures of synchrony

Much of the research on synchrony in music has focused on
_sensorimotor synchronization_ (SMS) or just synchronization, a
processes occurring within 100-2000 ms timescales, which has commonly
been studied with tapping experiments [@repp2008sensorimotor, @repp2013sensorimotor] but also through analyses of music corpora, from string quartets [@wing2014optimal] to drum ensembles [@polak2016both] and to
Indian instrumental music performances [@clayton_et_al_2018b]. Synchronisation analyses can also be applied to field observations
such as the study of how walking drumming groups attempt to
resist synchronisation to the beats of other drum groups during their
walk across the city during a festival [@lucas2011inter].

A number of different measures of synchrony have been proposed. We
follow the definitions from [@clayton2020, p. 161] who adopt and
develop the measures defined by [@rasch1988timing, @rasch1979synchronization]:

Here we first define the onset time differences, $d$, where $I_{1, i}$ and $I_{2, i}$ refer to an onset at time point $i$ for instrument 1 and 2:

\begin{equation}
d_{i} = I_{1, i} - I_{2, i}
\end{equation}

**Precision measures** estimate how consistent the two parts are:

*Pairwise asynchronization*: standard deviation of the
  asynchronies of any pair of instruments, which is

\begin{equation}
Sd = \sqrt\frac{\sum{(d_i-\bar{d})^2}}{n-1}
\end{equation}

*Groupwise asynchronization*: Root mean square (RMS) of the
  pairwise asynchronizations calculated as

\begin{equation}
RMS = \sqrt{\Sigma_{i=1}^{T}{\Big(\frac{d_{1,i} -d_{2,i}}{T}\Big)^2}}
\end{equation}

*Mean absolute asynchrony*: Mean of all unsigned asynchrony
  values which is

\begin{equation}
\hat{|s|} = \frac{1}{n} \sum_{i=i}^{n} |d_{i}|
\end{equation}

**Accuracy measures** estimate how far one part is ahead of or behind
the other(s):

*Mean pairwise asynchrony*: mean difference in the onset values
  of two instruments, i.e., the signed asynchrony values, as in

\begin{equation}
\hat{s_p} = \frac{1}{n} \sum_{i=i}^{n} d_{i}
\end{equation}

*Mean relative asynchrony*: mean position of an instrument's
  onsets relative to average position ($\hat{w}$) of the group, where the average position is

\begin{equation}
\hat{w} = \frac{w_{1} + w_{2} + \ldots + w_{n}}{n}
\end{equation}

and the relative onset time $w$ is

\begin{equation}
v_n = w_n - \hat{w}
\end{equation}

and therefore mean relative asynchrony is

\begin{equation}
\hat{s_r} = \frac{1}{n} \sum_{i=i}^{n} v_{i}
\end{equation}

These methods are based on asynchronies between onsets occurring at the same beat positions; other methods exist for assessing synchrony such as utilising circular statistics [@berens2009circstat], recurrence quantification analysis [@wallot2018analyzing], or calculating the synchrony of spike trains [@kreuz2007measuring]. `onsetsync` is focussed on measures of asynchrony relying on onsets between instruments at the same beat positions, which is usually the relevant approach in music sense.

## 1.4 Availability

```{r secret_load, eval=TRUE, echo=FALSE}
# LOADS THIS SECRECTLY FROM THE LOCAL FOLDER
library(onsetsync)
ver <- packageVersion('onsetsync')
```

`onsetsync` is available at Github and can be loaded and installed using the code below. The current version is `r ver`.

```{r availability,eval=FALSE}
library(devtools)
devtools::install_github("tuomaseerola/onsetsync")
library(onsetsync)
print(packageVersion('onsetsync'))
```

In order to be able to build the carry out such comparisons, `onsetsync` has several auxiliary functions. There are functions that help assessing relative timings, to associate onsets with beat subdivisions and cycles in music structure, and to select and compare the instruments. The library has seven categories of functions; (1) input/output, (2)
annotation, (3) visualisation, (4) synchrony, (5) periodicity, (6)
summary, and (7) dataset functions (see Table 1). The functions are named to be easily associated with the categories.

| Category      | Examples                                             |
|---------------|------------------------------------------------------|
| Input/Output  | `get_OSF_csv`, `synthesise_onsets`                   |
| Annotation    | `add_annotation`, `add_isobeats`                     |
| Visualise     | `plot_by_beat`, `plot_by_pair`, `plot_by_variable`   |
| Synchrony     | `sync_sample_paired`, `sync_execute_pairs`           |
| Periodicity   | `periodicity`, `period_to_BPM`, `period_nPVI`        |
| Summarise     | `summarise_onsets`, `summarise_sync`                 |
| Datasets      | `Asere_OU_2`, `DebBh_Drut`, `CSS_IEMP`               |

Table: Function categories and example functions or datasets.

In the next sections we will demonstrate the analysis processes related to asynchrony between several performers. It is also worth pointing out that the library is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g. [Librosa](https://librosa.org), [Essentia](https://essentia.upf.edu), or [MIR toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms. Here we take it as granted that the onsets have been extracted already, someone has carried out a quality check, assigned relevant onsets to beat positions and saved them in csv files. For solid overviews on computational onset extraction from audio, see [@schluter2014improved] and [@bock2016madmom], and for a full workflow of how to combine onset detection and annotation of the musical information, see [@Clayton2020emr].

## 1.5 Onset representation

Here we will take one Cuban Son performance that has been included in the library. This song, _Palo Santo_, has been performed by seven performers using the following combination of instruments: bass, guitar, tres, and trumpet, and five percussion instruments, clave, bongo, bell, cajon, and conga (the instrumentation varies slightly between songs and sections within songs). The full data including other Cuban salsa and son performances is available from an open access repository at _Open Science Framework_ (OSF) and has been annotated and processed as part of the [Interpersonal Entrainment in Music Performance](https://musicscience.net/projects/timing/iemp/) project, available at [https://osf.io/sfxa2/](https://osf.io/sfxa2/).

The code example loads the onset data and annotations of metre of the son performance. The library comes with a set of five Cuban salsa and son (CSS) performances and this is the second example from the IEMP `CSS_IEMP` mini-corpus. The second line of the code selects the nine columns of interest for the analysis; three first are meta-data, followed by four instruments (selected to aid legibility of the tables and charts), and two timing related columns are included at the end.

```{r, results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library(dplyr)
library(ggplot2)
```

```{r getdata1,message=FALSE,warning=FALSE,results='asis'}
CSS_Song2 <- dplyr::select(onsetsync::CSS_IEMP[[2]],
  Piece, Section, Clave, Bass, Guitar, Tres, 
  SD, Cycle, Isochronous.SD.Time)
print(knitr::kable(head(CSS_Song2),digits = 2,
  caption = 'Onset data structure example.'))
```

Table 2 shows a portion of the onset data structure, just 6 rows out of 1568 rows in the data containing 3286 onsets. The first two columns are
*meta-data*, referring to the piece and section. The next four columns represent the onset times in seconds (rounded to two decimals for convenience) of four selected instruments, `Clave`, `Bass`, `Guitar` and `Tres`. The onset times for the instruments are available in seconds (with precision of 6 digits, although we prefer to use milliseconds (ms) in the analyses of synchrony to avoid reporting
unnecessarily many digits. The last three columns refer to information about the relative time: `SD` represents the beat subdivision within the cycle. In this music there are 16 subdivisions per cycle (which can be felt as 4 beats \(\times\) 4 subdivisions). `Cycle` refers to a running number of the beat cycles (each will have 16 beats in this music). `Isochronous.SD.Time` is an evenly temporally spaced time point for each beat subdivision (calculated by dividing each cycle duration by 16) that can act as temporal reference grid. The beat subdivisions and cycles are manually annotated and can be integrated to the data frame with a specific function (`add_annotation`) if given separately. Also a reference timing (`Isochronous.SD.Time`) has been provided for the cycles but it can also be estimated from the cycles and beat subdivisions using a function (`add_isobeats`) in the library. Information about cycles and beat subdivisions is generally based on manual annotation, and originates separately from the automatically extracted onsets; the two are combined to produce tables with onsets assigned to beat positions. It is also possible to read the csv files directly from OSF using `get_OSF_csv` function to allow transparent and reproducible analysis workflows with published datasets.

## 1.6 Onset summaries

Let's explore the overall structure of the onsets in the example piece.

```{r summarise,message=FALSE,warning=FALSE,results='asis'}
tab1 <- summarise_onsets(df = CSS_Song2, 
  instr = c('Clave','Bass','Guitar','Tres','Isochronous.SD.Time'))
print(knitr::kable(tab1,digits = 1,
  caption = 'Summary of onsets in the example data 
  (Md, M, Min, and Max in ms).'))
```

Table 3 shows the summary of the onsets, namely the number of onsets for each instrument, the descriptive statistics referring to time difference of successive onsets ($d$) within each instrument. We have provided the isochronous beat subdivisions as a reference, which shows how many beats there are and what the is the average duration of the beats (220 ms) in this performance lasting approximately 5 minutes and 45 seconds.

As a broad overview, we can use the function `plot_by_beat` visualise the asynchrony relative to an equal division subdivision of the beats for each instrument across the time. This visualisation gives a summary about when the instruments are playing relative to the beat division and the sections of the song, shown in Figure 1. For example, in many genres different instruments contribute to specific subdivision of the beat. This is of course the case on Cuban Son, but maybe even more extreme in Malian Jembe music [@polak2010rhythmic] or more variable in Western jazz [@wesolowski2016timing].

```{r synch2isochron,fig.width=7.0, fig.height=5.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
fig1 <- plot_by_beat(df = CSS_Song2, 
                     instr = c('Bass','Clave','Guitar','Tres'), 
                     beat = 'SD', 
                     virtual='Isochronous.SD.Time',
                     pcols=2)
print(fig1)
```

At first glance of Figure 1, it is easy to spot the distinctive clave pattern going through the song (the beginning of the song is at the bottom of the graphs, and every cycle is cascaded on top of the previous cycle). Notice how the bass plays mainly on subdivisions 5, 7, 13, 15, except in the last minute when the pattern changes into playing subdivisions 3, 4, 7, 8, 15, 16. The guitar plays almost on every beat subdivision, and towards the end of the piece tres --  a cuban guitar with 6-strings tuned to 3 pitches -- plays a pattern of 3 subdivisions and a break. Overall, this six-minute piece does not contain a high amount of variation in the patterns of onsets across the subdivisions. We can see that the bass elaborates the beats during the last minute or so while the clave, guitar and tres keep playing their patterns throughout the piece.

# 2 Analysis of synchrony

Moving on to the analysis of synchrony, we can first explore how much the onsets deviate from the isochronous beats or from the mean onset times. Figure 2 shows an example of the former for two guitars, guitar and tres.

```{r, fig.cap='Relative timing deviations from beat sub-divisions across the performance for guitar and tres.'}
print(plot_by_beat(df = CSS_Song2,
  instr = c("Guitar","Tres"),
  beat = "SD",
  virtual = "Isochronous.SD.Time", 
  griddeviations = TRUE))
```

Overall this suggests that these two instruments tend to play earlier when compared to the isochronous beat division of the cycle, although guitar is tightly aligned (<2%) with the reference beat on every fourth beat subdivisions (subdivisions of 1, 5, 9, and 13), which are also the salient metrical positions. Similar fluctuations across the beat subdivisions are not evident in tres, but the graph reveals that the tres tends to play a little earlier than guitar.

## 2.1 Synchrony between the instruments

Let's determine the overall synchrony between specific instruments. Here we continue the analysis guitar and tres and calculate the asynchrony between them.

```{r}
d1 <- sync_sample_paired(CSS_Song2, instr1 = "Guitar", instr2 = "Tres", 
  beat = "SD")
dplyr::summarise(data.frame(d1), 
  N = n(), Mean.ms = mean(asynch*1000), Sd.ms = sd(asynch*1000))
```

This analysis indicates that on average, the tres plays 13 ms ahead of the guitar, with a standard deviation of 27 ms. When one is comparing instruments that have radically different number of joint onsets (such as clave and bass in this example), it is possible to make the comparisons between the pairs of instruments easier by specifying a sample of onsets that will be taken from each instrument for the analysis. It also possible to establish the confidence intervals by bootstrapping the asynchrony calculations.

To carry out the comparison for all possible pairings of the instruments is possible with `sync_execute_pairs` function and the results can be visualised with a related function (`plot_by_pair`).

```{r,fig.cap='Asynchronies across multiple instrument pairs.'}
inst <- c("Clave","Bass","Guitar","Tres")
dn <- sync_execute_pairs(CSS_Song2, inst, beat = "SD")
print(plot_by_pair(dn))
```

In Figure 1 we see how different pairs of instruments have different asynchrony relationships; the bass is consistently ahead of the guitar, and clave and guitar play behind tres, to pick some of the extreme examples from the visualisation. The vertical lines stand for median asynchrony and the violin plot shows the distribution of onset time differences.

We can calculate the classic measures of synchronization once the pairing has been done. Here are these measures calculated for the synchrony between bass and clave.

```{r}
d <- sync_sample_paired(CSS_Song2,"Clave","Bass", beat = "SD")
print(t(summarise_sync(d)))
```

The output suggests that at least in this piece, these two instruments tend to have the synchronization precision around 20 ms, depending on the method of calculation. Pairwise asynchronisation is the standard deviation of the signed asynchrony; this tends to be highly correlated with the mean absolute asynchrony. The Mean pairwise asynchrony measure shows that on average, the clave plays 16 ms after the bass.

We can also calculate the relative asynchrony by comparing the onset times to the mean of the other instruments that we define. Here we compare the asynchrony of bass and clave to the mean of the rhythm section as defined by four instruments, guitar, bass, tres, and clave.

```{r}
dl <- sync_sample_paired_relative(df = CSS_Song2, 
  instr = "Bass", instr_ref = c("Guitar","Bass","Tres","Clave"),
  beat = "SD")
print(dl$`Mean pairwise asynchrony`)

dl <- sync_sample_paired_relative(df = CSS_Song2,
  instr = "Clave", instr_ref = c("Guitar","Bass","Tres","Clave"),
  beat = "SD")
print(dl$`Mean pairwise asynchrony`)
```

The relative measure (mean relative asynchrony) shows that bass is ahead (ie. -19 ms) of the mean of other instruments whereas clave is almost exactly in synchrony with the other instruments (<1 ms late). In different types of performances, musical pieces, instruments, and genres, the magnitude of these differences is highly variable [@clayton2020].

## 2.2 Synchrony between the instruments across beat subdivisions

In the first visualisation we already observed that asynchronies varied across beat subdivision (shown with respect to the equal subdivision) in the example with tres and guitar. Here we continue the analysis of the guitar and tres and calculate the asynchrony between them across the beat subdivisions. The summary of the differences in milliseconds is contained in the output. This output reports the asynchrony between the instruments across the subdivisions and the \emph{t-test} provides an index of whether the mean differences are significantly different from zero. In other words, the summary shows in which beat subdivisions the two musicians are performing asynchronously in a consistent manner.

```{r stats,results='asis'}
dn <- sync_execute_pairs(df = CSS_Song2,
  instruments = c("Guitar", "Tres"),
  beat = "SD")
print(knitr::kable(summarise_sync_by_pair(dn, bybeat = TRUE),
  digits = 2,caption = 'Asynchrony between tres and guitar 
  across beat sub-divisions.'))
```

The output of Table 4 suggests that about half of the asynchronies in the subdivisions are statistically significant (and this includes a correction for multiple comparisons using the Bonferroni adjustment). Asynchronies over 10 ms tend to be statistically significant and these fall into beat subdivisions 4, 8, 9, 12, 13, and 16. Here we are using t-test as an arbitrary indicator of the differences since the assumption for the independence of the observations is not met since the two instruments are temporally not independent.

Calculation of asynchronies between parts depends on the number of comparable onsets (points at which both are playing on the same beat subdivision). This will be relatively high in music with a homophonic texture, and low in an interlocking or hocketing pattern where the musicians are playing on alternating beats. In Cuban salsa and son, some instruments play on most subdivisions (as with the guitar and tres in this piece) while others are relatively sparse and may only coincide a few times per cycle. In this piece the clave plays on subdivisions 1, 4, 7, 11, 13 while the bass starts off playing on subdivisions 4, 5, 7, 13, 15; thus they coincide on subdivisions 4, 7 and 13 on most cycles and these subdivisions are the points at which we can calculate asynchronies. In sum, the number of joint onsets (onset occurring around the same beat) for each pair of instrument varies greatly. To keep the mean and standard deviations comparable, we can randomly sample an equal number of joint onsets for both instruments.

## 2.3 Synchrony across multiple performances

So far our analysis has considered synchrony within a single performance. We can also take several performances, choose the instrument pairings, and carry out the desired comparison, provided that this is conceptually meaningful and technically feasible (similar types of onset and annotation data are available). For this, we load five Cuban Son and Salsa performances that come with the library and run the same analysis as was done in prior examples using `sync_sample_paired` function across the performances.

```{r, fig.cap='Absolute asynchronies between guitar and tres in five performances within the CSS_IEMP dataset.'}
corpus <- onsetsync::CSS_IEMP
D <- sync_sample_paired(corpus,"Guitar","Tres", beat="SD")
D <- D$asynch
D$asynch_abs <- abs(D$asynch)*1000
print(plot_by_dataset(D,"asynch_abs","name", box = TRUE))
```

We can create a simple violin plot of the combined data ignoring the beat sub-divisions (Figure 4). All further analyses and summaries are then also possible.

Figure 4 suggests that the asynchrony of 13 ms between tres and guitar we saw in the analysis of \emph{Palo Alto} is largely in line with the overall asynchrony between the same instruments in the other instruments (overall mean asynchrony in all 5 pieces is 17 ms) although our example piece (\emph{Palo Santo}) shows the tightest synchrony in these examples.

## 2.4 Synchrony with other variables

It is relatively straightforward to explore whether synchrony is linked with another variable. For example, the precision of synchrony could change when the tempo of the performance is faster or slower. Or synchrony might be influenced by the dynamics or note density in a specific section. As an example, we illustrate synchrony across the time and tempo. As tempo is commonly defined by the beats per minute (BPM), we use the cycle information to convert the isochronous beat times to BPM.

```{r,fig.cap='Asynchrony between tres and guitar (Y-axis) across time (X-axis) and tempo (illustrated with marker colors).'}
CSS_Song2 <- CSS_Song2 %>%
  group_by(Cycle) %>%
  mutate(Tempo = 240/(max(Isochronous.SD.Time) - 
                        min(Isochronous.SD.Time))) %>%
  ungroup()

d2 <- sync_sample_paired(CSS_Song2,"Tres","Guitar", 
                         beat = "Tempo")
d3 <- sync_sample_paired(CSS_Song2,"Tres","Guitar", 
                         beat = "Isochronous.SD.Time")
tmp <- data.frame(asynch = d2$asynch*1000, 
                  Tempo = d2$beatL, 
                  Time = d3$beatL)
print(plot_by_var_time(df = tmp,
  var1 = "Time",
  var2 = "asynch",
  var3 = "Tempo",
  ylabel = "Asynchrony (ms)"))
```

As Figure 5 shows, synchrony fluctuates across time but these fluctuations do not seem to be directly related to the underlying tempo changes, although the last minute and half is performed slightly slower ($\approx71$ BPM) and the guitar is approximately 20 ms ahead of tres.

# 3 Analysis of periodicity

We now explore the periodic nature of the instrument onset times without reference to the metrical structure. Functions described in this part may in fact be particularly useful in other musical genres which do not exhibit a clear metrical structure; nonetheless they are illustrated here using the example used in the previous analyses.

Let's take the guitar onsets in the first example (\texttt{CSS\_Song2}) and define an extract of 5 seconds starting from 1 minute mark. We know from the previous analyses that guitar keeps a fairly steady beat, separated by about 0.25 seconds while leaving out some beats.

```{r}
extract <- dplyr::filter(CSS_Song2, Guitar >= 60 & Guitar < 65)
fig6A <- gaussify_onsets(extract$Guitar, wlen = 0.2, plot = TRUE)
```

For the periodicity analyses, we want to transform the discrete onsets to continuous time representation with an uniform sampling rate. This is done with the \texttt{gaussify\_onsets} function. This makes the comparisons of the models easier and allows us to handle time more explicitly. From this new continuous representation, we can estimate the periodicity of the guitar strokes by using different methods; Here we demonstrate (a) a simple time difference method, (b) method based on autocorrelation, and (c) fast fourier transform. We constrain the frequency a range of 0.05 to 0.50 seconds. After the extraction of periodicity with \texttt{periodicity} function, we will plot the estimated frequency curves (Figure 6).

```{r}
frq_range <- c(0.05,0.50)
fig6B <- periodicity(extract, instr = "Guitar", 
  freq_range=frq_range, method = "diff", title = "Difference")
fig6C <- periodicity(extract, instr="Guitar", 
  freq_range=frq_range, method = "acf",title = "Autocorrelation")
fig6D <- periodicity(extract, instr="Guitar", 
  freq_range=frq_range, method = "per",title = "FFT")
```

```{r periodicityfigure,echo=FALSE,fig.cap='Periodicity analysis of a 5-second extract of guitar playing in Cuban son. Panel A shows the onsets (red) and the conversion into gaussian distributions (blue) for a continuous representation. Panel B displays the difference in onset times as a primitive periodicity measure, Panel C shows the output from autocorrelation function, and Panel D shows the estimation of periodicity by FFT.'}

library(cowplot)
plots <- align_plots(fig6A$fig, align = 'v', axis = 'l')
bottom_row <- plot_grid(fig6B$Figure, fig6C$Figure, fig6D$Figure, 
  labels = c('B', 'C', 'D'), label_size = 12,nrow =1)
fig6 <- plot_grid(plots[[1]], bottom_row, labels = c('A', ''), 
  label_size = 12, ncol = 1)
print(fig6)
```

From the estimated periodicity curves, we can obtain the peak period in milliseconds, and if we so wish, we convert this into beats per minute (BPM). For the latter convention, we make use of the fact that the guitar plays four times per four beats.

```{r}
PM <- summarise_periodicity(fig6D$Curve)
print(paste('Period of',round(PM$Per*1000),'ms'))
print(paste(period_to_BPM(PM$Per)/4,'BPM'))
```

```{r fourperiodicies, echo=FALSE, fig.cap='Periodicity analysis of four selected instruments in the example song (Palo Alto). The grid lines indicate multiples of the period identified earlier as the underlying beat (217 ms).'}

excerpt <- dplyr::filter(CSS_Song2,Guitar>60 & Guitar < 60*2)
excerpt <- data.frame(excerpt)
p1 <- onsetsync::periodicity(excerpt,instr = "Guitar",
                             freq_range = c(0,1),sampling_rate = 100)
p2 <- onsetsync::periodicity(excerpt,instr = "Bass",
                             freq_range = c(0,1),sampling_rate = 100)
p3 <- onsetsync::periodicity(excerpt,instr = "Tres",
                             freq_range = c(0,1),sampling_rate = 100)
p4 <- onsetsync::periodicity(excerpt,instr = "Clave",
                             freq_range = c(0,1),sampling_rate = 100)
p1$Curve$Instrument <- "Guitar"
p2$Curve$Instrument <- "Bass"
p3$Curve$Instrument <- "Tres"
p4$Curve$Instrument <- "Clave"
df <- rbind(p1$Curve,p2$Curve,p3$Curve,p4$Curve)
df$Instrument<-factor(df$Instrument)

fig6 <- ggplot(df,aes(x=Time,y=Ampl))+
  geom_line(color='blue')+
  facet_wrap(.~Instrument,scales = "free",nrow = 2)+ # ,
  theme_linedraw()+
  scale_y_continuous(expand = c(0.01,0.01))+
  scale_x_continuous(breaks = seq(0,1,by=0.217/2),
                     labels = scales::number_format(accuracy = 0.01),expand = c(0.01,0.01))+
  xlab('Period (s)')+
  ylab('Amplitude')
print(fig6)

```

Being able to bring periodicity estimation of onset sequences into the analysis of asynchrony has many useful purposes. It can serve as a diagnostic into the periodic patterns performed by the instruments. This might be particularly useful if no annotation about the cycles and beat subdivisions is available and the analyst is trying to determine what are the periodicities exhibited by the different instruments. For instance, Figure 7 shows the periodicities of the selected four instruments in our salsa example. The analysis reveals that guitar and tres tend to play the fastest beat pulsation (0.22 s) while tres occasionally ventures into double (0.44 s) and four times the period (0.88 s). Bass shows three dominant periods, one at the fastest beat pulsation level (0.22), double of that period (0.44) and two uneven subdivisions of the underlying beat (0.35 s and 0.58 s). The clave's pattern includes a period of 0.66 secs, three times the fastest pulse, presumably thanks to its pattern at the start of the cycle (when it plays on subdivisions 1, 4 & 7).

# 4 Conclusions

`onsetsync` provides representations, visualisations and calculations of key measures relating to onset structures in music. The library does not take a stance on how the onset times have been estimated, but assumes that onset times are mapped onto beat positions, and that metre annotations are available. The strengths of the library are in being able to read onset data and annotations directly from OSF and to define and redefine the object of synchrony analysis (e.g., between all instruments or specific instruments, or annotated beat sub-division and specific instruments, or the mean onsets of the instruments and specific instruments) and help the user to calculate classic measures of synchrony between instruments. The visualisations are designed to facilitate the selection and exploration of the data that would be otherwise challenging. Some analysis options are provided to calculate the periodicity of the onsets. Future needs for development lie in creating a specific corpus format that would be open to many uses, to test the framework with other datasets, and to incorporate other analytical options such as circular statistics, granger causality or cross-recurrence analyses.

We hope scholars and students of music, music information retrieval and psychology can pursue topics related to synchrony in music with a transparent workflow assisted by the library and capitalise on existing datasets that are compatible with the library. Eventually we hope that tools such as these will allow a larger number of people to create and utilise open datasets related to musical behaviours, helping the topic become data-driven, empirical and collaborative.

# Acknowledgements

We acknowledge contributions from IEMP Fellows, specifically Adrian Poole, who created Cuban Son and Salsa dataset with Simone Tarsitani and Martin Clayton, but also Rainer Polak, [Richard Jankowsky](https://as.tufts.edu/music/people/faculty/richard-jankowsky), [Nori Jacoby](http://www.norijacoby.com), Mark Doffman, Luis Jure, Andy McGuiness, Nikki Moran, and [Martín Rocamora](https://iie.fing.edu.uy/~rocamora/)). We are thankful for technical support by [Simone Tarsitani](https://www.durham.ac.uk/staff/simone-tarsitani/) and useful feedback by Juliano Abramovay. For funding and time, we acknowledge EU Horizon 2020 FET project [EnTimeMent - ENtrainment & synchronization at multiple TIME scales in the MENTal foundations of expressive gesture](https://entimement.dibris.unige.it/) that was instrumental in turning the idea about the shared resource for temporal analyses such as musical onsets into an open access library.

The package is available from \url{https://tuomaseerola.github.io/onsetsync/}. The analyses shown in the article are contained in the vignette and article titled "Analysis Example".

# References

### 2023-01-17 paper/paper.Rmd
Renamed figures

@@
 
-```{r synch2isochron,fig.width=7.0, fig.height=5.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
+```{r figure1,fig.width=7.0, fig.height=5.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
 fig1 <- plot_by_beat(df = CSS_Song2, 
@@
 
-```{r, fig.cap='Relative timing deviations from beat sub-divisions across the performance for guitar and tres.'}
+```{r figure2, fig.cap='Relative timing deviations from beat sub-divisions across the performance for guitar and tres.'}
 print(plot_by_beat(df = CSS_Song2,
@@
 
-```{r,fig.cap='Asynchronies across multiple instrument pairs.'}
+```{r figure3,fig.cap='Asynchronies across multiple instrument pairs.'}
 inst <- c("Clave","Bass","Guitar","Tres")
@@
 
-```{r, fig.cap='Absolute asynchronies between guitar and tres in five performances within the CSS_IEMP dataset.'}
+```{r figure4, fig.cap='Absolute asynchronies between guitar and tres in five performances within the CSS_IEMP dataset.'}
 corpus <- onsetsync::CSS_IEMP
@@
 
-```{r,fig.cap='Asynchrony between tres and guitar (Y-axis) across time (X-axis) and tempo (illustrated with marker colors).'}
+```{r figure5, fig.cap='Asynchrony between tres and guitar (Y-axis) across time (X-axis) and tempo (illustrated with marker colors).'}
 CSS_Song2 <- CSS_Song2 %>%
@@
 
-```{r periodicityfigure,echo=FALSE,fig.cap='Periodicity analysis of a 5-second extract of guitar playing in Cuban son.
+```{r figure6, echo=FALSE, fig.cap='Periodicity analysis of a 5-second extract of guitar playing in Cuban son.
 Panel A shows the onsets (red) and the conversion into gaussian distributions (blue) for a continuous representation.
@@
 
-```{r fourperiodicies, echo=FALSE, fig.cap='Periodicity analysis of four selected instruments in the example song (Palo Alto).
+```{r figure7, echo=FALSE, fig.cap='Periodicity analysis of four selected instruments in the example song (Palo Alto).
 The grid lines indicate multiples of the period identified earlier as the underlying beat (217 ms).'}

### 2023-05-13 paper/paper.Rmd
Update paper.Rmd
Fixed a typo

@@
 through note onsets.
-Data is ingested in the form of CSV files with individual instrument onset times are aligned with labelled beat positions; separate metre annotations comprising lists of time estimates for the downbeats are also used for some functions.
+Data is ingested in the form of CSV files with individual instrument onset times aligned with labelled beat positions; separate metre annotations comprising lists of time estimates for the downbeats are also used for some functions.
 The package includes functions for common operations such as adding isochronous beats based on metrical structure, adding annotations, calculating classic measures of synchrony between two performers, assessing the periodicity of the onsets, and visualising synchrony across metrical cycles, time, or another property.

### 2023-05-13 paper/paper.Rmd
Fix typos and wrong citations types

@@
 The examples are taken from the Interpersonal Entrainment in Music Performance (IEMP) project collection that has compiled small ensemble performance data from six traditions (North Indian ragas, Malian jembe, Tunisian stambeli, Uruguayan candombe, European string quartet, and Cuban son and salsa).
-These collection contain tens of hours annotated and verified onsets and are available from Open Science Framework (OSF).
+These collections contain tens of hours annotated and verified onsets and are available from Open Science Framework (OSF).
 There are also noteworthy datasets containing audio and annotated contents of music relevant to assessing synchrony such as _Carnatic Music Rhythm Dataset_ [@srinivasamurthy2015particle], _Hindustani Music Rhythm Dataset_ [@srinivasamurthy2016generalized], _Arab-Andalusian (Maghreb) dataset_ [@sordo2014creating], _Scandinavian Fiddle Tunes with Accompanying Foot-Tapping_ [@lordelo2020adversarial] and _MAESTRO (MIDI and Audio Edited for Synchronous Tracks and Organisation)_ dataset which is an extensive collection (200+ hours) of professional piano performances [@hawthorne2018enabling].
-Some datasets such as Chopin performances compiled by [@goebl2010investigations] facilitate study of the synchrony between the two hands of the pianist.
+Some datasets such as Chopin performances compiled by @goebl2010investigations facilitate study of the synchrony between the two hands of the pianist.
 
@@
 _sensorimotor synchronization_ (SMS) or just synchronization, a
-processes occurring within 100-2000 ms timescales, which has commonly
+process occurring within 100-2000 ms timescales, which has commonly
 been studied with tapping experiments [@repp2008sensorimotor, @repp2013sensorimotor] but also through analyses of music corpora, from string quartets [@wing2014optimal] to drum ensembles [@polak2016both] and to

### 2023-05-28 paper/paper.Rmd
Improvements based on reviews
R1

@@
 
-# 1 Introduction
+# Summary
 
@@
 between performers. An important aspect of this interpersonal
-entrainment is the synchronisation of musical events between performers,
+entrainment is the synchronization of musical events between performers,
 and this can be analysed from the note onsets obtained from recorded
 performances. Analysis of the asynchronies between onsets gives an
-insight into synchronisation between performers, and the extent to which
+insight into synchronization between performers, and the extent to which
 it is accurate (how close together the events are in time) and precise
@@
 
-## 1.1 Statement of need
+## Statement of need
 
@@
 
-## 1.2 Onset datasets
-
 `onsetsync` package comes with example performances from two traditions, Cuban Son and Salsa performances [@Poole_Tarsitani_Clayton_2022] and North Hindustani performances [@Clayton_Leante_Tarsitani_2021].
@@
 
-## 1.3 Terminology and measures of synchrony
+### Terminology and measures of synchrony
 
@@
 process occurring within 100-2000 ms timescales, which has commonly
-been studied with tapping experiments [@repp2008sensorimotor, @repp2013sensorimotor] but also through analyses of music corpora, from string quartets [@wing2014optimal] to drum ensembles [@polak2016both] and to
-Indian instrumental music performances [@clayton_et_al_2018b]. Synchronisation analyses can also be applied to field observations
-such as the study of how walking drumming groups attempt to
-resist synchronisation to the beats of other drum groups during their
-walk across the city during a festival [@lucas2011inter].
+been studied with tapping experiments [@repp2008sensorimotor; @repp2013sensorimotor] but also through analyses of music corpora, from string quartets [@wing2014optimal] to drum ensembles [@polak2016both], and to
+Indian instrumental music performances [@clayton_et_al_2018b].
 
 A number of different measures of synchrony have been proposed. We
-follow the definitions from [@clayton2020, p. 161] who adopt and
-develop the measures defined by [@rasch1988timing, @rasch1979synchronization]:
+follow the definitions from @clayton2020 who adopt and
+develop the measures defined by Rasch [-@rasch1988timing; -@rasch1979synchronization]:
 
-Here we first define the onset time differences, $d$, where $I_{1, i}$ and $I_{2, i}$ refer to an onset at time point $i$ for instrument 1 and 2:
+Here we first define the onset time differences (or asynchronies), $d$, where $I_{1, i}$ and $I_{2, i}$ refer to an onset at time point $i$ for instrument 1 and 2:
 
@@
 \begin{equation}
-Sd = \sqrt\frac{\sum{(d_i-\bar{d})^2}}{n-1}
+Sd = \sqrt\frac{\sum_{i=1}{(d_i-\bar{d})^2}}{n-1}
 \end{equation}
@@
 \begin{equation}
-RMS = \sqrt{\Sigma_{i=1}^{T}{\Big(\frac{d_{1,i} -d_{2,i}}{T}\Big)^2}}
+RMS = \sqrt{\Sigma_{i=1}^{n}{\Big(\frac{d_{i}}{n}\Big)^2}}
 \end{equation}
+
+where $n$ is the number of onset time differences.
 
@@
 \begin{equation}
-\hat{|s|} = \frac{1}{n} \sum_{i=i}^{n} |d_{i}|
+\bar{|s|} = \frac{1}{n} \sum_{i=1}^{n} |d_{i}|
 \end{equation}
@@
 *Mean pairwise asynchrony*: mean difference in the onset values
-  of two instruments, i.e., the signed asynchrony values, as in
+  of two instruments, $\bar{d}$, i.e., the signed asynchrony values.
+
+*Mean relative asynchrony*: mean position of an instrument's
+  onsets relative to average position ($\bar{I}$) of the group, where the average position is
 
 \begin{equation}
-\hat{s_p} = \frac{1}{n} \sum_{i=i}^{n} d_{i}
+\bar{I} = \frac{I_{1} + I_{2} + \ldots + I_{n}}{n}
 \end{equation}
 
-*Mean relative asynchrony*: mean position of an instrument's
-  onsets relative to average position ($\hat{w}$) of the group, where the average position is
+and the relative onset time $R$ is
 
 \begin{equation}
-\hat{w} = \frac{w_{1} + w_{2} + \ldots + w_{n}}{n}
-\end{equation}
-
-and the relative onset time $w$ is
-
-\begin{equation}
-v_n = w_n - \hat{w}
+R_i = I_i - \bar{I}
 \end{equation}
@@
 \begin{equation}
-\hat{s_r} = \frac{1}{n} \sum_{i=i}^{n} v_{i}
+\bar{s} = \frac{1}{n} \sum_{i=1}^{n} R_{i}
 \end{equation}
@@
 These methods are based on asynchronies between onsets occurring at the same beat positions; other methods exist for assessing synchrony such as utilising circular statistics [@berens2009circstat], recurrence quantification analysis [@wallot2018analyzing], or calculating the synchrony of spike trains [@kreuz2007measuring].
-`onsetsync` is focussed on measures of asynchrony relying on onsets between instruments at the same beat positions, which is usually the relevant approach in music sense.
 
-## 1.4 Availability
+## Availability and functionality
 
@@
 library(onsetsync)
-print(packageVersion('onsetsync'))
 ```
 
-In order to be able to build the carry out such comparisons, `onsetsync` has several auxiliary functions.
+To carry out such comparisons, `onsetsync` has several auxiliary functions.
 There are functions that help assessing relative timings, to associate onsets with beat subdivisions and cycles in music structure, and to select and compare the instruments.
-The library has seven categories of functions; (1) input/output, (2)
-annotation, (3) visualisation, (4) synchrony, (5) periodicity, (6)
-summary, and (7) dataset functions (see Table 1). The functions are named to be easily associated with the categories.
+The library has seven categories of functions summarised in Table 1.
 
@@
 It is also worth pointing out that the library is not dedicated to extraction of onsets from audio as that can be done in other packages (e.g.
-[Librosa](https://librosa.org), [Essentia](https://essentia.upf.edu), or [MIR toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms.
+[Librosa](https://librosa.org), [Essentia](https://essentia.upf.edu), [MIR toolbox for Matlab](https://www.jyu.fi/hytk/fi/laitokset/mutku/en/research/materials/mirtoolbox), or [Sonic Visualiser](https://www.sonicvisualiser.org) using well-known onset detection algorithms).
 Here we take it as granted that the onsets have been extracted already, someone has carried out a quality check, assigned relevant onsets to beat positions and saved them in csv files.
-For solid overviews on computational onset extraction from audio, see [@schluter2014improved] and [@bock2016madmom], and for a full workflow of how to combine onset detection and annotation of the musical information, see [@Clayton2020emr].
+For solid overviews on computational onset extraction from audio, see @schluter2014improved or @bock2016madmom, and for a full workflow of how to combine onset detection and annotation of the musical information, see @Clayton2020emr.
 
-## 1.5 Onset representation
+### Onset representation
 
@@
 The library comes with a set of five Cuban salsa and son (CSS) performances and this is the second example from the IEMP `CSS_IEMP` mini-corpus.
-The second line of the code selects the nine columns of interest for the analysis; three first are meta-data, followed by four instruments (selected to aid legibility of the tables and charts), and two timing related columns are included at the end.
+The second line of the code selects the nine columns of interest for the analysis; the first two columns are meta-data, followed by four instruments, and three timing related columns are included at the end.
 
@@
 The onset times for the instruments are available in seconds (with precision of 6 digits, although we prefer to use milliseconds (ms) in the analyses of synchrony to avoid reporting
-unnecessarily many digits.
+unnecessarily many digits).
 The last three columns refer to information about the relative time: `SD` represents the beat subdivision within the cycle.
@@
 Information about cycles and beat subdivisions is generally based on manual annotation, and originates separately from the automatically extracted onsets; the two are combined to produce tables with onsets assigned to beat positions.
+This explains why the first row of the table has no onsets but it contains annotation information (Cycle 1 and beat sub-division 1).
 It is also possible to read the csv files directly from OSF using `get_OSF_csv` function to allow transparent and reproducible analysis workflows with published datasets.
 
-## 1.6 Onset summaries
+### Onset summaries
 
-Let's explore the overall structure of the onsets in the example piece.
-
-```{r summarise,message=FALSE,warning=FALSE,results='asis'}
-tab1 <- summarise_onsets(df = CSS_Song2, 
-  instr = c('Clave','Bass','Guitar','Tres','Isochronous.SD.Time'))
-print(knitr::kable(tab1,digits = 1,
-  caption = 'Summary of onsets in the example data 
-  (Md, M, Min, and Max in ms).'))
-```
-
-Table 3 shows the summary of the onsets, namely the number of onsets for each instrument, the descriptive statistics referring to time difference of successive onsets ($d$) within each instrument.
-We have provided the isochronous beat subdivisions as a reference, which shows how many beats there are and what the is the average duration of the beats (220 ms) in this performance lasting approximately 5 minutes and 45 seconds.
-
-As a broad overview, we can use the function `plot_by_beat` visualise the asynchrony relative to an equal division subdivision of the beats for each instrument across the time.
+Let's explore the overall structure of the onsets in the example piece using the function `plot_by_beat` to visualise the asynchrony relative to an equal division subdivision of the beats for each instrument across the time.
 This visualisation gives a summary about when the instruments are playing relative to the beat division and the sections of the song, shown in Figure 1.
-For example, in many genres different instruments contribute to specific subdivision of the beat.
-This is of course the case on Cuban Son, but maybe even more extreme in Malian Jembe music [@polak2010rhythmic] or more variable in Western jazz [@wesolowski2016timing].
 
 ```{r figure1,fig.width=7.0, fig.height=5.0,fig.cap='Onsets arranged for beat sub-divisions for four instruments across the whole piece.'}
+library(dplyr)
 fig1 <- plot_by_beat(df = CSS_Song2, 
@@
 Notice how the bass plays mainly on subdivisions 5, 7, 13, 15, except in the last minute when the pattern changes into playing subdivisions 3, 4, 7, 8, 15, 16.
-The guitar plays almost on every beat subdivision, and towards the end of the piece tres --  a cuban guitar with 6-strings tuned to 3 pitches -- plays a pattern of 3 subdivisions and a break.
-Overall, this six-minute piece does not contain a high amount of variation in the patterns of onsets across the subdivisions.
-We can see that the bass elaborates the beats during the last minute or so while the clave, guitar and tres keep playing their patterns throughout the piece.
+The guitar plays almost on every beat subdivision, and towards the end of the piece the tres --  a cuban guitar with 6-strings tuned to 3 pitches -- plays a pattern of 3 subdivisions and a break.
 
-# 2 Analysis of synchrony
+# Analysis of synchrony
 
@@
 
-Overall this suggests that these two instruments tend to play earlier when compared to the isochronous beat division of the cycle, although guitar is tightly aligned (<2%) with the reference beat on every fourth beat subdivisions (subdivisions of 1, 5, 9, and 13), which are also the salient metrical positions.
-Similar fluctuations across the beat subdivisions are not evident in tres, but the graph reveals that the tres tends to play a little earlier than guitar.
+Overall this suggests that these two instruments tend to play earlier when compared to the isochronous beat division of the cycle, although the guitar is tightly aligned (<2%) with the reference beat on every fourth beat subdivisions (subdivisions of 1, 5, 9, and 13), which are also the salient metrical positions.
+Similar fluctuations across the beat subdivisions are not evident in the tres, but the graph reveals that the tres tends to play a little earlier than guitar.
 
-## 2.1 Synchrony between the instruments
+## Synchrony between the instruments
 
@@
 
-In Figure 1 we see how different pairs of instruments have different asynchrony relationships; the bass is consistently ahead of the guitar, and clave and guitar play behind tres, to pick some of the extreme examples from the visualisation.
+In Figure 3 we see how different pairs of instruments have different asynchrony relationships; the bass is consistently ahead of the guitar, and clave and guitar play behind tres, to pick some of the extreme examples from the visualisation.
 The vertical lines stand for median asynchrony and the violin plot shows the distribution of onset time differences.
@@
 The output suggests that at least in this piece, these two instruments tend to have the synchronization precision around 20 ms, depending on the method of calculation.
-Pairwise asynchronisation is the standard deviation of the signed asynchrony; this tends to be highly correlated with the mean absolute asynchrony.
+Pairwise asynchronization is the standard deviation of the signed asynchrony; this tends to be highly correlated with the mean absolute asynchrony.
 The Mean pairwise asynchrony measure shows that on average, the clave plays 16 ms after the bass.
@@
 
-The relative measure (mean relative asynchrony) shows that bass is ahead (ie.
+The relative measure (mean relative asynchrony) shows that bass is ahead (i.e.
 -19 ms) of the mean of other instruments whereas clave is almost exactly in synchrony with the other instruments (<1 ms late).
@@
 
-## 2.2 Synchrony between the instruments across beat subdivisions
+It is also possible to [calculate synchrony between instruments across beat sub-divisions](https://tuomaseerola.github.io/onsetsync/articles/advanced_topics.html), [assess synchrony across multiple performances](https://tuomaseerola.github.io/onsetsync/articles/advanced_topics.html), and [estimate and visualise synchrony with other variables](https://tuomaseerola.github.io/onsetsync/articles/advanced_topics.html).
+The library also comes with functionality to [analyse the periodicity of the onsets](https://tuomaseerola.github.io/onsetsync/articles/analysisis_of_periodicity.html) and [synthesising the onsets](https://tuomaseerola.github.io/onsetsync/articles/synthesise_onsets.html).
 
-In the first visualisation we already observed that asynchronies varied across beat subdivision (shown with respect to the equal subdivision) in the example with tres and guitar.
-Here we continue the analysis of the guitar and tres and calculate the asynchrony between them across the beat subdivisions.
-The summary of the differences in milliseconds is contained in the output.
-This output reports the asynchrony between the instruments across the subdivisions and the \emph{t-test} provides an index of whether the mean differences are significantly different from zero.
-In other words, the summary shows in which beat subdivisions the two musicians are performing asynchronously in a consistent manner.
-
-```{r stats,results='asis'}
-dn <- sync_execute_pairs(df = CSS_Song2,
-  instruments = c("Guitar", "Tres"),
-  beat = "SD")
-print(knitr::kable(summarise_sync_by_pair(dn, bybeat = TRUE),
-  digits = 2,caption = 'Asynchrony between tres and guitar 
-  across beat sub-divisions.'))
-```
-
-The output of Table 4 suggests that about half of the asynchronies in the subdivisions are statistically significant (and this includes a correction for multiple comparisons using the Bonferroni adjustment).
-Asynchronies over 10 ms tend to be statistically significant and these fall into beat subdivisions 4, 8, 9, 12, 13, and 16.
-Here we are using t-test as an arbitrary indicator of the differences since the assumption for the independence of the observations is not met since the two instruments are temporally not independent.
-
-Calculation of asynchronies between parts depends on the number of comparable onsets (points at which both are playing on the same beat subdivision).
-This will be relatively high in music with a homophonic texture, and low in an interlocking or hocketing pattern where the musicians are playing on alternating beats.
-In Cuban salsa and son, some instruments play on most subdivisions (as with the guitar and tres in this piece) while others are relatively sparse and may only coincide a few times per cycle.
-In this piece the clave plays on subdivisions 1, 4, 7, 11, 13 while the bass starts off playing on subdivisions 4, 5, 7, 13, 15; thus they coincide on subdivisions 4, 7 and 13 on most cycles and these subdivisions are the points at which we can calculate asynchronies.
-In sum, the number of joint onsets (onset occurring around the same beat) for each pair of instrument varies greatly.
-To keep the mean and standard deviations comparable, we can randomly sample an equal number of joint onsets for both instruments.
-
-## 2.3 Synchrony across multiple performances
-
-So far our analysis has considered synchrony within a single performance.
-We can also take several performances, choose the instrument pairings, and carry out the desired comparison, provided that this is conceptually meaningful and technically feasible (similar types of onset and annotation data are available).
-For this, we load five Cuban Son and Salsa performances that come with the library and run the same analysis as was done in prior examples using `sync_sample_paired` function across the performances.
-
-```{r figure4, fig.cap='Absolute asynchronies between guitar and tres in five performances within the CSS_IEMP dataset.'}
-corpus <- onsetsync::CSS_IEMP
-D <- sync_sample_paired(corpus,"Guitar","Tres", beat="SD")
-D <- D$asynch
-D$asynch_abs <- abs(D$asynch)*1000
-print(plot_by_dataset(D,"asynch_abs","name", box = TRUE))
-```
-
-We can create a simple violin plot of the combined data ignoring the beat sub-divisions (Figure 4). All further analyses and summaries are then also possible.
-
-Figure 4 suggests that the asynchrony of 13 ms between tres and guitar we saw in the analysis of \emph{Palo Alto} is largely in line with the overall asynchrony between the same instruments in the other instruments (overall mean asynchrony in all 5 pieces is 17 ms) although our example piece (\emph{Palo Santo}) shows the tightest synchrony in these examples.
-
-## 2.4 Synchrony with other variables
-
-It is relatively straightforward to explore whether synchrony is linked with another variable.
-For example, the precision of synchrony could change when the tempo of the performance is faster or slower.
-Or synchrony might be influenced by the dynamics or note density in a specific section.
-As an example, we illustrate synchrony across the time and tempo.
-As tempo is commonly defined by the beats per minute (BPM), we use the cycle information to convert the isochronous beat times to BPM.
-
-```{r figure5, fig.cap='Asynchrony between tres and guitar (Y-axis) across time (X-axis) and tempo (illustrated with marker colors).'}
-CSS_Song2 <- CSS_Song2 %>%
-  group_by(Cycle) %>%
-  mutate(Tempo = 240/(max(Isochronous.SD.Time) - 
-                        min(Isochronous.SD.Time))) %>%
-  ungroup()
-
-d2 <- sync_sample_paired(CSS_Song2,"Tres","Guitar", 
-                         beat = "Tempo")
-d3 <- sync_sample_paired(CSS_Song2,"Tres","Guitar", 
-                         beat = "Isochronous.SD.Time")
-tmp <- data.frame(asynch = d2$asynch*1000, 
-                  Tempo = d2$beatL, 
-                  Time = d3$beatL)
-print(plot_by_var_time(df = tmp,
-  var1 = "Time",
-  var2 = "asynch",
-  var3 = "Tempo",
-  ylabel = "Asynchrony (ms)"))
-```
-
-As Figure 5 shows, synchrony fluctuates across time but these fluctuations do not seem to be directly related to the underlying tempo changes, although the last minute and half is performed slightly slower ($\approx71$ BPM) and the guitar is approximately 20 ms ahead of tres.
-
-# 3 Analysis of periodicity
-
-We now explore the periodic nature of the instrument onset times without reference to the metrical structure.
-Functions described in this part may in fact be particularly useful in other musical genres which do not exhibit a clear metrical structure; nonetheless they are illustrated here using the example used in the previous analyses.
-
-Let's take the guitar onsets in the first example (\texttt{CSS\_Song2}) and define an extract of 5 seconds starting from 1 minute mark.
-We know from the previous analyses that guitar keeps a fairly steady beat, separated by about 0.25 seconds while leaving out some beats.
-
-```{r}
-extract <- dplyr::filter(CSS_Song2, Guitar >= 60 & Guitar < 65)
-fig6A <- gaussify_onsets(extract$Guitar, wlen = 0.2, plot = TRUE)
-```
-
-For the periodicity analyses, we want to transform the discrete onsets to continuous time representation with an uniform sampling rate.
-This is done with the \texttt{gaussify\_onsets} function.
-This makes the comparisons of the models easier and allows us to handle time more explicitly.
-From this new continuous representation, we can estimate the periodicity of the guitar strokes by using different methods; Here we demonstrate (a) a simple time difference method, (b) method based on autocorrelation, and (c) fast fourier transform.
-We constrain the frequency a range of 0.05 to 0.50 seconds.
-After the extraction of periodicity with \texttt{periodicity} function, we will plot the estimated frequency curves (Figure 6).
-
-```{r}
-frq_range <- c(0.05,0.50)
-fig6B <- periodicity(extract, instr = "Guitar", 
-  freq_range=frq_range, method = "diff", title = "Difference")
-fig6C <- periodicity(extract, instr="Guitar", 
-  freq_range=frq_range, method = "acf",title = "Autocorrelation")
-fig6D <- periodicity(extract, instr="Guitar", 
-  freq_range=frq_range, method = "per",title = "FFT")
-```
-
-```{r figure6, echo=FALSE, fig.cap='Periodicity analysis of a 5-second extract of guitar playing in Cuban son.
-Panel A shows the onsets (red) and the conversion into gaussian distributions (blue) for a continuous representation.
-Panel B displays the difference in onset times as a primitive periodicity measure, Panel C shows the output from autocorrelation function, and Panel D shows the estimation of periodicity by FFT.'}
-
-library(cowplot)
-plots <- align_plots(fig6A$fig, align = 'v', axis = 'l')
-bottom_row <- plot_grid(fig6B$Figure, fig6C$Figure, fig6D$Figure, 
-  labels = c('B', 'C', 'D'), label_size = 12,nrow =1)
-fig6 <- plot_grid(plots[[1]], bottom_row, labels = c('A', ''), 
-  label_size = 12, ncol = 1)
-print(fig6)
-```
-
-From the estimated periodicity curves, we can obtain the peak period in milliseconds, and if we so wish, we convert this into beats per minute (BPM).
-For the latter convention, we make use of the fact that the guitar plays four times per four beats.
-
-```{r}
-PM <- summarise_periodicity(fig6D$Curve)
-print(paste('Period of',round(PM$Per*1000),'ms'))
-print(paste(period_to_BPM(PM$Per)/4,'BPM'))
-```
-
-```{r figure7, echo=FALSE, fig.cap='Periodicity analysis of four selected instruments in the example song (Palo Alto).
-The grid lines indicate multiples of the period identified earlier as the underlying beat (217 ms).'}
-
-excerpt <- dplyr::filter(CSS_Song2,Guitar>60 & Guitar < 60*2)
-excerpt <- data.frame(excerpt)
-p1 <- onsetsync::periodicity(excerpt,instr = "Guitar",
-                             freq_range = c(0,1),sampling_rate = 100)
-p2 <- onsetsync::periodicity(excerpt,instr = "Bass",
-                             freq_range = c(0,1),sampling_rate = 100)
-p3 <- onsetsync::periodicity(excerpt,instr = "Tres",
-                             freq_range = c(0,1),sampling_rate = 100)
-p4 <- onsetsync::periodicity(excerpt,instr = "Clave",
-                             freq_range = c(0,1),sampling_rate = 100)
-p1$Curve$Instrument <- "Guitar"
-p2$Curve$Instrument <- "Bass"
-p3$Curve$Instrument <- "Tres"
-p4$Curve$Instrument <- "Clave"
-df <- rbind(p1$Curve,p2$Curve,p3$Curve,p4$Curve)
-df$Instrument<-factor(df$Instrument)
-
-fig6 <- ggplot(df,aes(x=Time,y=Ampl))+
-  geom_line(color='blue')+
-  facet_wrap(.~Instrument,scales = "free",nrow = 2)+ # ,
-  theme_linedraw()+
-  scale_y_continuous(expand = c(0.01,0.01))+
-  scale_x_continuous(breaks = seq(0,1,by=0.217/2),
-                     labels = scales::number_format(accuracy = 0.01),expand = c(0.01,0.01))+
-  xlab('Period (s)')+
-  ylab('Amplitude')
-print(fig6)
-
-```
-
-Being able to bring periodicity estimation of onset sequences into the analysis of asynchrony has many useful purposes.
-It can serve as a diagnostic into the periodic patterns performed by the instruments.
-This might be particularly useful if no annotation about the cycles and beat subdivisions is available and the analyst is trying to determine what are the periodicities exhibited by the different instruments.
-For instance, Figure 7 shows the periodicities of the selected four instruments in our salsa example.
-The analysis reveals that guitar and tres tend to play the fastest beat pulsation (0.22 s) while tres occasionally ventures into double (0.44 s) and four times the period (0.88 s).
-Bass shows three dominant periods, one at the fastest beat pulsation level (0.22), double of that period (0.44) and two uneven subdivisions of the underlying beat (0.35 s and 0.58 s).
-The clave's pattern includes a period of 0.66 secs, three times the fastest pulse, presumably thanks to its pattern at the start of the cycle (when it plays on subdivisions 1, 4 & 7).
-
-# 4 Conclusions
+# Conclusions
 
@@
 The library does not take a stance on how the onset times have been estimated, but assumes that onset times are mapped onto beat positions, and that metre annotations are available.
-The strengths of the library are in being able to read onset data and annotations directly from OSF and to define and redefine the object of synchrony analysis (e.g., between all instruments or specific instruments, or annotated beat sub-division and specific instruments, or the mean onsets of the instruments and specific instruments) and help the user to calculate classic measures of synchrony between instruments.
-The visualisations are designed to facilitate the selection and exploration of the data that would be otherwise challenging.
-Some analysis options are provided to calculate the periodicity of the onsets.
-Future needs for development lie in creating a specific corpus format that would be open to many uses, to test the framework with other datasets, and to incorporate other analytical options such as circular statistics, granger causality or cross-recurrence analyses.
+Future needs for development lie in creating a specific corpus format that would be open to many uses, to test the framework with other datasets, and to incorporate other analytical options such as circular statistics, Granger causality or cross-recurrence analyses.
 

### 2023-05-28 tests/testthat/test-02-asynchrony.R

@@ -1,0 +1,17 @@
+# test-02-asynchrony.R
+
+d1 <- sync_sample_paired(df = CSS_Song2, 
+                         instr1 = 'Clave',
+                         instr2 = 'Bass',
+                         n = 0, 
+                         beat = 'SD')
+
+test_that("Demo data the correct length", {
+  expect_equal(length(d1$asynch),241)
+})
+
+test_that("Demo data has the correct mean value", {
+  expect_equal(mean(d1$asynch),0.0162328797)
+})
+
+

### 2023-05-28 tests/testthat/test-03-figures.R

@@ -1,0 +1,20 @@
+# test-03-figures.R
+
+fig1 <- plot_by_beat(df = CSS_Song2, 
+                     instr = c('Bass','Clave','Guitar','Tres'), 
+                     beat = 'SD', 
+                     virtual='Isochronous.SD.Time',
+                     pcols=2)
+
+test_that("Plot is created", {
+  expect_s3_class(fig1,"ggplot")
+})
+
+
+inst <- c('Clave','Bass','Guitar','Tres') # Define instruments 
+dn <- sync_execute_pairs(CSS_Song2,inst,beat = "SD")
+fig2 <- plot_by_pair(dn)
+
+test_that("Plot is created", {
+  expect_s3_class(fig2,"ggplot")
+})

### 2023-05-28 tests/testthat/test-04-periodicity.R

@@ -1,0 +1,32 @@
+# test-04-periodicity.R
+
+extract <- dplyr::filter(CSS_Song2, Guitar >= 60 & Guitar < 65)
+extract <- ungroup(extract) # drop grouping structure
+frq_range <- c(0.10,0.50)
+
+per1 <- periodicity(extract,instr='Guitar',freq_range=frq_range,
+                    method = 'diff', title = 'Difference')
+PM1 <- summarise_periodicity(per1$Curve)
+period1 <- round(PM1$Per*1000)
+
+test_that("Periodicity data has the correct value", {
+  expect_equal(period1,195)
+})
+
+per2 <- periodicity(extract,instr='Guitar',freq_range=frq_range,
+                    method = 'acf')
+PM2 <- summarise_periodicity(per2$Curve)
+period2 <- round(PM2$Per*1000)
+
+test_that("Periodicity data has the correct value", {
+  expect_equal(period2,204)
+})
+
+per3 <- periodicity(extract,instr='Guitar',freq_range=frq_range,
+                    method = 'fft')
+PM3 <- summarise_periodicity(per3$Curve)
+period3 <- round(PM3$Per*1000)
+
+test_that("Periodicity data has the correct value", {
+  expect_equal(period3,217)
+})

### 2023-05-29 R/plot_by_dataset.R
Additional articles

@@ -24,25 +24,11 @@ plot_by_dataset <-
            colour = 'lightblue',
            colorpalette = 'Set2') {
     # T. Eerola, Durham University, IEMP project
-    # 23/1/2018
-    # needs work
-    
-    # for testing
-    # df<-asere
-    # instr<-c('Bass','Tres','Bell')
-    # beat<-'SD'
-    # virtual<-'Virtual.SD'
-    # pcols<-2
-    # box<-TRUE
-    # griddeviations<-FALSE
-    #print(data)
-    #print(asynchronies)
     asynch <- n <- NULL
     
     DF <- dplyr::select(df, asynchronies, data)
     colnames(DF) <- c('asynch', 'data')
-#    print(head(DF))
-    
+
     g1 <-
       ggplot2::ggplot(DF, ggplot2::aes(asynch, fill = data), colour = 'black') +
       ggplot2::geom_density(alpha = 1, show.legend = FALSE) +

### 2023-08-14 paper/paper.Rmd
Add files via upload
fixed typos

@@
 \begin{equation}
-Sd = \sqrt\frac{\sum_{i=1}{(d_i-\bar{d})^2}}{n-1}
+SD = \sqrt\frac{\sum_{i=1}{(d_i-\bar{d})^2}}{n-1}
 \end{equation}
@@
 Overall this suggests that these two instruments tend to play earlier when compared to the isochronous beat division of the cycle, although the guitar is tightly aligned (<2%) with the reference beat on every fourth beat subdivisions (subdivisions of 1, 5, 9, and 13), which are also the salient metrical positions.
-Similar fluctuations across the beat subdivisions are not evident in the tres, but the graph reveals that the tres tends to play a little earlier than guitar.
+Similar fluctuations across the beat subdivisions are not evident in the tres, but the graph reveals that the tres tends to play a little earlier than the guitar.
 
@@
 
-Let's determine the overall synchrony between specific instruments. Here we continue the analysis guitar and tres and calculate the asynchrony between them.
+Let's determine the overall synchrony between specific instruments. Here we continue the analysis of the guitar and the tres and calculate the asynchrony between them.
 
@@
 This analysis indicates that on average, the tres plays 13 ms ahead of the guitar, with a standard deviation of 27 ms.
-When one is comparing instruments that have radically different number of joint onsets (such as clave and bass in this example), it is possible to make the comparisons between the pairs of instruments easier by specifying a sample of onsets that will be taken from each instrument for the analysis.
+When one is comparing instruments that have radically different number of joint onsets (such as the clave and the bass in this example), it is possible to make the comparisons between the pairs of instruments easier by specifying a sample of onsets that will be taken from each instrument for the analysis.
 It also possible to establish the confidence intervals by bootstrapping the asynchrony calculations.
@@
 
-In Figure 3 we see how different pairs of instruments have different asynchrony relationships; the bass is consistently ahead of the guitar, and clave and guitar play behind tres, to pick some of the extreme examples from the visualisation.
+In Figure 3 we see how different pairs of instruments have different asynchrony relationships; the bass is consistently ahead of the guitar, and the clave and the guitar play behind the tres, to pick some of the extreme examples from the visualisation.
 The vertical lines stand for median asynchrony and the violin plot shows the distribution of onset time differences.
 
-We can calculate the classic measures of synchronization once the pairing has been done. Here are these measures calculated for the synchrony between bass and clave.
+We can calculate the classic measures of synchronization once the pairing has been done. Here are these measures calculated for the synchrony between the bass and the clave.
 
@@
 We can also calculate the relative asynchrony by comparing the onset times to the mean of the other instruments that we define.
-Here we compare the asynchrony of bass and clave to the mean of the rhythm section as defined by four instruments, guitar, bass, tres, and clave.
+Here we compare the asynchrony of the bass and the clave to the mean of the rhythm section as defined by the mean onset times of the four instruments.
 
@@
 
-The relative measure (mean relative asynchrony) shows that bass is ahead (i.e.
--19 ms) of the mean of other instruments whereas clave is almost exactly in synchrony with the other instruments (<1 ms late).
+The relative measure (mean relative asynchrony) shows that the bass is ahead (i.e.
+-19 ms) of the mean of other the instruments whereas the clave is almost exactly in synchrony with the other instruments (<1 ms late).
 In different types of performances, musical pieces, instruments, and genres, the magnitude of these differences is highly variable [@clayton2020].

### 2023-09-09 README.Rmd
Improvements based on reviews
version 0.5.0

@@
 if (!require(devtools)) install.packages("devtools")
-devtools::install_github("tuomaseerola/onsetsync")
+devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
 ```
@@
 For more examples, see [Get started](https://tuomaseerola.github.io/onsetsync/articles/onsetsync.html).
+
+Read the [community guidelines](CONTRIBUTING.md) for contributing to the package or asking questions.

### 2023-09-09 paper/paper.Rmd
version 0.5.0

@@
 # LOADS THIS SECRECTLY FROM THE LOCAL FOLDER
-library(onsetsync)
+devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
+#library(onsetsync)
 ver <- packageVersion('onsetsync')
@@
 library(devtools)
-devtools::install_github("tuomaseerola/onsetsync")
+devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
 library(onsetsync)

### 2023-09-09 R/plot_by_beat.R

@@ -107,13 +107,8 @@ plot_by_beat <-
     
     ## PLOT with metrical grid deviations
     if (griddeviations == TRUE) {
-<<<<<<< HEAD
 #      S$label <- paste(round(S$M, digits = 0), '%', sep = '')
       S$label <- paste0(ifelse(S$M >= 0, "+", ""), round(S$M, digits = 0), "%")
-=======
-      S$label <- paste0(ifelse(S$M >= 0, "+", ""), round(S$M, digits = 0), "%")
-#      S$label <- paste(round(S$M, digits = 0), '%', sep = '')
->>>>>>> 5fd5cee830e6993950779dc6907839fff70bd9fc
       S$beat <- as.integer(S$beatF)
       
       

### 2023-09-09 paper/paper.Rmd
Update paper.Rmd
install options changes

--- 2023-08-14 paper/paper.Rmd
@@
 ```{r secret_load, eval=TRUE, echo=FALSE}
-# LOADS THIS SECRECTLY FROM THE LOCAL FOLDER
-library(onsetsync)
+devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
+#library(onsetsync)
 ver <- packageVersion('onsetsync')
@@
 library(devtools)
-devtools::install_github("tuomaseerola/onsetsync")
+devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE")
 library(onsetsync)

### 2023-09-09 paper/paper.Rmd
updated visuals on plot_by_beat

--- 2023-09-09 paper/paper.Rmd
@@
 devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
-#library(onsetsync)
+library(onsetsync)
 ver <- packageVersion('onsetsync')

### 2023-09-14 paper/paper.Rmd
suppressed installation message

--- 2023-09-09 paper/paper.Rmd
@@
 
-```{r secret_load, eval=TRUE, echo=FALSE}
-devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
-#library(onsetsync)
+```{r secret_load, eval=TRUE, echo=FALSE,message=FALSE,warning=FALSE}
+devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE,quiet=TRUE)
+library(onsetsync)
 ver <- packageVersion('onsetsync')
@@
 
-```{r availability,eval=FALSE}
+```{r availability,eval=FALSE,message=FALSE,warning=FALSE}
 library(devtools)
-devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE")
+devtools::install_github("tuomaseerola/onsetsync")
 library(onsetsync)

### 2023-10-10 R/add_isobeats.R
update to 0.5.1

@@ -1,13 +1,21 @@
-#' Add isochronous beats to the file
+#' Add isochronous times or mean onset times to the data frame
 #'
-#' Add isochronous beats to a onset datafame based on beat levels and average onset of the 1st beat
+#' Given one instrument, this is the reference time which adds isochronous beats to a
+#' onset dataframe based on beat levels and average onset of the 1st beat.
+#' Given two or instruments, the mean onset time is added as a reference time.
 #'
-#' @param df data frame to be processed
-#' @param instr variable which contains the instruments
-#' @param beat variable which contains the beats
-#' @param beatlabel variable which labels the newly created beats
+#' @param df data frame to be processed, where the minimal requirements are:
+#' (1) a reference time OR two or more instruments (instr input) and (2)
+#' beat sub-division (beat input)
+#' @param instr required element that specifies how the reference time is constructed:
+#' if given two or more instruments, will create timings based on mean onset times of the instruments.
+#' if given one input as an argument, this is assumed to be an external cycle time and
+#' the reference time is then inferred from beat sub-divisions, cycles and mean onset times 
+#' of the first beat. Required input.
+#' @param beat Beat sub-divisions (required)
+#' @param beatlabel variable which labels the newly created beats (optional)
 #' @seealso \code{\link{add_annotation}}
-#' @return dataframe with added column of isochronous beats and cycle numbers
+#' @return dataframe with added column of isochronous beats ("Iso.Time") OR mean onset times ("Mean.Time")
 #' @import ggplot2
 #' @import tidyr
 #' @import magrittr
@@ -15,49 +23,61 @@
 add_isobeats <- function(df = NULL,
                          instr = NULL,
                          beat = NULL,
-                         beatlabel = 'Isochronous.SD.Time'){
-
[…]

### 2023-10-10 R/gaussify_onsets.R

@@ -2,11 +2,11 @@
 #'
 #' This is a conversion function that turns onsets into sampled time-series.
 #'
-#' @param data Data frame to be processed with discrete time points (onsets)
+#' @param data Data frame to be processed with discrete time points (required)
 #' @param sr Sampling rate (default 500 Hz)
 #' @param wlen Window length of percentage of the sr (default 4)
-#' @param plot If a plot is needed
-#' @param time Time samples to be included in the output
+#' @param plot If a plot is needed (optional)
+#' @param time Time samples to be included in the output (optional)
 #' @return Output contain vector of onset times
 #' @import signal
 #' @import ggplot2

### 2023-10-10 R/period_to_BPM.R

@@ -2,8 +2,8 @@
 #'
 #' Converts period in seconds to BPM
 #'
-#' @param period data frame to be processed
-#' @param beats_per_minute baseline assumption of how many beats there are in min.
+#' @param period data frame to be processed (required)
+#' @param beats_per_minute baseline assumption of how many beats there are in minute (default 60)
 #' @seealso \code{\link{summarise_periodicity}},\code{\link{periodicity}}
 #' @return Numeric output
 #' @export

### 2023-10-10 R/periodicity.R

@@ -2,14 +2,14 @@
 #'
 #' This function plots the calculated asynchronies of instruments and labels these by instruments.
 #'
-#' @param df data frame to be processed
-#' @param instr Instrument name to be processed
+#' @param df data frame to be processed  (required)
+#' @param instr Instrument name to be processed  (required)
 #' @param method Periodicity analysis method: per (default), diff, acf, or fft
-#' @param sampling_rate Sampling rate (Hz)
-#' @param freq_range Frequency range to be included in the periodicity analysis (in seconds)
-#' @param resolution Resolution for some of the analyses (in seconds)
-#' @param colour Line colour for plotting 
-#' @param title Title for plotting
+#' @param sampling_rate Sampling rate (Hz), default 500 Hz
+#' @param freq_range Frequency range to be included in the periodicity analysis (in seconds), default 0 to 2
+#' @param resolution Resolution for some of the analyses (in seconds), default 0.01
+#' @param colour Line colour for plotting (optional)
+#' @param title Title for plotting (optional)
 #' @seealso \code{\link{summarise_periodicity}}
 #' @return Graphic output
 #' @export

### 2023-10-10 R/plot_by_dataset.R

@@ -2,13 +2,13 @@
 #'
 #' This function plots the calculated asynchronies of instruments and labels these by instruments.
 #'
-#' @param df data frame to be processed
-#' @param asynchronies Variable name where the asynchronies are
-#' @param data Variable name where the beats are
+#' @param df data frame to be processed (required)
+#' @param asynchronies Variable name where the asynchronies are (required)
+#' @param data Variable name where the beats are (required)
 #' @param pcols Number of columns for multiple plots (default 1)
-#' @param box Do the graphics by boxplot
-#' @param colour colour for the boxplot
-#' @param colorpalette colourpalette for the boxplot
+#' @param box Do the graphics by boxplot (optional)
+#' @param colour colour for the boxplot (optional)
+#' @param colorpalette colourpalette for the boxplot (optional)
 #' @seealso \code{\link{plot_by_variable}}, \code{\link{plot_by_pair}}, \code{\link{plot_by_beat}}
 #' @return Graphic output
 #' @import ggplot2
@@ -24,11 +24,25 @@ plot_by_dataset <-
            colour = 'lightblue',
            colorpalette = 'Set2') {
     # T. Eerola, Durham University, IEMP project
+    # 23/1/2018
+    # needs work
+    
+    # for testing
+    # df<-asere
+    # instr<-c('Bass','Tres','Bell')
+    # beat<-'SD'
+    # virtual<-'Virtual.SD'
+    # pcols<-2
+    # box<-TRUE
+    # griddeviations<-FALSE
+    #print(data)
+    #print(asynchronies)
     asynch <- n <- NULL
     
     DF <- dplyr::select(df, asynchronies, data)
     colnames(DF) <- c('asynch', 'data')
-
+#    print(head(DF))
+    
     g1 <-
       ggplot2::ggplot(DF, ggplot2::aes(asynch, fill = data), colour = 'black') +
       ggplot2::geom_density(alpha = 1, show.legend = FALSE) +

### 2023-10-10 R/plot_by_var_time.R

@@ -4,15 +4,15 @@
 #' calculated asynchronies of instruments across time and
 #' another variable.
 #'
-#' @param df data frame to be processed
-#' @param var1 Variable 1 (Time)
-#' @param var2 Variable 2
+#' @param df data frame to be processed (required)
+#' @param var1 Variable 1 (Time)  (required)
+#' @param var2 Variable 2 (required)
 #' @param var3 Variable 3 (optional)
-#' @param colour Colour option
-#' @param smoothline_colour Colour option
-#' @param xlabel label for X axis
-#' @param ylabel label for Y axis
-#' @param zlabel label for Z axis
+#' @param colour Colour option (optional)
+#' @param smoothline_colour Colour option (optional)
+#' @param xlabel label for X axis (optional)
+#' @param ylabel label for Y axis (optional)
+#' @param zlabel label for Z axis (optional)
 #' @return Graphic output
 #' @seealso \code{\link{sync_sample_paired}} for synchronies between
 #' instruments, \code{\link{plot_by_beat}} for plotting.

### 2023-10-10 R/plot_timeline.R

@@ -2,8 +2,8 @@
 #'
 #' This function plots the calculated asynchronies of instruments and labels these by instruments.
 #'
-#' @param data data frame to be processed
-#' @param instr Instrument name to be processed
+#' @param data data frame to be processed (required)
+#' @param instr Instrument name to be processed (required)
 #' @param ampl Amplitude of the onsets (optional)
 #' @param grouping Grouping factor (optional)
 #' @seealso \code{\link{plot_by_variable}}

### 2023-10-10 R/summarise_onsets.R

@@ -4,12 +4,12 @@
 #' summarises the time difference between successive onsets
 #' across instruments.
 #'
-#' @param df data frame to be processed
-#' @param instr Instrument names to be processed
-#' @param filter_lower Lower limit for onset differences to be summarised
-#' @param filter_upper Upper limit for onset differences to be summarised
-#' @param binw bin width for histogram
-#' @param plot Option for plotting (FALSE default)
+#' @param df data frame to be processed (required)
+#' @param instr Instrument names to be processed  (required)
+#' @param filter_lower Lower limit for onset differences to be summarised (default 0)
+#' @param filter_upper Upper limit for onset differences to be summarised (default 2)
+#' @param binw bin width for histogram (default 0.25)
+#' @param plot Option for plotting (defaul FALSE)
 #' @return Output contain summarises of the onset differences
 #' @import ggplot2
 #' @import tidyr

### 2023-10-10 R/summarise_sync_by_pair.R

@@ -2,7 +2,8 @@
 #'
 #' This function calculates simple t-test of the asynchronies of instruments (whether they differ from zero)
 #'
-#' @param df data frame to be processed
+#' @param df data frame to be processed, where the minimal requirements are:
+#' (1) two or more instruments (`instr`) and (2) beat sub-division (`beat`)
 #' @param bybeat logical value whether beat information is used
 #' @param adjust Adjust p values with Bonferroni correction (default)
 #' @seealso \code{\link{sync_sample_paired}}, \code{\link{sync_execute_pairs}}

### 2023-10-10 R/sync_sample_paired.R

@@ -2,9 +2,9 @@
 #'
 #' taken N samples of two instruments (where they both have onsets)
 #'
-#' @param df data frame to be processed
-#' @param instr1 Instrument 1 name to be processed
-#' @param instr2 Instrument 2 name to be processed
+#' @param df data frame to be processed (required)
+#' @param instr1 Instrument 1 name to be processed (required)
+#' @param instr2 Instrument 2 name to be processed (required)
 #' @param n Number of samples to be drawn from the pool of joint onsets. If 0, do not sample (default)
 #' @param bootn How many bootstraps are drawn (default none, which is 1)
 #' @param beat Beat structure (subdivisions) to be included

### 2023-10-10 R/synthesise_onsets.R

@@ -2,8 +2,8 @@
 #'
 #' Take the specified instruments and create wave file with sounds where the onsets are
 #'
-#' @param data data frame to be processed
-#' @param instruments Instruments names to be included
+#' @param data data frame to be processed (required)
+#' @param instruments Instruments names to be included (required)
 #' @param sr sampling rate (default 22050)
 #' @param dur duration of sounds (default 30 ms)
 #' @param type type of sounds (default noise, try click or synth from seewave)

### 2023-10-12 paper/paper.Rmd

--- 2023-09-09 paper/paper.Rmd
@@
 
-```{r secret_load, eval=TRUE, echo=FALSE}
-devtools::install_github("tuomaseerola/onsetsync@R1",force = TRUE)
-#library(onsetsync)
+```{r secret_load, eval=TRUE, echo=FALSE,message=FALSE,warning=FALSE}
+devtools::install_github("tuomaseerola/onsetsync",force = TRUE, quiet=TRUE)
+library(onsetsync)
 ver <- packageVersion('onsetsync')
@@
 
-```{r availability,eval=FALSE}
+```{r availability,eval=FALSE,message=FALSE,warning=FALSE}
 library(devtools)
